---
title: "Demographic Biases in Naturalistic Language Recordings in the CHILDES Database"
shorttitle        : "WEIRD CHILDES"

author: 
  - name          : "Camila Scaff*"
    affiliation   : "1,2"
  - name          : "Georgia Loukatou*"
    affiliation   : "2"
  - name          : "Alejandrina Cristia"
    affiliation   : "2"
  - name          : "Naomi Havron"
    affiliation   : "3"
    
affiliation:
  - id: "1"
    institution: University of Zurich, Institute of Evolutionary Medicine (IEM), Switzerland
  - id: "2"
    institution: PSL University, Laboratoire de Sciences Cognitives et de Psycholinguistique (ENS, EHESS, CNRS, DEC), France
  - id: "3"
    institution: Haifa Univerity, Israel



  # - code: "2"
  #   address: ENS, EHESS, CNRS, PSL University, Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d études cognitives, 29 rue d ulm, Paris, 75005, France

# corresp_author_name:  "ANON"
# corresp_author_email: "ANON"

abstract: |
  In recent years, the importance of estimating demographic biases in research has become apparent. Here we provide a systematic review of the CHILDES database, the major source of naturalistic recordings of children’s linguistic environment. We analyzed the database according to four dimensions considered central to language learning: SES, urbanization, family structure, and language. We present descriptive statistics of each dimension to assess whether naturalistic recordings were biased regarding the demographics of the countries and the families recorded within them. We find that CHILDES’s recordings overrepresented wealthier countries and higher parental education levels; urban settings; and smaller households. Middle- and higher-class participants were likewise over-represented. The corpora were not representative of their countries in terms of urbanization either - with a larger percentage of families residing in urban settings than is overall true for their respective countries. In terms of family structure, nuclear families were more prevalent than in the countries where the data was collected. Last, we found that corpora were linguistically diverse, but we estimate that these recordings underrepresented bilingual and multilingual households. We conclude that researchers should be mindful when generalizing from naturalistic recordings of children's input and output obtained from CHILDES, and make recommendations for the future of the use of CHILDES. 



## Remove this if not required
funding: |
  AC acknowledges financial and institutional support from Agence Nationale de la Recherche (ANR-16-DATA-0004 ACLEW, ANR-17-EURE-0017, ANR-14-CE30-0003 MechELex); the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award; and the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (ExELang, Grant agreement No. 101001095). GL acknowledges financial support from Fondation Fyssen.




## Remove this if not required
acknowledgements: |
  We are grateful to Lara Oliel and Clara Daoud for help entering corpus-level data; Sara Pisani for organizing communication with corpus curators; Brian MacWhinney for support and encouragement when retrieving information from corpus curators. We particularly want to thank the literally hundreds of researchers who contributed data to CHILDES, without which this project would be unthinkable; and the 103 curators who replied to our request.

# ac2cs shall we add their names? Do we have permission for that?


keywords          : "childes, verbal input, infant-directed speech, language, weird"


bibliography      : ["childes.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_doc
---


```{r setup, include=FALSE}
Sys.setlocale("LC_CTYPE", "en_US_POSIX.UTF-8")
#Sys.setlocale("LANG", "en_US_POSIX.UTF-8")
#Sys.setlocale(category = "LC_ALL", locale = "en_SE.utf8")

knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE)
# Install remotes package if necessary
if(!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")

# Install the stable development version from GitHub
#remotes::install_github("crsh/papaja")
library("papaja")
library(psych)
library(car)
library(ggplot2)
library(ggthemes)
library(scales)
library(gridExtra)
library(lme4)
library(ggpubr)
library(lmerTest)

library(knitr)
options(digits=1)
library(kableExtra)
library(dplyr)
library(effects)
library(gvlma)
library(influence.ME)
library("wesanderson")
#devtools::install_github("psyteachr/introdataviz")
#library(ggalt)
library(Hmisc)

annotations_inc<- read.csv("../derived/annotations_included", sep="")
ind<- read.csv("../derived/ind", sep="")
ind_all<- read.csv("../derived/ind_all", sep="")
ocde <- read.csv("../data/ocde_country.csv", sep=",")


##countries and corpora
corpora <- as.data.frame((table(annotations_inc$country)))
names(corpora)<-c("country","Freq")
#write.table(corpora,"derived/corpora", row.names = FALSE, col.names = TRUE)


merge(x=corpora, y=ocde,  by = "country", all.x = T) -> ocde_country

annotations<- read.csv("../data/Table for authors - annotations.csv") 
#original link https://docs.google.com/spreadsheets/d/1s-ytfQf7WsZFDDZ6QkOQnZhTodHpQp7D2Wja8YFvjQY/edit?usp=sharing

#some cleaning  of the col names
colnames(annotations)=gsub("\\.\\..*","",colnames(annotations))
colnames(annotations)[colnames(annotations)=="Number.of.participants"]<-"Nb.of.participants"
colnames(annotations)[colnames(annotations)=="Language.or.Languages.spoken.in.recordings..be.specific.if.possible.e.g..French.Quebec."]<-"Language"
colnames(annotations)[colnames(annotations)=="Language.or.Languages.spoken.in.recordings..be.specific.if.possible.e.g..French.Quebec."]<-"Language"
colnames(annotations)[colnames(annotations)=="Location..Neighbourhood..village.city..province..state..country."]<-"Location"
colnames(annotations)[colnames(annotations)=="X..of.children.with.siblings"]<-"Nb.of.children.with.siblings"
colnames(annotations)[colnames(annotations)=="X..of.children.with.older.siblings"]<-"Nb.of.children.with.older.siblings"
colnames(annotations)[colnames(annotations)=="Our.coding.of"]<-"Our.coding.of.community.type"

#Inclusion
annotations$Inclusion[annotations$Inclusion %in% c("No","no")] <- "exclude"
annotations$Inclusion[annotations$Inclusion %in% c("Yes","yes")] <- "include"
#why_exclude
excluded <-  subset(annotations, (Inclusion %in% c("exclude")))
#xtabs(~why_exclude, excluded)




```

## Research highlights

* We examined CHILDES, the major source of data on children’s linguistic environment, for potential sampling bias in terms of SES, urbanization, family structure, and language.
* We found that the 48 countries represented in CHILDES overrepresented higher-GDP, educated, urbanized countries with small nuclear families.
* Within these countries, corpora overrepresented higher-SES, educated, urban, nuclear, and monolingual families.
* Interpretation of studies using CHILDES should acknowledge these biases.


\newpage

  Since its foundation in 1984, CHILDES [the Child Language Data Exchange System, @macwhinney2000childes] has been the major source of naturalistic recordings and transcript data for researchers studying language acquisition. Moreover, it has contributed to establishing seminal work in various domains such as Theory of Mind [@bartsch1995children] and memory [@anderson1991reflections]. It has also inspired theories that help understand the connection between language input and development [@christiansen1998learning]. As such, CHILDES has been and is one of the most valuable assets of our field. While much of previous work relied on one or a handful of corpora, even when all relevant corpora are used [@ibbotson2024effects; @tan2024predicting], conclusions are more likely to be generalizable if they are based on representative data. We examine the extent to which CHILDES' naturalistic recordings, containing information about both input and output, come from a representative sample of the world's populations.
 
Researchers in many fields are increasingly aware of the bias their samples may have towards WEIRD populations [i.e., Western, Educated, Industrial, Rich, Democratic, @henrich2010most; @nielsen2017persistent;@cychosz2022using ;@moriguchi2022beyond; @singh2023diversity]. Recent calls have been made to diversify research in psychology and developmental science to address this issue [@Majid2010; @kidd2022diverse; @blasi2022over]. For instance, @kidd2022diverse systematically reviewed publications from child-language journals. They revealed biases towards specific continents (North America and Europe) and languages (English, Spanish, French), highlighting the need for increased diversity in the populations and languages studied.
	
Every area of research conceptualizes different dimensions relevant to explaining variance for a given research topic. In what follows we present the four most relevant dimensions thought to play a role in language acquisition, particularly related to naturalistic recordings: socioeconomic status, urbanization, family structure, and language.


### Socioeconomic Status
It is beyond the scope of this paper to detail all the theories that attempt to account for the complex causal pathways that may connect socio-economic status (SES) to children's language environments [see @rowe2018understanding for a discussion, and @golinkoff2019language; @sperry2019reexamining for diverse theoretical perspectives]. While we do not take a stance on whether, how, or why SES and language environments relate to each other, SES is one of the factors that has been repeatedly studied in the context of early language acquisition, and as such, deserves attention. For example, @hoff2003specificity compared the speech of high- versus middle-SES American mothers. College-educated mothers produced more utterances to their children, with more diverse vocabulary, longer phrases, and a higher number of utterances continuing a topic the child had brought up, which some claim to also be related to language outcomes [see also @hart1995meaningful;@hoff1990maternal;@hoff2014causes; @huttenlocher2002language; @huttenlocher2007varieties; see also @dailey2022language; @leonardo2022socioeconomic; for meta-analyses supporting the link; and @bergelson2023everyday for a large-scale study finding no SES effects]. 

In the developmental literature, SES is primarily indexed by parents’ education [@ensminger2014decade; @hoff2003specificity], but can also be indexed by parental income, occupation, or a composite measure of these three [e.g. using the Hollingshead index, @hollingshead1975four].


### Urbanization
This dimension represents differences across populations in terms of the extent to which they are organized around urban sites, which often entail differences in job opportunities, living conditions, and infrastructure. There have been proposals that societies varying in their urbanization have differing views and values about the role of children in conversations [e.g., @sharma1998child ; @richman1992cultural; @draper2017parent; @keller2012autonomy]. For instance, @keller2012autonomy discusses three prototypical cases, urban, rural, and hybrid, which differ in terms of their goals for children: Urban families aim for child psychological independence, rural families for child physical autonomy and interdependence, and hybrid families for some mix. @vogt2020multimodal found that the number of gestures, gesture-speech alignment, and gesture types all vary across the three groups in ways that can be related to Keller's typology, in urban Dutch, urban Mozambique, and rural Mozambique. Similarly, @cristia2023systematic systematic review uses urbanization to describe the frequency of child-directed vocalization, finding that it is far less frequent in rural than urban communities.

### Family structure
In this paper, we use the concept of family structure to encompass a variety of related aspects such as birth order and the way that caregiving is structured, each of which has an impact on child and language development [e.g., @blake1981family; @duncan2020home; @havron2022effect; @havron2019effect; @hoff1991older; @tomasello1985pragmatics; @bornstein2019mother]. For instance, children with older siblings exhibit lower language skills compared to first-born children across various cultures [e.g., @peyre2016differential in France; @havron2022effect for Singapore; @zambrana2012impact for Norway], although second-born children might benefit from overhearing conversations between their caregivers and older sibling [@oshima1996birth]. 
As for caregiving structures, in many middle-class Euro-American families, parents typically assume primary responsibility for children, with the mother as the primary caregiver [e.g., @bianchi2010work]. However, certain cultures, like the Turkish families described by @isleyen2021marital, adopt a different approach, with nuclear families living in separate apartments but sharing common spaces and caregiving responsibilities, resulting in extensive support networks.

### Languages 
Linguistic factors likely influence the development of language skills in children [@kidd2022diverse; @blasi2022over]. For example, based on transcriptions of conversations, Pye and colleagues documented that K’iche' Mayan children frequently use and understand passive constructions from a very young age, unlike their English-speaking peers, refuting the idea that passive constructions can only emerge later in development [@pye1980acquisition; @pye1988precocious].
Similarly, many Indo-European languages show a strong noun bias in early vocabularies (a bias for acquiring words for concrete referential objects rather than actions), which has been claimed to be a universal feature of early language acquisition. However, studies have shown that in some Mayan languages, including Tseltal and Tsotsil [@casillas2024little; @de1999verbs], there is a verb bias instead. Thus, generalizations about language acquisition could be inaccurate when they rely on a linguistically biased database.    	
Another major dimension is the study of bilingualism or multilingualism [for an introductory review see @hohle2020variability, and @mccabe2013multilingual]. There is evidence that monolingual and bilingual early language development differ in some aspects, particularly regarding phonological acquisition and word learning. For example, monolingual infants' ability to discriminate non-native sounds declines during the first year of age, whereas infants exposed to more language(s) maintain the discrimination window for a longer period [@singh2022diversity]. In terms of input, bilingual linguistic exposure is divided between two or more native languages. It has been shown that the amount of exposure to each native language can affect bilingual infants' speech discrimination abilities [@garcia2011bilingual].


### The current study
Inspired by calls to document the demographic and linguistic diversity of participant samples [@kidd2022diverse; @singh2023unified], we sought to provide a systematic analysis of the naturalistic speech corpora in the CHILDES database by quantifying the diversity of each dimension introduced above (SES, urbanization, family structure, and languages). Following @ghai2021s recommendation to enhance the description of diversity in behavioral sciences, we examined different levels, from a macro-level, involving broad country comparisons, to a corpus-level, where we analyzed individual corpora for a more detailed and nuanced understanding of the data.

# Methods

### Inclusion criteria
We extracted information from the online CHILDES database in 2021, at which point there were 430 corpora. We excluded corpora on clinical populations (N = 69), in-lab recordings (N = 26), and elicitation tasks (N = 114), because they would have introduced additional sources of variation and biases. For example, the inclusion of clinical populations may highlight differences in access to diagnosis in different countries. Given our interest in naturalistic input and output represented in CHILDES, we also excluded corpora where only child speech was transcribed (N = 19) and diary studies (N = 11). Finally, we excluded 11 corpora that were not available, resulting in a final sample of 180 corpora containing naturalistic recordings of children's input and output.

### Demographic Data Collection
Data on participant demographics was extracted from the references provided for each corpus in CHILDES such as articles, book chapters, dissertations, and the actual content of transcribed recordings (for more information please see SM1; Anonymized, 2024). When key variables could not be determined from these sources, we contacted corpus curators for missing information, and over a third (103 corpora, 32%) provided additional data or confirmed precoded information. They were also invited to correct our coding if it was inaccurate to their knowledge, and we took their coding as definitive. Unfortunately, curators for 50 corpora (15%) could not be contacted due to unresponsive email addresses (38 corpora), or the curator’s passing (12 corpora).

### Analyses
Analyses and visualizations were carried out using R (version 4.4.2, R Core Team, 2020) and ggplot2 [@wickham2016getting]. Data, scripts, and Supplementary Materials are available on the (OSF)[https://osf.io/q9w82/].


# Results

```{r demoinfo, echo = FALSE, eval=T}

#summary of counts by country , Nb.of.participants and region

annotations_inc$Corpus[annotations_inc$Language == "Spanish/Catalan" & annotations_inc$Corpus == "Vila"] <- "Vila2" #changed for n utteraance graph

corpora.ind <- annotations_inc %>%
 select(country, Corpus, Nb.of.participants, region)

corpora.ind$region[corpora.ind$country =="Spain & Hungary"] = "Europe"
corpora.ind$region[corpora.ind$country =="Sweden & Portugal"] = "Europe"

#summary of counts by country 

summary_corpora_COUNTRY <- corpora.ind %>%
 group_by(country) %>%
 summarise(
  TimesRepeated = n(),
  TotalParticipants_country = sum(Nb.of.participants)
 )

#summary of counts by continent 
summary_corpora_CONT <- corpora.ind %>%
 group_by(region) %>%
 summarise(
  TimesRepeated = n(),
  TotalParticipants_continent = sum(Nb.of.participants)
 )

#categorize siblings - singletons
annotations_inc$siblings[annotations_inc$Nb.of.children.with.siblings != 0] <- "yes"
annotations_inc$siblings[annotations_inc$Nb.of.children.with.siblings == 0] <- "no"
annotations_inc$siblings[annotations_inc$Nb.of.children.with.siblings %in% c("all firstborn (some newborn sibs)", "at least 1 (triplet)")] <- "yes"
annotations_inc$siblings = factor(annotations_inc$siblings, levels=c("yes", "no"))


#some cleaning for OCDE counts
merge(x=annotations_inc,y=ocde_country,by= "country",all=T) -> ocde_inc
ocde_inc$ocde[ocde_inc$country =="Spain & Hungary"] <- "yes"
ocde_inc$ocde[ocde_inc$country =="Sweden & Portugal"] <- "yes"
ocde_inc$ocde[ocde_inc$country =="Hong Kong"] <- "no"

# changing order for count purposes
annotations_inc$Language[annotations_inc$Language == "Catalan/Spanish"] <- "Spanish/Catalan"

annotations_inc$Language[annotations_inc$Language == "Spanish/English"] <- "English/Spanish"

freq_languages <- as.data.frame(table(annotations_inc$Language))

lang_dif <- annotations_inc %>%
 group_by(Language) %>%
 summarise(Total_Participants = sum(Nb.of.participants, na.rm = TRUE))


#fixing spellings so that childes db finds the corpus

corpora.ind$Corpus2 <- corpora.ind$Corpus
corpora.ind$Corpus2[corpora.ind$Corpus =="De Houwer/Bornstein"] = "DeHouwerBornstein"
corpora.ind$Corpus2[corpora.ind$Corpus =="Chromá" ] = "Chroma"
corpora.ind$Corpus2[corpora.ind$Corpus == "Leveillé"] <- "Leveille"
corpora.ind$Corpus2[corpora.ind$Corpus == "PhonBLA"] <- "PhonBLA" ##not available trancript transcripts
corpora.ind$Corpus2[corpora.ind$Corpus == "PAIDUS-German"] <- "PAIDUS_German" ##not available trancript
corpora.ind$Corpus2[corpora.ind$Corpus == "Miipro"] <- "MiiPro"
corpora.ind$Corpus2[corpora.ind$Corpus == "Kern-Romanian"] <- "KernRomanian"
corpora.ind$Corpus2[corpora.ind$Corpus == "EsteveGibert/Prieto"] <- "EstevePrieto"
corpora.ind$Corpus2[corpora.ind$Corpus == "PAIDUS Spanish"] <- "PAIDUS_Spanish" #not available trancript
corpora.ind$Corpus2[corpora.ind$Corpus == "LlinasOjea (IRENE)"] <- "LlinasOjea"
corpora.ind$Corpus2[corpora.ind$Corpus == "Andrén"] <- "Andren"
corpora.ind$Corpus2[corpora.ind$Corpus == "CRSLP-MARCS"] <- "CRSLP"
corpora.ind$Corpus2[corpora.ind$Corpus == "Providence "] <- "Providence"
corpora.ind$Corpus2[corpora.ind$Corpus == "Demetras - Trevor"] <- "Demetras1"
corpora.ind$Corpus2[corpora.ind$Corpus == "Feldman (Andrea) "] <- "Feldman"
corpora.ind$Corpus2[corpora.ind$Corpus == "Demetras - Working"] <- "Demetras2"
corpora.ind$Corpus2[corpora.ind$Corpus == "Warren-Leubecker "] <- "Warren"
corpora.ind$Corpus2[corpora.ind$Corpus == "Peters/Wilson"] <- "Peters"
#corpora.ind$Corpus2[corpora.ind$Corpus == "Rondal-TD"] <- "TD" #missing- cant find from others
#rondal=subset(transcripts_, (transcripts_$corpus_name %in% "TD")) problem with corpus name is mixed with other ones

#same name of corpus : "Vila | Vila" ; "LlinasOjea | "LlinasOjea (IRENE) ; "	Serra & Sole| SerraSole"
#creating a different name corpus so that they are not counted twice
corpora.ind$Corpus2[corpora.ind$Corpus =="De Houwer" ] = "De Houwer"#not fixing because DeHouwer do not distinguishable from the other Dehouwer corpora - and this would count twice its values and at the end the all counts for Beligum
corpora.ind$Corpus2[corpora.ind$Corpus == "Serra & Sole"] <- "Serra & Sole"
corpora.ind$Corpus2[corpora.ind$Corpus == "LlinasOjea (IRENE)"] <- "LlinasOjea (IRENE)"
#+vila


```


### Do Countries in Our CHILDES Sample Represent the World's Countries?

```{r worldmap, echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Geographical Distribution of the Corpora in CHILDES. Each circle represents a corpus, and its size is proportional to the number of participants."}

library(maps)
## Figure
# Creating a subset with specific columns
map_info <- annotations_inc[, c("country", "Corpus", "Nb.of.participants", "Location")]
#write.table(map_info, file = "derived/map_info.txt", sep = "\t", quote = FALSE)

country_coords<- read.csv("../data/map_coordinates.csv", sep=",")

world_map <- map_data("world")

ggplot() +
 geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "lightgray") +
 geom_point(data = country_coords, aes(x = Longitude, y = Latitude, size = Nb.of.participants), color = "blue", alpha = 0.6) +
 theme_void() +
 labs(title = "Participants by Country")

ggsave("../figures/childes_country_participants.png", width = 6, height = 4, dpi = 300)

```

The `r nrow(annotations_inc)` corpora included in this study represent `r length(levels(as.factor(annotations_inc$country)))-2` <!-- minus 2 bilingual corpora counting twice Hungary/ Spain and Sweden/Portugal --> different countries or territories across all populated continents (see Figure 1). `r sum(ocde_country$ocde == "yes", na.rm = TRUE)` countries, representing `r sum(ocde_inc$ocde == "yes",na.rm=T)` corpora (`r round((sum(ocde_inc$ocde == "yes",na.rm=T)*100)/nrow(annotations_inc))`%), belong to the OECD (Organisation for Economic Co-operation and Development). Meanwhile, OECD countries represent only 19.5% of the world’s countries. The country with the most corpora is the United States with `r length(annotations_inc$country[annotations_inc$country == "United States"])` different corpora, followed by Spain with `r length(annotations_inc$country[annotations_inc$country == "Spain"]) + 1`<!-- +1 bilingual corpora -->, followed by the United Kingdom and France with `r length(annotations_inc$country[annotations_inc$country == "France"])`.

Analysis of the number of utterances per country presents a different picture, with the United Kingdom having the largest number of utterances, followed by the United States, Japan, Indonesia and Spain (see Figure 2).


```{r getnutterances, echo = FALSE, eval=T}

## Figure n uterrances
# library(childesr)
# library(stringr)
# #
# 
# # takes a long time to retrieve !!!
#  #use this code below to reproduce the corpora_utterances.csv
#  #if you want to reproduce the graph just use the corpora_utterances.csvfile
#  transcripts_ <-get_participants()
# 
# # Creating a subset with specific columns included in paper
#  include_thesecorp<- corpora.ind$Corpus2
#  ddd=subset(transcripts_, (transcripts_$corpus_name %in% include_thesecorp))
# 
# # create loop for all utterances of subset of corpora
#  utterances_childes<- NULL
#  for (name in include_thesecorp) {
#   print(paste("Analyzing corpus:", name))  #Print the corpus name
#   countutts <- get_utterances(corpus = name)
#   tabletmp <- countutts
#   utterances_childes <- rbind(utterances_childes, tabletmp)
#  }
# 
# # summarise by corpus
#   summary_table_childesutts <- utterances_childes %>%
#    group_by(corpus_name) %>%
#    summarise(count = n())  #Count the number of utterances per corpus
# 
# # identidy missing corpora
#  missing_corpora <- setdiff(corpora.ind$Corpus2, summary_table_childesutts$corpus_name)
# 
#  if (length(missing_corpora) == 0) {
#   print("All corpora are present in summary_table")
#  } else {
#   print("Missing corpora:")
#   print(missing_corpora)
#  }
# 
#  #check
#  summary_table_childesutts$count[summary_table_childesutts$corpus_name=="Remedi"]
#  get_utterances(corpus="Remedi")
# 
# write.csv(utterances_childes,"../derived/utterances_childes.csv",row.names=F)
# write.csv(summary_table_childesutts,"../derived/corpora_utterances.csv",row.names=F)

 ## Count words in each utterance (split by space)
 #  wow$word_count <- str_count(wow$gloss, "\\S+")
 
 # # Sum the total number of words
 #  total_words <- sum(wow$word_count, na.rm = TRUE)

# MISSING CORPORA:  

#info about utternaces for 172 corpus :
#missing 8: "Vila | Vila" ; "LlinasOjea | "LlinasOjea (IRENE) ; "	Serra & Sole| SerraSole" - ""De Houwer" doubled in name so +4
#problem with name in childessb RondatlTD +1
#no trancripts available for PAIDUS_Spanish | PAIDUS_German | PhonBLA + 3
# so 172 + 8 = 180

```

```{r plotnutterances, echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Number of utterances per country in CHILDES."}


childes_utts<- read.csv("../derived/corpora_utterances.csv", sep=",")

merge(x=corpora.ind, y=childes_utts, by.x="Corpus2", by.y="corpus_name") -> nutts

summary_corpora_country_utt <- nutts %>%
 group_by(country) %>%
 summarise(
  TimesRepeated = n(),
  Totalutts_country = sum(count)
 )

summary_corpora_country_utt <- summary_corpora_country_utt[order(summary_corpora_country_utt$Totalutts_country, decreasing = TRUE),] 

options(scipen = 999)
ggplot(summary_corpora_country_utt, aes(x = reorder(country, Totalutts_country), y = Totalutts_country)) +
  geom_bar(stat = "identity", fill = "blue",width = 0.8) +
  scale_y_discrete(expand = expansion(mult = 0.5)) +  # Increase space between bars
  coord_flip() +
  scale_y_continuous( breaks = c(500000, 1000000, 1500000, 2000000), labels = c("500k", "1000k", "1500k", "2000k"))+
  labs(title = "",
  x = "Country",
  y = "Total Number of Utterances") +
  theme_minimal()

 
ggsave("../figures/childes_utterancesbycountry_bycorpus.png", width = 6, height = 4, dpi = 300)


```


```{r densityplot, echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Comparison of Country-Level Descriptors Between World Countries and Our CHILDES Sample. Note. Density plots show the distribution of country-level descriptors for all world countries (gray) versus countries represented in our sample of CHILDES (orange). % compl. LSS is the percentage of the country's population completing lower secondary school. % urban is the percent of the country's population residing in urban (as opposed to rural) locations. GDP stands for Gross Domestic Product. Avg Household size stands for average household size. To clarify, the calculations are based on unweighted country-level statistics (each country contributes equally to the distribution irrespective of population size). For example, panel A indicates that, worldwide, only a minority of countries have over 90% completion of lower secondary school, whereas in the CHILDES sample, the density curve shows a higher concentration of countries in this range."}

library(ggplot2)


#GDP
dens_dat=data.frame(rbind(cbind(ind_all$log_gdp[!is.na(ind_all$log_gdp)],0),
             cbind(ind$log_gdp[!is.na(ind$log_gdp)],1) ))
colnames(dens_dat)<-c("GDP.per.capita","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
gdpt = t.test(dens_dat[,1]~dens_dat$in.childes)

rich <- ggplot(dens_dat, aes(x = GDP.per.capita, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "GDP per capita (log 10)")+ scale_fill_colorblind()


#Education
dens_dat=data.frame(rbind(cbind(ind_all$Compl.LS[!is.na(ind_all$Compl.LS)],0),
             cbind(ind$Compl.LS[!is.na(ind$Compl.LS)],1) ))
colnames(dens_dat)<-c("completed.LSS","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
edt = t.test(dens_dat[,1]~dens_dat$in.childes)

ed <- ggplot(dens_dat, aes(x = completed.LSS, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = c(0.25, 0.8), axis.title.y=element_blank() , legend.title = element_blank())+labs( x = "% compl. LSS")+ scale_fill_colorblind()

#urbanization
dens_dat=data.frame(rbind(cbind(ind_all$SP.URB.TOTL.ZS[!is.na(ind_all$SP.URB.TOTL.ZS)],0),
             cbind(ind$SP.URB.TOTL.ZS[!is.na(ind$SP.URB.TOTL.ZS)],1) ))
colnames(dens_dat)<-c("pc.urban","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
urbt = t.test(dens_dat[,1]~dens_dat$in.childes)

urb <- ggplot(dens_dat, aes(x = pc.urban, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() )+labs( x = "% urban")+ scale_fill_colorblind()

#Family structure
#% nuclear family
ind_all$nuclear = (ind_all$Couple.with.children + ind_all$Single.parent.with.children)
ind$nuclear = (ind$Couple.with.children + ind$Single.parent.with.children)

dens_dat=data.frame(rbind(cbind(ind_all$nuclear[!is.na(ind_all$nuclear)],0),
             cbind(ind$nuclear[!is.na(ind$nuclear)],1) ))
colnames(dens_dat)<-c("nuclear","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
nuclear = t.test(dens_dat[,1]~dens_dat$in.childes)

NUCLEAR <- ggplot(dens_dat, aes(x = nuclear, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "% Nuclear household with children")+ scale_fill_colorblind()

#% non nuclear family
ind_all$nnuclear = (ind_all$Extended.family + ind_all$Non.relatives)
ind$nnuclear = (ind$Extended.family + ind$Non.relatives)

dens_dat=data.frame(rbind(cbind(ind_all$nnuclear[!is.na(ind_all$nnuclear)],0),
             cbind(ind$nnuclear[!is.na(ind$nnuclear)],1) ))
colnames(dens_dat)<-c("nnuclear","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
nnuclear = t.test(dens_dat[,1]~dens_dat$in.childes)

NNUCLEAR <- ggplot(dens_dat, aes(x = nnuclear, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "% Extended or Non-relatives in the household")+ scale_fill_colorblind()


#Average household size
dens_dat=data.frame(rbind(cbind(ind_all$Avg.hs[!is.na(ind_all$Avg.hs)],0),
             cbind(ind$Avg.hs[!is.na(ind$Avg.hs)],1) ))
colnames(dens_dat)<-c("Avg.hs","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
hh = t.test(dens_dat[,1]~dens_dat$in.childes)

HH <- ggplot(dens_dat, aes(x = Avg.hs, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "Avg Household size")+ scale_fill_colorblind()

#Average household size
dens_dat=data.frame(rbind(cbind(ind_all$Avg.und15[!is.na(ind_all$Avg.und15)],0),
             cbind(ind$Avg.und15[!is.na(ind$Avg.und15)],1) ))
colnames(dens_dat)<-c("Avg.und15","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
und15 = t.test(dens_dat[,1]~dens_dat$in.childes)

UND15 <- ggplot(dens_dat, aes(x = Avg.und15, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "Avg # members >15 years per household ")+ scale_fill_colorblind()


# combined
density_plot <- ggarrange(ed, urb, rich, HH, 
        labels = c("A", "B", "C", "D"),
        ncol = 2, nrow = 2)
density_plot

ggsave("../figures/density_plot.png", width = 6, height = 4, dpi = 300)
```

We first assessed the representativeness of our CHILDES sample by comparing country-level descriptors across - SES (education and GDP), urbanization, and family structure - against global distributions. We used public datasets from Our World in Data, the World Bank (World Development Indicators, WDI, Arel-Bundock, 2021), and the UN Household Size and Composition database (United Nations, 2022, see full details in SM1) to extract these global descriptors.

The countries in our sample of CHILDES differed significantly from the distribution of countries in the world on our key descriptors according to unpaired tests that did not assume equality of variances (Welch's t): a higher proportion of the population residing in our sample’s countries completed lower secondary school than the worldwide sample (% compl. LSS, t(`r round(edt$parameter,2)`)=`r round(edt$statistic,2)`; the countries were more urban (% urban, t(`r round(urbt$parameter,2)`)=`r round(urbt$statistic,2)`, p < .001) and richer (log GDP per capita, t(`r round(gdpt$parameter,2)`)=`r round(gdpt$statistic,2)`, p < .001); and they had smaller households (average household size, t(`r round(hh$parameter,2)`)=`r round(hh$statistic,2)`,p < .001; see Figure 3).<!--FIX P value reporting, currently broken -->

### How Representative and Diverse are Participant Samples?
We next investigated variables characterizing the specific families within each corpus, as far as possible given substantial missing data (see Table 1 and figures below for definitions of the variables and the number of corpora represented in each analysis, see SM1 for more details on variables). 


```{r tab1,echo = FALSE, as.is=T, ig.pos = "T!",results="asis"}
library(kableExtra)


#SES
SES = sum(!is.na(annotations_inc$SES.STDZD))
SES_missing =round(sum(is.na(annotations_inc$SES.STDZD))/length(annotations_inc$SES.STDZD)*100)
SES_present = 100- SES_missing
#ed
ED_ses = sum(!is.na(annotations_inc$Education.STDZD))
ED_ses_missing =round(sum(is.na(annotations_inc$Education.STDZD))/length(annotations_inc$Education.STDZD)*100)
ED_ses_present = 100- ED_ses_missing

#occupation
Ocup_ses = sum(!is.na(annotations_inc$Parental.profession))
Ocup_ses_missing =round(sum(is.na(annotations_inc$Parental.profession))/length(annotations_inc$Parental.profession)*100)
Ocup_ses_present = 100- Ocup_ses_missing

#urb Type of community
urbrur = sum(!is.na(annotations_inc$Type.of.community))
urbrur_missing =round(sum(is.na(annotations_inc$Type.of.community))/length(annotations_inc$Type.of.community)*100)
urbrur_present = 100- urbrur_missing


#family structure 
#Household composition
Household.structure = sum(!is.na(annotations_inc$Household.structure))
Household.structure_missing =round(sum(is.na(annotations_inc$Household.structure))/length(annotations_inc$Household.structure)*100)
Household.structure_present = 100- Household.structure_missing

#Percent children with sibling(s)", 
per_sibs = sum(!is.na(annotations_inc$Proportion.With.Siblings))
per_sibs_missing =round(sum(is.na(annotations_inc$Proportion.With.Siblings))/length(annotations_inc$Proportion.With.Siblings)*100)
per_sibs_present = 100- per_sibs_missing

#"Average number of siblings", 
avg_sibs = sum(!is.na(annotations_inc$Average.number.of.siblings))
avg_sibs_missing =round(sum(is.na(annotations_inc$Average.number.of.siblings))/length(annotations_inc$Average.number.of.siblings)*100)
avg_sibs_present = 100- avg_sibs_missing
#"Language(s) spoken", 
Langspoken = sum(!is.na(annotations_inc$Language))
Langspoken_missing =round(sum(is.na(annotations_inc$Language))/length(annotations_inc$Language)*100)
Langspoken_present = 100- Langspoken_missing
#"Lingual status
multiling = sum(!is.na(annotations_inc$Bilingualism.Multilingualism.in.corpus))
multiling_missing =round(sum(is.na(annotations_inc$Bilingualism.Multilingualism.in.corpus))/length(annotations_inc$Bilingualism.Multilingualism.in.corpus)*100)
multiling_present = 100- multiling_missing
#Language

intermediate_SES <- paste(SES, " (", round(SES_present), "%)", sep = "")
intermediate_ED <- paste(ED_ses, " (", round(ED_ses_present), "%)", sep = "")
intermediate_ocup <- paste(Ocup_ses, " (", round(Ocup_ses_present), "%)", sep = "")
intermediate_urb <- paste(urbrur, " (", round(urbrur_present), "%)", sep = "")
intermediate_HHS <- paste(Household.structure, " (", round(Household.structure_present), "%)", sep = "")
intermediate_percsib <- paste(per_sibs, " (", round(per_sibs_present), "%)", sep = "")
intermediate_avgsib <- paste(avg_sibs, " (", round(avg_sibs_present), "%)", sep = "")
intermediate_langspo <- paste(Langspoken, " (", round(Langspoken_present), "%)", sep = "")
intermediate_multiling <- paste(multiling, " (", round(multiling_present), "%)", sep = "")


# Create a data frame
mytab1 <- data.frame(
  Dimension = c("SES", "SES", "SES", "Urbanization", "Family structure", "Family structure", "Family structure", "Language", "Language"),
  `Corpus-level variable` = c("Parents' socio-economic status", "Parents' education level", "Parents' occupation", "Type of community", "Household composition", "Percent children with sibling(s)", "Average number of siblings", "Language(s) spoken", "Lingual status"),
  `Definition or levels` = c("Low, mid and/or high", "Highest level of education completed: Primary, High school, Undergraduate, Masters, PhD", "Parents' activity or profession", "Urban, rural, mixed", "Whether the family was composed primarily of caregivers and children (nuclear) or it also included other family members: grandparents, cousins, uncles, and/or aunts (extended)", "The percentage of children in the corpus who had at least one sibling", "How many siblings children had on average (including children with zero siblings)", "Which languages were spoken in the transcripts and by the family", "Whether more than one language is spoken in the corpus (regardless of how many languages are spoken in the family or the community at large)"),
  `N of corpus (%)` = c(intermediate_SES, intermediate_ED , intermediate_ocup, intermediate_urb, intermediate_HHS, intermediate_percsib, intermediate_avgsib, intermediate_langspo, intermediate_multiling)
)

# Create the table using kableExtra # currently not working
# mytab1 %>%
#   kbl(caption = "Definition of the corpus-level variables", align = c("lccccccccr")) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
#   column_spec(1, width = "2cm") %>%
#   column_spec(2, width = "3cm") %>%
#   column_spec(3, width = "6cm") %>%
#   column_spec(4, width = "3cm") %>%
#   scroll_box(width = "100%", height = "500px")

apa_table(mytab1,caption="Variables of interest", longtable = F)
#write.csv(mytab1,"output/table1.csv")

```


### Socio-economic Status: SES, Education, and Occupation

The vast majority of families were of middle or higher SES. In about a fifth of the corpora samples were described as mixed (spanning lower to middle or higher SES). A very small percentage were described as families of low SES. As we show in SM2, this general distribution over-represents higher SES families.
A smaller proportion of corpora (additionally) reported parental education level (Figure 4). Among families in CHILDES, three-quarters included at least some children whose parents had a graduate or postgraduate degree. In fact, over half of the corpora contained data solely from children whose parents had at least a graduate degree. Overall, more educated parents were overrepresented when compared to their source countries, which is reinforced by the observations that parents had professions linked to graduate-level education in over half of corpora (see SM3).


```{r figure4,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Distribution of Minimum Levels of Parental Education by Country. Spain & Hungary corresponds to a corpus collected in two countries. The order of countries is reversed-alphabetical. Note that the figure displays data only from countries for which we have available information. Corpora with missing data are not represented in the figure The same applies for Figures 5-8."}

annotations_inc$Education.ac = factor(annotations_inc$Education.ac, levels=c("College and above", "Some college", "Some secondary","Some primary"  ))

# Plot this categorywise bar chart
g <- ggplot(data=subset(annotations_inc, !is.na(Education.ac)), aes(country))
g + geom_bar(aes(fill=Education.ac), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  
  coord_flip() + 
  theme_minimal()  +
  theme(plot.title = element_text(size = 10, face = "bold"),
        strip.text.x = element_text(size = 10)) +
  scale_fill_manual(breaks=c("Some primary", "Some secondary", "Some college", "College and above"),
                    values = c("#E6FFFF", "#99E6FF", "#4CA6FF", "#0040FF" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of corpora with Education data", 
                     breaks = seq(0,12 , by=1)) +
  labs(title="Education of caregivers", fill = "Education Level")

```

### Urbanization

The majority of corpora were urban (Figure 5), with an over-representation of urban households given the source countries (see SM4).

```{r figure5,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Distribution of Type of Community by Country."}
annotations_inc$Type.of.community[annotations_inc$Type.of.community %in% c("academic","capital city of the Soviet Union","city","industial","industrial ","work-for-pay",
                                                                                                        "work-for-pay/industrial","industrial","Northeastern US urban","urban western", "industrial & service &trade activities...", "Mediterranean city of 1.6 million inhabitants", "modern city families, usually both working parents","Seaside tourist town","Industrial city (population: around 272,000 inhabitants)","mainly urban", "industrial city (3.000.000 inhabitants")]<-"urban"

annotations_inc$Type.of.community[annotations_inc$Type.of.community %in% c("farmer","rural","rural farming village")]<-"rural"
annotations_inc$Type.of.community[annotations_inc$Type.of.community %in% c("industrial, farmers")]<-"both"
annotations_inc$Type.of.community = factor(annotations_inc$Type.of.community, levels=c("urban", "both", "rural"))



# Plot this categorywise bar chart
gurb <- ggplot(data=subset(annotations_inc, !is.na(Type.of.community)), aes(country))
# g + geom_bar(aes(fill=Education.ac), width = 0.6, position = "fill") + 
#   theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() +
#   labs(title="Education of caregivers") 
gurb + geom_bar(aes(fill=Type.of.community), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() + 
  #facet_grid(~ scenario_f) +
  theme_minimal()  +
  #ggtitle("Education of caregivers")  + 
  theme(plot.title = element_text(size = 10, face = "bold"),strip.text.x = element_text(size = 12)) +
  coord_flip() + 
  #guides(fill=FALSE ) + 
  scale_fill_manual(breaks=c("rural", "both", "urban"),values = c("#BFAB57", "#966F56", "#808A8F" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of corpora with Urbanization data", breaks = seq(0, 16, by=1)) +
  labs(title="Urbanization", fill = "Type of community")



```


### Family Structure: Household composition and number of siblings

As Figure 6 shows, the overwhelming majority of corpora contained only nuclear families (i.e., parents and their children), and only a tiny percentage had a mixture of nuclear and extended families. We were not able to find a systematic database allowing us to check whether this was representative of the source countries.

Next, about a quarter of the corpora in our CHILDES sample consist exclusively of singletons (children with no siblings), while the remaining include children who have at least one sibling (Figure 7; see SM5 for why singletons might not be underrepresented).

```{r figure6,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Distribution of Household Arrangement by Country. Sweden & Portugal corresponds to a corpus collected in two countries. The same applies for Figure 8."}
annotations_inc$Household.structure = factor(annotations_inc$Household.structure, levels=c("nuclear", "extended", "varied"))

# Plot this categorywise bar chart
ghouse <- ggplot(data=subset(annotations_inc, !is.na(Household.structure)), aes(country))
# g + geom_bar(aes(fill=Education.ac), width = 0.6, position = "fill") + 
#   theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() +
#   labs(title="Education of caregivers") 
ghouse + geom_bar(aes(fill=Household.structure), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() + 
  #facet_grid(~ scenario_f) +
  theme_minimal()  +
  #ggtitle("Education of caregivers")  + 
  theme(plot.title = element_text(size = 10, face = "bold"),strip.text.x = element_text(size = 10)) +
  coord_flip() + 
  #guides(fill=FALSE ) + 
  scale_fill_manual(breaks=c("nuclear", "varied", "extended"),values = c("#ADDFAD", "#D1BA97", "#F08080" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of corpora with Household data", breaks = seq(0, 13, by=1)) +
  labs(title="Households", fill = "Type of household arrangement")

```

```{r figure7,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Distribution of Singletons by Country."}

# Plot this categorywise bar chart
gsib <- ggplot(data=subset(annotations_inc, !is.na(siblings)), aes(country))
# g + geom_bar(aes(fill=Education.ac), width = 0.6, position = "fill") + 
#   theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() +
#   labs(title="Education of caregivers") 
gsib + geom_bar(aes(fill=siblings), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() + 
  #facet_grid(~ scenario_f) +
  theme_minimal()  +
  #ggtitle("Education of caregivers")  + 
  theme(plot.title = element_text(size = 10, face = "bold"),strip.text.x = element_text(size = 10)) +
  coord_flip() + 
  #guides(fill=FALSE ) + 
  scale_fill_manual(breaks=c("yes", "no"),values = c("#B5A6DF", "#000080" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of corpora with Singletons data", breaks = seq(0, 11, by=1)) +
  labs(title="Singletons", fill = "Only singletons")

```

### Languages

```{r figure8,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Distribution of Indo-European Languages by Country."}
lang_fam<- read.csv("../derived/annotations_included_languagefamily.csv") 
#xtabs(~indo_european, lang_fam) #indoeuropean 122 / no 58
#xtabs(~language_family, lang_fam)

English_corpora_all<- sum(freq_languages$Freq[freq_languages$Var1 == "English" | freq_languages$Var1 == "Portuguese/Swedish/English" |freq_languages$Var1 == "Hungarian/Farsi/English" |freq_languages$Var1 == "English/Spanish" |freq_languages$Var1 == "English/Russian" |freq_languages$Var1 == "English/Mandarin/Cantonese" |freq_languages$Var1 == "English/Mandarin" |freq_languages$Var1 == "Portuguese/Swedish/English" |freq_languages$Var1 == "Portuguese/Swedish/English" |freq_languages$Var1 == "English/Japanese/Danish" |freq_languages$Var1 == "English/Japanese" |freq_languages$Var1 == "English/Hebrew" |freq_languages$Var1 == "English/French" |freq_languages$Var1 == "English/Dutch" |freq_languages$Var1 == "English/Cantonese" |freq_languages$Var1 == "Dutch/English, Dutch/French"])

English_ind_all<- sum(lang_dif$Total_Participants[lang_dif$Language == "English" | lang_dif$Language == "Portuguese/Swedish/English" |lang_dif$Language == "Hungarian/Farsi/English" |lang_dif$Language == "English/Spanish" |lang_dif$Language == "English/Russian" |lang_dif$Language == "English/Mandarin/Cantonese" |lang_dif$Language == "English/Mandarin" |lang_dif$Language == "Portuguese/Swedish/English" |lang_dif$Language == "Portuguese/Swedish/English" |lang_dif$Language == "English/Japanese/Danish" |lang_dif$Language == "English/Japanese" |lang_dif$Language == "English/Hebrew" |lang_dif$Language == "English/French" |lang_dif$Language == "English/Dutch" |lang_dif$Language == "English/Cantonese" |lang_dif$Language == "Dutch/English, Dutch/French"])

Spanish_corpora_all<- sum(freq_languages$Freq[freq_languages$Var1 == "Spanish" | freq_languages$Var1 == "Spanish/Catalan" |freq_languages$Var1 == "Spanish/Galician" |freq_languages$Var1 == "German/Spanish" |freq_languages$Var1 == "English/Spanish" |freq_languages$Var1 == "Hungarian/Catalan/Spanish"])

Spanish_ind_all<- sum(lang_dif$Total_Participants[lang_dif$Language == "Spanish" | lang_dif$Language == "Spanish/Catalan" |lang_dif$Language == "English/Spanish" |lang_dif$Language == "Hungarian/Catalan/Spanish" |  lang_dif$Language == "Spanish/Galician"| lang_dif$Language == "German/Spanish" ])


# Plot this categorywise bar chart
glang_fam <- ggplot(data=lang_fam, aes(country))
# g + geom_bar(aes(fill=Education.ac), width = 0.6, position = "fill") + 
#   theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() +
#   labs(title="Education of caregivers") 
glang_fam + geom_bar(aes(fill=language_family), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() + 
  #facet_grid(~ scenario_f) +
  theme_minimal()  +
  #ggtitle("Education of caregivers")  + 
  theme(plot.title = element_text(size = 10, face = "bold"),strip.text.x = element_text(size = 10)) +
  coord_flip() + 
  #guides(fill=FALSE ) + 
  scale_fill_manual(breaks=c("indo_european", "other"),labels=c("Indo-European", "Other"), values = c("#CC5700", "#008080" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of corpora", breaks = seq(0, 31, by=1)) + 
  labs(title="Language Family", fill = "Language Family")

```

We found a total of `r length(levels(as.factor(annotations_inc$Language)))` different languages or language combinations (for bilingual and multilingual children; see SM6). About a third of the included corpora were not monolingual. According to information in SM7, our sample of CHILDES underrepresents bilingual and multilingual households in the world.
As for linguistic diversity, most of the corpora pertain to Indo-European languages (Figure 8). English is the most prevalent language, with `r English_corpora_all` corpora (`r sum(annotations_inc$Language=="English")` exclusively monolingual), comprising `r round(English_corpora_all/nrow(annotations_inc)*100) `% of the included corpora. Additionally, it is the language most studied for bilingualism or multilingualism, with English being combined with: Cantonese, Danish, Dutch, Farsi, French, Hebrew, Hungarian, Japanese, Mandarin, Portuguese, Russian, Spanish, and Swedish. The next most represented language in terms of different corpora is Spanish, with `r Spanish_corpora_all` corpora (`r sum(annotations_inc$Language=="Spanish")` monolingual).


# Discussion

The current paper provides a systematic review of corpora of naturalistic recordings in the CHILDES database, a priceless public research resource. We assessed and characterized sampling bias by examining the database through four dimensions believed to be relevant for early language acquisition: SES, urbanization, family structure, and languages. 

Thanks to the collective work of nearly 300 individuals (with deep thanks to the people acknowledged in SM8), we included 180 corpora, in 63 different languages or language combinations, from 48 different countries, primarily OECD (84%). Despite a remarkable diversity in terms of languages and countries, analyses of the demographic characteristics of our CHILDES sample revealed several measures where the corpora do not represent the world population, nor faithfully represent the countries in which data were collected, in terms of SES, urbanization, family structure, and linguistic dimensions.

In terms of SES, our macro-level analyses showed that represented countries had a higher percentage of individuals completing lower secondary school, and higher Gross Domestic Product (GDP) per capita, compared to the worldwide distribution. Our analyses further suggested that higher SES families with highly educated parents were more common than would have been expected given their source countries. Considering the extensive literature linking SES and language acquisition, these are important biases to take into account.

Notably, approximately half of the corpora feature parents engaged in research-related professions, with academics frequently studying their own (grand-)children. In comparison, in the US general population, in 2020<!-- in 2020, 47,000 people were scientists in the USA (including both in academia and the private sector), 135,000 were professors, 20 million were Masters students, and 3 million were PhD students. Given that the USA’s population is 392 million, -->, only 6% of American citizens would fall under such definitions. The danger of extrapolating from such a sample goes beyond issues previously discussed in the context of convenience (homogeneous or heterogeneous) samples [@bornstein2013sampling], with parents pursuing a career in academia (and likely in language-related areas) possibly being unusual in their linguistic behavior.


Regarding urbanization, `r round(sum(annotations_inc$Type.of.community=="urban" & !is.na(annotations_inc$Type.of.community))/sum(!is.na(annotations_inc$Type.of.community))*100)`% of corpora were urban, compared to `r round(mean(ind_all$SP.URB.TOTL.ZS, na.rm=T))`% of the world population, with only `r round(sum(annotations_inc$Type.of.community=="rural" & !is.na(annotations_inc$Type.of.community)))` corpora exclusively rural. The latter seemed to be reasonable given urbanization statistics for the countries where data were collected. Thus, in this case, the bias pertains only to which countries are represented, and not how data are sampled within the country. One challenge for future work will involve teasing apart the ways in which urbanization may relate to language outcomes, including not only the parental goals mentioned in the Introduction [@keller2012autonomy] but also factors that may be confounded with urbanization. For example, a study using long-form recordings suggested that infants growing up in formerly rural families that had migrated to urban centers were afforded less spoken input, and had smaller vocabularies, than infants whose families had remained rural [@ma2024associations].

In terms of family structure, the represented countries had smaller households compared to the world average. When comparing household composition in our CHILDES sample with the countries where data had been collected, we found a possible trend for underrepresenting extended families. A bias towards nuclear families limits our understanding of diverse caregiving structures, particularly alloparental involvement, in language acquisition. 

While multi-child families are well-represented in our CHILDES sample, the possible absence of other household members' speech in some transcripts [e.g., siblings' speech, @loukatou2021child] could hinder insights into typical daily language interactions. Limited data on birth order also restricts robust conclusions on sibling configuration in CHILDES and underscores the need for broader exploration of non-parental caregiving roles typically overlooked in studies focusing on predominantly WEIRD households.

Finally, we found that English and more broadly Indo-European languages were overrepresented in our CHILDES sample, a point already made by others [e.g., @christiansen2022toward]. This aligns with trends in child development literature [@kidd2022diverse; @singh2023diversity]. Interestingly, out of the 10 countries with the most utterances (Figure 2), 4 were countries or territories where non-Indo-European languages are the main language spoken: Japan, Indonesia, China and Taiwan. Additionally, although a third of the corpora were not monolingual, this nonetheless underrepresents multilingualism in the world's population [@grosjean2024statistics].

These biases likely have both proximal and distal consequences for child language. Proximally, studies on typologically diverse languages [e.g., @casillas2024little; @de1999verbs; @demuth2022acquisition; @pye1980acquisition] highlight the limitations of current acquisition theories to the extent they rely on biased samples. For example, theories like syntactic bootstrapping, which assume consistent exposure to specific sentence structures, were developed in the context of Indo-European language. These can struggle to generalize to free word order languages such as Warlpiri, and tend to stress the syntactic rather than morphological side of bootstrapping, as English is relatively morphologically impoverished language [@babineau2023learning].

Broader socioecological diversity, such as variations in SES that influence familial stress [@troller2024relations], or polyadic caregiving environments that shape input sources and conversational dynamics [@sagna2022acquisition], may open new ways of thinking about the forces driving some of our classical findings on SES and input. Multilingual settings add another layer of complexity; for instance, Thai-English bilinguals demonstrate distinct narrative styles and gestural patterns depending on the language used, reflecting cultural norms and linguistic features [@rochanavibhata2022diversity].
More distal factors, such as noise pollution—more common in dense urban areas— linked with early brain development [@simon2022environmental] can have downstream effects on language acquisition. Skewed samples not only hinder the testing and generalization of language acquisition theories, but also risk mischaracterizing developmental mechanisms more broadly. 

Although our CHILDES sample is more homogeneous than expected given global and national variations, there remains substantial diversity across all the dimensions we studied. However, these dimensions are also confounded in these CHILDES corpora. Put simply, since most samples are from nuclear families with well-educated parents living in urban sites, it is impossible to tease apart these three dimensions. Additionally, this confound makes it hard to understand the intersectionality of such factors. 

While making some strides in documenting the demographic characteristics of the people represented in naturalistic corpora in CHILDES, our work has several limitations. First, we focused on SES, urbanization, family structure, and language, but other dimensions may well be relevant for language acquisition research. We hope that our sharing of materials will facilitate others' addition of novel descriptors. Second, the absence of comprehensive data for certain variables restricted a thorough analysis of representativeness on some dimensions. This should not be seen as the fault of the corpus curators, as these data were not always available or were not considered relevant at the time. It is only recently that researchers started sounding the alarm regarding underreporting of relevant demographic characteristics [e.g., @singh2023unified; @singh2023diversity].

Our paper emphasizes the importance of evaluating the diversity of the samples we rely on. This issue is not unique to CHILDES, and applies broadly to any developmental archive [e.g., Databrary, @databrary]. In the era of big data, AI, and the rise of child-focused technologies, it is essential to consider the generalizations that models based on such data might produce for language acquisition - and reflect on how these models and generalizations impact the entire developmental field. 

Researchers using CHILDES data should familiarize themselves with the demographic characteristics of their samples and acknowledge any limitations that could affect their findings. For future contributions, we recommend including demographic metadata following CHILDES’s already existing metadata system. Requiring data entry would ensure the inclusion of key descriptors at both the speaker and corpus levels, improving the database’s demographic information. When collecting data, purposive sampling can help include all key sub-groups for a representative sample [see @bornstein2013sampling]. Collaborating with local communities and adopting participatory practices [@gehlert2018seeing] can enhance data collection in diverse contexts.

Finally, systemic changes in academia are crucial for fostering long-term diversity. Journals, editorial boards, and funding agencies should create opportunities and provide support for studies on underrepresented populations and researchers. Adopting key performance indicators for diversity—focusing on linguistic, geographic, and socioeconomic representation [@havron2022community]—and increasing the inclusion of non-Western authors in developmental science [@aravena2024towards] are actionable steps to address systemic biases.

To conclude, we found that many countries and languages are represented in CHILDES, but that these countries did not constitute a representative sample of the world's countries, and that for many dimensions, families were not representative of their own countries. We also noted low variability along certain dimensions (e.g., SES and family structure) and the predominance of Indo-European languages, with English having the most recorded corpora. Despite these limitations, CHILDES remains an unparalleled resource that has contributed significantly to the field over decades, exemplifying the success of open and collaborative science. Our discussion leads us to argue for the systematic inclusion of descriptors in speaker-level and corpus-level CHILDES metadata, to track and quantify both diversity and representativity.

```{r writedown, include=FALSE}

lang_fam$indo_european <- NULL
lang_fam$X <- NULL

lang_fam <- lang_fam[, !(names(lang_fam) %in% c("indo_european", "X"))]

write.csv(lang_fam,"../derived/clean/Database_data.csv",row.names=F)

```


\newpage
# References
