---
title: "How WEIRD-biased is CHILDES' data on children's linguistic input"
shorttitle        : "WEIRD CHILDES"

author: 
  - name          : "Camila Scaff*"
    affiliation   : "1,2"
  - name          : "Georgia Loukatou*"
    affiliation   : "2"
  - name          : "Alejandrina Cristia"
    affiliation   : "2"
  - name          : "Naomi Havron"
    affiliation   : "3"
    
affiliation:
  - id: "1"
    institution: University of Zurich, Institute of Evolutionary Medicine (IEM), Switzerland
  - id: "2"
    institution: PSL University, Laboratoire de Sciences Cognitives et de Psycholinguistique (ENS, EHESS, CNRS, DEC), France
  - id: "3"
    institution: Haifa Univerity, Israel



  # - code: "2"
  #   address: ENS, EHESS, CNRS, PSL University, Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d études cognitives, 29 rue d ulm, Paris, 75005, France

# corresp_author_name:  "ANON"
# corresp_author_email: "ANON"

abstract: |
  In recent years, the importance of estimating demographic biases in research has become apparent. Here we provide a systematic review of the CHILDES archive, the major source of data on naturalistic recordings of children’s linguistic environment. We analyzed the archive at the country and corpus level for four dimensions considered central for language learning: SES, urbanization, family structure and language. We compared these descriptive statistics to world statistics to assess whether the archive was biased in terms of the demographics of the countries represented and the families recorded within them. We found that at the country level, the 47 countries from which there were recordings in CHILDES overrepresented countries with higher educational level; were more urban; and had smaller households with less children. At the corpus level,  middle- and higher-class participants were over-represented in relation to the statistics of their own countries. Corpora also included more educated families, with academics being especially over-represented. The corpora were not representative of their countries in terms of urbanization either - with a larger percentage of families residing in urban settings than is overall true for the respective countries. In terms of family structure, nuclear families were more prevalent than in the countries the data was collected in, and - surprisingly - children with no siblings appeared to be under-represented. Last, we found that corpora were linguistically diverse, but we estimate that data in CHILDES under-represents bilingual and multilingual households. We conclude that when generalizing from analysis of data obtained from CHILDES, researchers should acknowledge the potential biases of the archive. 



## Remove this if not required
funding: |
  AC acknowledges financial and institutional support from Agence Nationale de la Recherche (ANR-16-DATA-0004 ACLEW, ANR-17-EURE-0017, ANR-14-CE30-0003 MechELex); the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award; and the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (ExELang, Grant agreement No. 101001095). GL acknowledges financial support from Fondation Fyssen.




## Remove this if not required
acknowledgements: |
  We are grateful to Lara Oliel and Clara Daoud for help entering corpus-level data; Sara Pisani for organizing communication with corpus curators; Brian MacWhinney for support and encouragement when retrieving information from corpus curators. We particularly want to thank the literally hundreds of researchers who contributed data to CHILDES, without which this project would be unthinkable; and the 103 curators who replied to our request.

# ac2cs shall we add their names? Do we have permission for that?


keywords          : "childes, verbal input, infant-directed speech, language, weird"
wordcount         : " XXwords"


bibliography      : ["childes.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---


```{r setup, include=FALSE}
Sys.setlocale("LC_CTYPE", "en_US_POSIX.UTF-8")
#Sys.setlocale("LANG", "en_US_POSIX.UTF-8")
#Sys.setlocale(category = "LC_ALL", locale = "en_SE.utf8")

knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE)
# Install remotes package if necessary
if(!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")

# Install the stable development version from GitHub
#remotes::install_github("crsh/papaja")
library("papaja")
library(psych)
library(car)
library(ggplot2)
library(ggthemes)
library(scales)
library(gridExtra)
library(lme4)
library(ggpubr)
library(lmerTest)

library(knitr)
options(digits=1)
library(kableExtra)
library(dplyr)
library(effects)
library(gvlma)
library(influence.ME)
library("wesanderson")
#devtools::install_github("psyteachr/introdataviz")
#library(ggalt)
library(Hmisc)

annotations_inc<- read.csv("../derived/annotations_included", sep="")
ind<- read.csv("../derived/ind", sep="")
ind_all<- read.csv("../derived/ind_all", sep="")
ocde <- read.csv("../data/ocde_country.csv", sep=",")


##countries and corpora
corpora <- as.data.frame((table(annotations_inc$country)))
names(corpora)<-c("country","Freq")
#write.table(corpora,"derived/corpora", row.names = FALSE, col.names = TRUE)


merge(x=corpora, y=ocde,  by = "country", all.x = T) -> ocde_country

annotations<- read.csv("../data/Table for authors - annotations.csv") 
#original link https://docs.google.com/spreadsheets/d/1s-ytfQf7WsZFDDZ6QkOQnZhTodHpQp7D2Wja8YFvjQY/edit?usp=sharing

#some cleaning  of the col names
colnames(annotations)=gsub("\\.\\..*","",colnames(annotations))
colnames(annotations)[colnames(annotations)=="Number.of.participants"]<-"Nb.of.participants"
colnames(annotations)[colnames(annotations)=="Language.or.Languages.spoken.in.recordings..be.specific.if.possible.e.g..French.Quebec."]<-"Language"
colnames(annotations)[colnames(annotations)=="Language.or.Languages.spoken.in.recordings..be.specific.if.possible.e.g..French.Quebec."]<-"Language"
colnames(annotations)[colnames(annotations)=="Location..Neighbourhood..village.city..province..state..country."]<-"Location"
colnames(annotations)[colnames(annotations)=="X..of.children.with.siblings"]<-"Nb.of.children.with.siblings"
colnames(annotations)[colnames(annotations)=="X..of.children.with.older.siblings"]<-"Nb.of.children.with.older.siblings"
colnames(annotations)[colnames(annotations)=="Our.coding.of"]<-"Our.coding.of.community.type"

#Inclusion
annotations$Inclusion[annotations$Inclusion %in% c("No","no")] <- "exclude"
annotations$Inclusion[annotations$Inclusion %in% c("Yes","yes")] <- "include"
#why_exclude
excluded <-  subset(annotations, (Inclusion %in% c("exclude")))
#xtabs(~why_exclude, excluded)




```

## Research highlights

*We examined CHILDES, the major source of data on children’s linguistic environment for potential sampling bias in terms of SES, urbanization, family structure and languages.
*We found that the 47 countries present in CHILDES overrepresented rich, educated, urbanized countries with small nuclear families.
*Within these countries, corpora overrepresented rich, educated, urban and nuclear families - and single-child families, as well as bilingual families were underrepresented. 
*Interpretation of studies based on CHILDES should acknowledge these biases.

\newpage


 Since its foundation in 1984, CHILDES [the Child Language Data Exchange System, @macwhinney2000childes] has been the major source of naturalistic recordings and transcript data for researchers studying language acquisition. Naturalistic recordings provide insight into how children acquire language in everyday contexts and capture the richness and complexity of language use in everyday conversations. They also constitute a valuable and ecologically valid source of data necessary for computational models seeking to reverse engineer language acquisition or simulate natural language development. This repository of naturalistic recordings has been the foundation for many influential concepts in developmental science and beyond, illustrated by more than 7,000 scientific publications [@macwhinney2019understanding]. It has contributed to establishing seminal work in various domains such as the theory of mind [@bartsch1995children] and human memory [@anderson1991reflections]. It has also inspired theories that help us understand the connection between language input and language development [@christiansen1998learning].
 
However, naturalistic recordings may also contain potential biases. One notable concern is that participants might not represent the diverse socio-economic backgrounds, cultures, or linguistic environments children typically experience. As a result, the generalizability of the related findings may be limited, necessitating cautious interpretation. Researchers in many fields are increasingly aware of the bias towards WEIRD populations [Western, Educated, Industrial, Rich, Democratic, @henrich2010most] in the samples they study [@nielsen2017persistent;@cychosz2022using ;@moriguchi2022beyond; @singh2023diversity]. Recent calls have been made to diversify research in psychology, cognitive, and developmental science to address this issue [@Majid2010; @kidd2022diverse; @blasi2022over]. For instance, @kidd2022diverse  systematically reviewed publications from child-language journals. They revealed biases towards specific continents (North America and Europe) and languages (English. Spanish, French), highlighting the need for increased diversity in populations and languages studied. Similarly, @blasi2022over emphasize the potential consequences of generalizing observations derived solely from English speakers and how fit they are to represent our entire human species. While these studies focused on biases towards specific languages, other sources of variance might be as important for research on language development.
Every area of research conceptualizes different dimensions relevant to explaining variance for a given research topic. In what follows we present the four most relevant dimensions thought to play a role in child language acquisition, particularly related to naturalistic recordings: socioeconomic status, urbanization, family structure, and language. 

### Socioeconomic Status
Decades of research have examined the linguistic differences among families with varying socio-economic status (SES). In the developmental literature, this is primarily indexed by parents’ education [@ensminger2014decade; @hoff2014causes],  but can also be indexed otherwise, such as by parental income, occupation, or a composite measure of these three (e.g. Hollingshead, 1975).
It is beyond the scope of this paper to detail all the theories that attempt to account for the complex causal pathways that may connect SES to children's language environments. We recommend @rowe2018understanding as a starting point for readers interested in this literature, along with @golinkoff2019language and @sperry2019reexamining for diverse theoretical perspectives. It should be noted, however, that some of this literature has been found to reflect Global North biases, including what kinds of language input are counted, and what features of linguistic experiences are valued [@sperry2019reexamining; @scaff2024characterization]. Without desiring to take a stance on how SES and language environments relate to each other, we merely indicate here that SES is undoubtedly one of the factors that has been repeatedly studied in the context of early language acquisition, particularly related to input. For example,  @hoff2014causes compared the speech of high- versus low-SES American mothers. College-educated parents produced more utterances to their child, with more diverse vocabulary, longer phrases, and higher number of utterances continuing a topic the child had brought up. Similar findings can be seen in other studies [@hart1995meaningful; @hoff1990maternal; @hoff2003specificity; @huttenlocher2007varieties; see also @dailey2022language; @leonardo2022socioeconomic; for meta-analyses supporting the link; and @bergelson2023everyday for a large-scale study finding non-significant SES effects]. 

### Urbanization
Urbanization is the process of moving from rural to urban areas along with noticeable changes in job opportunities and living conditions. It involves the growth and development of cities, leading to increased access to infrastructure and amenities. Within the general theoretical framework of language socialization, there have been proposals that societies varying in their urbanization process have differing views and values about the role of children in conversations, and more generally in the community [@sharma1998child ; @richman1992cultural; @draper2017parent; @keller2012autonomy]. For instance, @keller2012autonomy discusses three prototypical cases: urban, rural, and hybrid. These three groups differ in terms of their goals for children, with urban families aiming for child psychological independence, rural families for child physical autonomy and interdependence, and hybrid families aiming for some mix across these values. @vogt2020multimodal employ this conceptual classification to interpret their results on multimodal language use across three samples: urban Dutch, urban Mozambique, and rural Mozambique. They found that the number of gestures, gesture-speech alignment, and gesture types all vary across the three groups in ways that can be related to Keller's typology. Similarly, @cristia2023systematic systematic review concludes that children's urbanization status maps onto the amount of input afforded by caregivers: children from rural communities are exposed to less input from caregivers than children in urban ones.

### Family structure
Family structure refers to the arrangement within a household, forming the basis of a family unit. This dimension encompasses aspects such as the number of siblings, birth order, the number of caregivers in the household, and the number of individuals sharing or competing for household resources (including caregiving attention); each of which has a significant impact on child and language development [@blake1981family; @duncan2020home; @havron2022effect;  @havron2019effect; @hoff1991older; @tomasello1985pragmatics; @bornstein2019mother]. For example, birth order effects reveal that children with older siblings show lower language skills than first-born children in various cultures [e.g., @peyre2016differential in France; @havron2022effect for Singapore; @zambrana2012impact for Norway]. Other birth order effects suggest that second-born children might benefit in production through overheard speech from their caregivers and older siblings [@oshima1996birth].
Regarding household composition, in many middle-class Euro-American families, parents typically assume primary responsibility for children, often focusing on the mother as the primary caregiver (e.g. Bakermans-Kranenburg et al., 2004; Huttenlocher et al., 2010; Ispa et al., 2004; Pan et al., 2005). However, certain cultures, like Turkish families described by @isleyen2021marital, may adopt a different approach, with nuclear families living in separate apartments but sharing common spaces and caregiving responsibilities, resulting in extensive support networks. 

### Languages 
Characterizing the diversity of participant samples in terms of language [@kidd2022diverse; @blasi2022over] is an important factor in language acquisition, as variations in language exposure and language use among different groups allow to explore how purely linguistic factors shape and influence the development of language skills in children. For example, based on transcriptions of conversations, it was shown that K’iche' Mayan children frequently use and understand passive constructions from a very young, unlike their English-speaking peers, refuting the idea that passive constructions can only emerge later in development [@pye1980acquisition; @pye1988precocious].
Similarly, many Indo-European languages show a strong noun bias in early vocabularies (a bias for acquiring words for concrete referential objects rather than actions), it has been claimed to be a universal feature of early language acquisition. However, studies have shown that in some Mayan languages, including Tseltal and Tsotsil [@casillas2024little; @de1999verbs], there is little to no evidence for a noun bias and argue for a verb bias instead. This highlights significant cross-linguistic variation and underscores the importance of studying naturalistic children's recordings for describing different linguistic developmental trajectories [@casillas2024little].


Another major dimension and entire sub-field in developmental science is the study of bilingualism or multilingualism [for an introductory review see @hohle2020variability, also @mccabe2013multilingual]. There is evidence that monolingual and bilingual early language development differs in some aspects, particularly regarding phonological acquisition and word learning. For example, monolingual infants' ability to discriminate non-native sounds declines during the first year of age, whereas infants exposed to one or more languages maintain the discrimination window for a longer period. Also, in terms of input, bilingual linguistic exposure is divided between two or more native languages. It has been shown that the amount of exposure to each native language can affect bilingual infants' speech discrimination abilities [@garcia2011bilingual]. 


### The current study
Here, we provide a systematic analysis of the naturalistic speech corpora of the CHILDES database by quantifying the diversity of each dimension presented above (see Table 1). Though some of these dimensions overlap with each other (See Figures SM.X and SM.x in the supplementary materials for illustrations), we decided to illustrate each as best as possible independently.
Our systematic analysis follows three steps. First, we screened the CHILDES database, excluding the clinical and task-driven recordings. Second, we extracted information related to the four central factors in language acquisition described in the introduction. Finally, we follow @ghai2021s recommendation to improve the description of diversity in behavioral sciences by looking at different levels: from a macro-level, with broad country comparisons to a corpus-level, where we delved into individual corpora to gain a more detailed and nuanced understanding of the data.

# Methods

Analyses and visualizations were carried out using R (version 4.1.2, R Core Team, 2020) and ggplot2 [wickham2011ggplot2]. Data, scripts, and online Supplementary Materials are available on OSF [https://osf.io/q9w82/?view_only=a013f1b25b8c4556b8248f12870402c9].

### Inclusion criteria
Although valuable insights occur from corpora on clinical populations and elicitation tasks, we deliberately excluded them from our review. This is because they introduce additional sources of variation and biases. For example, the inclusion of clinical populations may highlight differences in the prevalence and access to diagnosis of language disorders in different regions or countries. Additionally, the structure of elicitation tasks may not reflect spontaneous language use in everyday conversations. We thus excluded the following corpora: a) clinical populations or non-typically developing children, b) structured tasks such as toy narratives, personal narratives, frog stories, picture or movie descriptions, structured storytelling, and other elicitation tasks; c) only child or adult speech without a conversational partner; and d) non-naturalistic setups, such as recordings conducted exclusively in a lab environment or in a diary format.

### Screening
Following a thorough examination of each corpus, which involved reviewing the corpus description available on the CHILDES website, checking for any accompanying references such as articles, book chapters, or dissertations, and conducting spot-checks on associated transcripts, we identified `r nrow(annotations_inc)` corpora that met our inclusion criteria mentioned above. Please refer to the flowchart in Supplementary Materials (SM2) for a detailed breakdown of the included corpora.


### Descriptive-statistics

Descriptive statistics
Firstly, we show descriptive statistics of the countries represented in the CHILDES database presented in FigureS 1 and 2 of the results. Our goal is to provide insights into the representativeness of our CHILDES sub-sample when compared to global statistics. Figure 2 draws data from official sources: the CWorld Bank, Our World in Data (WDI), and the United Nations (UN). For additional details on Figure 2 and the corresponding variables, please refer to Supplementary Materials (SM3).

Secondly, we present corpus-level statistics to assess the representativeness of our sub-sample of CHILDES in a more detailed manner across our four dimensions: SES, urbanization, family structure, and languages (See Table 1 for the complete list of variables and definitions).

Data was extracted from the provided sources mentioned in CHILDES such as articles, book chapters, dissertations, and transcripts to pre-fill the categories. Corpus curators were contacted to request missing or incomplete information, and an online table was provided to facilitate data entry. Over a third of the contacted curators (103 corpora, `r round((103*100)/324)`%) provided additional data or confirmed missing information. Unfortunately, curators for 50 corpora  (`r round((50*100)/324)`%) could not be contacted due to unresponsive email addresses (38), or the curator’s passing (12). 


```{r tab1,echo = FALSE, as.is=T, ig.pos = "T!",results="asis"}
library(kableExtra)


#SES
SES = sum(!is.na(annotations_inc$SES.STDZD))
SES_missing =round(sum(is.na(annotations_inc$SES.STDZD))/length(annotations_inc$SES.STDZD)*100)
SES_present = 100- SES_missing
#ed
ED_ses = sum(!is.na(annotations_inc$Education.STDZD))
ED_ses_missing =round(sum(is.na(annotations_inc$Education.STDZD))/length(annotations_inc$Education.STDZD)*100)
ED_ses_present = 100- ED_ses_missing

#occupation
Ocup_ses = sum(!is.na(annotations_inc$Parental.profession))
Ocup_ses_missing =round(sum(is.na(annotations_inc$Parental.profession))/length(annotations_inc$Parental.profession)*100)
Ocup_ses_present = 100- Ocup_ses_missing

#urb Type of community
urbrur = sum(!is.na(annotations_inc$Type.of.community))
urbrur_missing =round(sum(is.na(annotations_inc$Type.of.community))/length(annotations_inc$Type.of.community)*100)
urbrur_present = 100- urbrur_missing


#family structure 
#Household composition
Household.structure = sum(!is.na(annotations_inc$Household.structure))
Household.structure_missing =round(sum(is.na(annotations_inc$Household.structure))/length(annotations_inc$Household.structure)*100)
Household.structure_present = 100- Household.structure_missing

#Percent children with sibling(s)", 
per_sibs = sum(!is.na(annotations_inc$Proportion.With.Siblings))
per_sibs_missing =round(sum(is.na(annotations_inc$Proportion.With.Siblings))/length(annotations_inc$Proportion.With.Siblings)*100)
per_sibs_present = 100- per_sibs_missing

#"Average number of siblings", 
avg_sibs = sum(!is.na(annotations_inc$Average.number.of.siblings))
avg_sibs_missing =round(sum(is.na(annotations_inc$Average.number.of.siblings))/length(annotations_inc$Average.number.of.siblings)*100)
avg_sibs_present = 100- avg_sibs_missing
#"Language(s) spoken", 
Langspoken = sum(!is.na(annotations_inc$Language))
Langspoken_missing =round(sum(is.na(annotations_inc$Language))/length(annotations_inc$Language)*100)
Langspoken_present = 100- Langspoken_missing
#"Lingual status
multiling = sum(!is.na(annotations_inc$Bilingualism.Multilingualism.in.corpus))
multiling_missing =round(sum(is.na(annotations_inc$Bilingualism.Multilingualism.in.corpus))/length(annotations_inc$Bilingualism.Multilingualism.in.corpus)*100)
multiling_present = 100- multiling_missing
#Language

intermediate_SES <- paste(SES, " (", round(SES_present), "%)", sep = "")
intermediate_ED <- paste(ED_ses, " (", round(ED_ses_present), "%)", sep = "")
intermediate_ocup <- paste(Ocup_ses, " (", round(Ocup_ses_present), "%)", sep = "")
intermediate_urb <- paste(urbrur, " (", round(urbrur_present), "%)", sep = "")
intermediate_HHS <- paste(Household.structure, " (", round(Household.structure_present), "%)", sep = "")
intermediate_percsib <- paste(per_sibs, " (", round(per_sibs_present), "%)", sep = "")
intermediate_avgsib <- paste(avg_sibs, " (", round(avg_sibs_present), "%)", sep = "")
intermediate_langspo <- paste(Langspoken, " (", round(Langspoken_present), "%)", sep = "")
intermediate_multiling <- paste(multiling, " (", round(multiling_present), "%)", sep = "")


# Create a data frame
mytab1 <- data.frame(
  Dimension = c("SES", "SES", "SES", "Urbanization", "Family structure", "Family structure", "Family structure", "Language", "Language"),
  `Corpus-level variable` = c("Parents' socio-economic status", "Parents' education level", "Parents' occupation", "Type of community", "Household composition", "Percent children with sibling(s)", "Average number of siblings", "Language(s) spoken", "Lingual status"),
  `Definition or levels` = c("Low, mid and/or high", "Highest level of education completed: Primary, High school, Undergraduate, Masters, PhD", "Parents' activity or profession", "Urban, rural, mixed", "Whether the family was composed primarily of caregivers and children (nuclear) or it also included other family members: grandparents, cousins, uncles, and aunts (extended)", "The percentage of children in the corpus who had at least one sibling", "How many siblings children had on average (including children with zero siblings)", "Which languages were spoken in the transcripts and by the family", "Whether more than one language is spoken in the corpus (regardless of how many languages are spoken in the family or the community at large)"),
  `N of corpus (%)` = c(intermediate_SES, intermediate_ED , intermediate_ocup, intermediate_urb, intermediate_HHS, intermediate_percsib, intermediate_avgsib, intermediate_langspo, intermediate_multiling)
)

# Create the table using kableExtra # currently not working
# mytab1 %>%
#   kbl(caption = "Definition of the corpus-level variables", align = c("lccccccccr")) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
#   column_spec(1, width = "2cm") %>%
#   column_spec(2, width = "3cm") %>%
#   column_spec(3, width = "6cm") %>%
#   column_spec(4, width = "3cm") %>%
#   scroll_box(width = "100%", height = "500px")

apa_table(mytab1,caption="Definition of the corpus-level variables", longtable = F)
#write.csv(mytab1,"output/table1.csv")

```


# Results

```{r demoinfo, echo = FALSE, eval=T}

corpora.ind <- annotations_inc %>%
  select(country, Corpus, Nb.of.participants, region)

corpora.ind$region[corpora.ind$country =="Spain & Hungary"] = "Europe"
corpora.ind$region[corpora.ind$country =="Sweden & Portugal"] = "Europe"

summary_corpora_COUNTRY <- corpora.ind %>%
  group_by(country) %>%
  summarise(
    TimesRepeated = n(),
    TotalParticipants_country = sum(Nb.of.participants)
  )

summary_corpora_CONT <- corpora.ind %>%
  group_by(region) %>%
  summarise(
    TimesRepeated = n(),
    TotalParticipants_continent = sum(Nb.of.participants)
  )


```

The `r nrow(annotations_inc)` corpora included in this study represent `r length(levels(as.factor(annotations_inc$country)))-2` <!-- minus 2 bilingual corpora counting twice Hungary/ Spain and Sweden/Portugal -->different countries or territories across all populated continents. The country with the most corpora is the United States with `r length(annotations_inc$country[annotations_inc$country == "United States"])` different corpora, followed by Spain with `r length(annotations_inc$country[annotations_inc$country == "Spain"]) + 1` (including a bilingual corpus), and a tie between the United Kingdom and France with `r length(annotations_inc$country[annotations_inc$country == "France"])` (see SM4 for a full summary by country and continent). The most represented country in terms of individual children recorded is the United Kingdom with `r summary_corpora_COUNTRY$TotalParticipants_country[summary_corpora_COUNTRY$country == "United Kingdom"]`, followed by the United States with  `r summary_corpora_COUNTRY$TotalParticipants_country[summary_corpora_COUNTRY$country == "United States"]` and then Spain with  `r summary_corpora_COUNTRY$TotalParticipants_country[summary_corpora_COUNTRY$country == "Spain"] + summary_corpora_COUNTRY$TotalParticipants_country[summary_corpora_COUNTRY$country == "Spain & Hungary"] `. The most represented continent in terms of individual children recorded is Europe with`r summary_corpora_CONT$TotalParticipants_continent[summary_corpora_CONT$region == "Europe"]`. The least represented continent is Oceania with only `r summary_corpora_CONT$TotalParticipants_continent[summary_corpora_CONT$region == "Oceania"]` recorded children [all from Papua New Guinea, @sarvasy2017grammar].

Out of the `r length(levels(as.factor(annotations_inc$country)))-2` countries or territories,`r sum(ocde_country$ocde == "yes", na.rm = TRUE)` belong to the OECD (Organisation for Economic Co-operation and Development). This corresponds to `r sum(ocde_country$Freq[ocde_country$ocde == "yes"],na.rm=T)` out of `r nrow(annotations_inc)` corpora or `r round((sum(ocde_country$Freq[ocde_country$ocde == "yes"],na.rm=T)*100)/nrow(annotations_inc))`%  of this sample of CHILDES. Meanwhile, OECD countries represent only 19.5% of world countries. 



```{r worldmapecho = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Geographical distribution of the different corpora of CHILDES. Each circle represents a corpus, and the size of the circle is proportional to the number of participants."}

## Figure
# Creating a subset with specific columns
map_info <- annotations_inc[, c("country", "Corpus", "Nb.of.participants", "Location")]
#write.table(map_info, file = "derived/map_info.txt", sep = "\t", quote = FALSE)

country_coords<- read.csv("../data/map_coordinates.csv", sep=",")

world_map <- map_data("world")

ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "lightgray") +
  geom_point(data = country_coords, aes(x = Longitude, y = Latitude, size = Nb.of.participants), color = "blue", alpha = 0.6) +
  theme_void() +
  labs(title = "Participants by Country")

ggsave("../figures/childes_country_participants.png", width = 6, height = 4, dpi = 300)

```

We assessed the extent to which the countries with data in our sub-sample of CHILDES were a representative sample of countries in the world. Density plots are portrayed in Figure 2. 

```{r densityplot, echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Density plots show the distribution of country-level descriptors for all world countries (gray) versus countries represented in our sub-sample of CHILDES (orange). % complete. LSS is the percentage of the country's population completing lower secondary school. % urban is the percent of the country's population residing in urban (as opposed to rural) locations. GDP stands for Gross Domestic Product. Avg Household size stands for average household size."}
library(ggplot2)


#GDP
dens_dat=data.frame(rbind(cbind(ind_all$log_gdp[!is.na(ind_all$log_gdp)],0),
                          cbind(ind$log_gdp[!is.na(ind$log_gdp)],1) ))
colnames(dens_dat)<-c("GDP.per.capita","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
gdpt = t.test(dens_dat[,1]~dens_dat$in.childes)

rich <- ggplot(dens_dat, aes(x = GDP.per.capita, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "GDP per capita (log 10)")+ scale_fill_colorblind()


#Education
dens_dat=data.frame(rbind(cbind(ind_all$Compl.LS[!is.na(ind_all$Compl.LS)],0),
                          cbind(ind$Compl.LS[!is.na(ind$Compl.LS)],1) ))
colnames(dens_dat)<-c("completed.LSS","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
edt = t.test(dens_dat[,1]~dens_dat$in.childes)

ed <- ggplot(dens_dat, aes(x = completed.LSS, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = c(0.25, 0.8), axis.title.y=element_blank() , legend.title = element_blank())+labs( x = "% compl. LSS")+ scale_fill_colorblind()

#urbanization
dens_dat=data.frame(rbind(cbind(ind_all$SP.URB.TOTL.ZS[!is.na(ind_all$SP.URB.TOTL.ZS)],0),
                          cbind(ind$SP.URB.TOTL.ZS[!is.na(ind$SP.URB.TOTL.ZS)],1) ))
colnames(dens_dat)<-c("pc.urban","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
urbt = t.test(dens_dat[,1]~dens_dat$in.childes)

urb <- ggplot(dens_dat, aes(x = pc.urban, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() )+labs( x = "% urban")+ scale_fill_colorblind()

#Family structure
#% nuclear family
ind_all$nuclear = (ind_all$Couple.with.children + ind_all$Single.parent.with.children)
ind$nuclear = (ind$Couple.with.children + ind$Single.parent.with.children)

dens_dat=data.frame(rbind(cbind(ind_all$nuclear[!is.na(ind_all$nuclear)],0),
                          cbind(ind$nuclear[!is.na(ind$nuclear)],1) ))
colnames(dens_dat)<-c("nuclear","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
nuclear = t.test(dens_dat[,1]~dens_dat$in.childes)

NUCLEAR <- ggplot(dens_dat, aes(x = nuclear, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "% Nuclear household with children")+ scale_fill_colorblind()

#% non nuclear family
ind_all$nnuclear = (ind_all$Extended.family + ind_all$Non.relatives)
ind$nnuclear = (ind$Extended.family + ind$Non.relatives)

dens_dat=data.frame(rbind(cbind(ind_all$nnuclear[!is.na(ind_all$nnuclear)],0),
                          cbind(ind$nnuclear[!is.na(ind$nnuclear)],1) ))
colnames(dens_dat)<-c("nnuclear","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
nnuclear = t.test(dens_dat[,1]~dens_dat$in.childes)

NNUCLEAR <- ggplot(dens_dat, aes(x = nnuclear, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "% Extended or Non-relatives in the household")+ scale_fill_colorblind()


#Average household size
dens_dat=data.frame(rbind(cbind(ind_all$Avg.hs[!is.na(ind_all$Avg.hs)],0),
                          cbind(ind$Avg.hs[!is.na(ind$Avg.hs)],1) ))
colnames(dens_dat)<-c("Avg.hs","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
hh = t.test(dens_dat[,1]~dens_dat$in.childes)

HH <- ggplot(dens_dat, aes(x = Avg.hs, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "Avg Household size")+ scale_fill_colorblind()

#Average household size
dens_dat=data.frame(rbind(cbind(ind_all$Avg.und15[!is.na(ind_all$Avg.und15)],0),
                          cbind(ind$Avg.und15[!is.na(ind$Avg.und15)],1) ))
colnames(dens_dat)<-c("Avg.und15","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
und15 = t.test(dens_dat[,1]~dens_dat$in.childes)

UND15 <- ggplot(dens_dat, aes(x = Avg.und15, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "Avg # members >15 years per household ")+ scale_fill_colorblind()


# combined
density_plot <- ggarrange(ed, urb, rich, HH,  
               labels = c("A", "B", "C", "D"),
               ncol = 2, nrow = 2)

ggsave("../figures/density_plot.png", width = 6, height = 4, dpi = 300)

```

By comparing our sub-sample of CHILDES to the world statistics using unpaired samples t-tests without assuming equality of variance (Welch's t). Countries in our sub-sample of CHILDES had a higher proportion of the population completing lower secondary school than the world wide sample (% compl. LSS, t(`r round(edt$parameter,2)`)=`r round(edt$statistic,2)`, p = `r round(edt$p.value,2)`); they were more urban (% urban, t(`r round(urbt$parameter,2)`)=`r round(urbt$statistic,2)`, p = `r round(urbt$p.value,2)`); richer (log GDP per capita, t(`r round(gdpt$parameter,2)`)=`r round(gdpt$statistic,2)`, p = `r round(gdpt$p.value,2)`) and had smaller households (average household size, t(`r round(hh$parameter,2)`)=`r round(hh$statistic,2)`, p = `r round(hh$p.value,2)`). 
<!--FIX P value reporting -->

We then investigated at the corpus level the variables described in Table 1.  The information collected gives us more detail about the characteristics of the individual families that compose our sub-sample of CHILDES. However, it is important to acknowledge that this information was missing for a substantial number of the corpora. 

### SES

Our analysis of `r length(annotations_inc$SES.STDZD) - sum(is.na(annotations_inc$SES.STDZD))` corpora including SES information reveals that only `r round((sum(annotations_inc$SES.STDZD==1 & !is.na(annotations_inc$SES.STDZD))/ (length(annotations_inc$SES.STDZD) - sum(is.na(annotations_inc$SES.STDZD))))*100)`% (n = `r sum(annotations_inc$SES.STDZD==1 & !is.na(annotations_inc$SES.STDZD))` corpora) were described as families of low SES,  `r round((sum(annotations_inc$SES.STDZD=="1-2" & !is.na(annotations_inc$SES.STDZD))/ (length(annotations_inc$SES.STDZD) - sum(is.na(annotations_inc$SES.STDZD))))*100)` (n = `r sum(annotations_inc$SES.STDZD=="1-2" & !is.na(annotations_inc$SES.STDZD))`) described as a mixed sample spanning between lower and middle or higher SES. The vast majority, `r round((sum(annotations_inc$SES.STDZD==2 & !is.na(annotations_inc$SES.STDZD))/ (length(annotations_inc$SES.STDZD) - sum(is.na(annotations_inc$SES.STDZD))))*100)`% (n = `r sum(annotations_inc$SES.STDZD==2 & !is.na(annotations_inc$SES.STDZD))`), exclusively represented families of middle or higher SES.
Most countries in our CHILDES sample are members of the OECD. According to a 2016 report, "Almost two-thirds of people live in middle-income households in OECD countries." While less than 66% of families from OECD countries are considered middle-to-high SES, over 80% of our sample falls into this category.
It's important to note that OECD countries generally have higher SES populations compared to the global average. Given this context, it's likely that CHILDES overrepresents middle-to-high SES families to an even greater extent when considered on a global scale.

#### Education
 `r round(sum(annotations_inc$Education.min>3 & !is.na(annotations_inc$Education.min))/sum(!is.na(annotations_inc$Education.min))*100)`% of the corpora (n =  `r round(sum(annotations_inc$Education.min>3 & !is.na(annotations_inc$Education.min)))`) include children whose parents had at least a graduate, if not a postgraduate, degree.  `r round(sum(annotations_inc$Education.min==1 & !is.na(annotations_inc$Education.min))/sum(!is.na(annotations_inc$Education.min))*100)`% (n=`r sum(annotations_inc$Education.min==1 & !is.na(annotations_inc$Education.min))` ) had at least some parents with primary-level education;  `r round(sum(annotations_inc$Education.min==2 & !is.na(annotations_inc$Education.min))/sum(!is.na(annotations_inc$Education.min))*100)`% (n=`r round(sum(annotations_inc$Education.min==2 & !is.na(annotations_inc$Education.min)))`) had parents with secondary school education as the lower bound of the education range, and a further `r round(sum(annotations_inc$Education.min==3 & !is.na(annotations_inc$Education.min))/sum(!is.na(annotations_inc$Education.min))*100)`% (n=`r round(sum(annotations_inc$Education.min==3 & !is.na(annotations_inc$Education.min)))`) had some college as the lower bound. `r sum(annotations_inc$Education.STDZD=="diverse",na.rm=T)` corpora were described as "diverse", without clarifying the range of education covered. These numbers do not accurately represent the demographics of the countries they were obtained from. For instance, while the U.S. Census Bureau reported that only 36% of the adult population held a bachelor’s degree or higher in 2020, our data indicates that 100% of the corpora of the parents from the United States had a college education or higher. As seen in Figure 3, the same result is seen for corpus from China, Denmark, Egypt, Hong Kong, Hungary, Israel, Italy, Japan, Mexico, The Netherlands, Poland, Russia, Singapore, South Korea, Switzerland and the United States where 100% of parents with data are college-educated or above. Thus, it seems that our sub-samples of CHILDES are very skewed toward higher-educated parents. 


```{r figure3 ,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Distribution of parental education by country."}

annotations_inc$Education.ac = factor(annotations_inc$Education.ac, levels=c("College and above", "Some college", "Some secondary","Some primary"  ))

# Plot this categorywise bar chart
g <- ggplot(data=subset(annotations_inc, !is.na(Education.ac)), aes(country))
# g + geom_bar(aes(fill=Education.ac), width = 0.6, position = "fill") + 
#   theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() +
#   labs(title="Education of caregivers") 
g + geom_bar(aes(fill=Education.ac), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() + 
  #facet_grid(~ scenario_f) +
  theme_minimal()  +
  #ggtitle("Education of caregivers")  + 
  theme(plot.title = element_text(size = 10, face = "bold"),strip.text.x = element_text(size = 10)) +
  coord_flip() + 
  #guides(fill=FALSE ) + 
  scale_fill_manual(breaks=c("Some primary", "Some secondary", "Some college", "College and above"),values = c("#E6FFFF", "#99E6FF", "#4CA6FF", "#0040FF" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of corpora with Education data") +
  labs(title="Education of caregivers", fill = "Education Level")


```

#### Occupation
Professions were overall varied. The majority, comprising`r round(sum(annotations_inc$is.education=="yes" & !is.na(annotations_inc$is.education))/sum(!is.na(annotations_inc$Parental.profession))*100)`% (n=`r sum(annotations_inc$is.education=="yes" & !is.na(annotations_inc$is.education))`), was associated with the field of education. Notably, within this category, `r round(sum(annotations_inc$is.academic=="yes" & !is.na(annotations_inc$is.academic))/sum(!is.na(annotations_inc$Parental.profession))*100)`% (n=`r sum(annotations_inc$is.academic=="yes" & !is.na(annotations_inc$is.academic))`) of the individuals were identified as parents with professions linked to graduate-level education. This included roles such as Master's or Ph.D. students, professors, linguists, researchers, scientists, and academics. Some of the other occupations reported were psychologists, speech therapist or home makers.



### Urbanization

`r round(sum(annotations_inc$Type.of.community=="urban" & !is.na(annotations_inc$Type.of.community))/sum(!is.na(annotations_inc$Type.of.community))*100)`% of the corpora (n=`r sum(annotations_inc$Type.of.community=="urban" & !is.na(annotations_inc$Type.of.community))`) were described as industrialized or urban, with one additional corpus described as both rural and urban [Rigol corpus, @lieven2013early]. Only 8% of the corpora (n = 5) were labeled as farming or rural, representing countries such as Canada, Lesotho, Jamaica, Papua New Guinea, and the United States. While this categorization may be arbitrary, it offers insight into the representation of each country within their corpora. For instance, countries like Lesotho and Papua New Guinea, where approximately 70% and 86% of the population resides in rural areas (World Bank estimates based on the United Nations Population Division's World Urbanization Prospects: 2018 Revision)<!-- not sure how to cite this https://data.worldbank.org/indicator/SP.RUR.TOTL.ZS?locations=LS -->, are categorized as rural corpora [@demuth2022acquisition ; @sarvasy2017grammar]. Conversely, the most urbanized countries in our sample, including Kuwait (`r na.omit(ind_all$SP.URB.TOTL.ZS[ind_all$Entity == "Kuwait"])[1]`%), Belgium (`r na.omit(ind_all$SP.URB.TOTL.ZS[ind_all$Entity == "Belgium"])[1]`%), and the Netherlands (`r na.omit(ind_all$SP.URB.TOTL.ZS[ind_all$Entity == "Netherlands"])[1]`%), are represented only by urban corpora.
Overall, the sample is predominantly composed of urban countries, with more than half of the countries being over 80% urban (see SM5 for a table of urban indicators by country), and the data reflects this urbanization tendency accordingly (see Figure 4).

```{r figure4,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Urbanization by country."}
annotations_inc$Type.of.community[annotations_inc$Type.of.community %in% c("academic","capital city of the Soviet Union","city","industial","industrial ","work-for-pay",
                                                                                                        "work-for-pay/industrial","industrial","Northeastern US urban","urban western", "industrial & service &trade activities...", "Mediterranean city of 1.6 million inhabitants", "modern city families, usually both working parents","Seaside tourist town","Industrial city (population: around 272,000 inhabitants)","mainly urban", "industrial city (3.000.000 inhabitants")]<-"urban"

annotations_inc$Type.of.community[annotations_inc$Type.of.community %in% c("farmer","rural","rural farming village")]<-"rural"
annotations_inc$Type.of.community[annotations_inc$Type.of.community %in% c("industrial, farmers")]<-"both"
annotations_inc$Type.of.community = factor(annotations_inc$Type.of.community, levels=c("urban", "both", "rural"))



# Plot this categorywise bar chart
gurb <- ggplot(data=subset(annotations_inc, !is.na(Type.of.community)), aes(country))
# g + geom_bar(aes(fill=Education.ac), width = 0.6, position = "fill") + 
#   theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() +
#   labs(title="Education of caregivers") 
gurb + geom_bar(aes(fill=Type.of.community), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() + 
  #facet_grid(~ scenario_f) +
  theme_minimal()  +
  #ggtitle("Education of caregivers")  + 
  theme(plot.title = element_text(size = 10, face = "bold"),strip.text.x = element_text(size = 10)) +
  coord_flip() + 
  #guides(fill=FALSE ) + 
  scale_fill_manual(breaks=c("rural", "both", "urban"),values = c("#BFAB57", "#966F56", "#808A8F" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of corpora with Urbanization data") +
  labs(title="Urbanization", fill = "Type of community")



```



### Family structure

```{r figure5,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Household by country."}
annotations_inc$Household.structure = factor(annotations_inc$Household.structure, levels=c("nuclear", "extended", "varied"))

# Plot this categorywise bar chart
ghouse <- ggplot(data=subset(annotations_inc, !is.na(Household.structure)), aes(country))
# g + geom_bar(aes(fill=Education.ac), width = 0.6, position = "fill") + 
#   theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() +
#   labs(title="Education of caregivers") 
ghouse + geom_bar(aes(fill=Household.structure), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() + 
  #facet_grid(~ scenario_f) +
  theme_minimal()  +
  #ggtitle("Education of caregivers")  + 
  theme(plot.title = element_text(size = 10, face = "bold"),strip.text.x = element_text(size = 10)) +
  coord_flip() + 
  #guides(fill=FALSE ) + 
  scale_fill_manual(breaks=c("nuclear", "varied", "extended"),values = c("#ADDFAD", "#D1BA97", "#F08080" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of corpora with Household data") +
  labs(title="Households", fill = "Type of househols arrangement")

```
#### Nuclear households
As for household composition, `r  sum(annotations_inc$Household.structure=="nuclear" & !is.na(annotations_inc$Household.structure))` corpora (`r round(sum(annotations_inc$Household.structure=="nuclear" & !is.na(annotations_inc$Household.structure))/sum(!is.na(annotations_inc$Household.structure))*100)`%) were based on nuclear families;  `r  sum(annotations_inc$Household.structure=="extended" & !is.na(annotations_inc$Household.structure))` extended families(`r round(sum(annotations_inc$Household.structure=="extended" & !is.na(annotations_inc$Household.structure))/sum(!is.na(annotations_inc$Household.structure))*100)`%); and in `r  sum(annotations_inc$Household.structure=="varied" & !is.na(annotations_inc$Household.structure))` corpora the structure was varied (`r round(sum(annotations_inc$Household.structure=="varied" & !is.na(annotations_inc$Household.structure))/sum(!is.na(annotations_inc$Household.structure))*100)`%). We compare side by side the average per country of the nuclear households (the sum of the percentages of couples with children households, and single parents with children households; United Nations, 2022). 


#### Sibling presence
A majority of corpora in our sub-sample of CHILDES include children who have siblings. In fact, only 28% of corpora (among the 93 corpora having information on siblings) were constituted exclusively of children with no siblings, and the remaining had at least one sibling, with the overall average being 0.8 siblings. Since 83% of countries in CHILDES are in the OECD, we draw a comparison point for such countries: 46% of children had no siblings in OECD countries according to 2015 data. In this sense, children with no siblings appear to be under-represented in CHILDES.

```{r figure6,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Siblings presence by country."}
annotations_inc$siblings[annotations_inc$Nb.of.children.with.siblings >= 1] <- "yes"
annotations_inc$siblings[annotations_inc$Nb.of.children.with.siblings == 0] <- "no"
annotations_inc$siblings[annotations_inc$Nb.of.children.with.siblings %in% c("all firstborn (some newborn sibs)", "at least 1 (triplet)")] <- "yes"
annotations_inc$siblings = factor(annotations_inc$siblings, levels=c("yes", "no"))


# Plot this categorywise bar chart
gsib <- ggplot(data=subset(annotations_inc, !is.na(siblings)), aes(country))
# g + geom_bar(aes(fill=Education.ac), width = 0.6, position = "fill") + 
#   theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() +
#   labs(title="Education of caregivers") 
gsib + geom_bar(aes(fill=siblings), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() + 
  #facet_grid(~ scenario_f) +
  theme_minimal()  +
  #ggtitle("Education of caregivers")  + 
  theme(plot.title = element_text(size = 10, face = "bold"),strip.text.x = element_text(size = 10)) +
  coord_flip() + 
  #guides(fill=FALSE ) + 
  scale_fill_manual(breaks=c("yes", "no"),values = c("#B5A6DF", "#000080" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of corpora with siblings data") +
  labs(title="Siblings presence", fill = "")

```
### Languages
We had two variables of interest here: language spoken in the corpus and lingual status. A total of 62 different languages or language combinations (for bilingual and multilingual children) were reportedly spoken in the corpora. 
```{r tab4,echo = FALSE, as.is=T, ig.pos = "T!",results="asis"}

# table4 <- annotations_inc %>%
#   group_by(Language) %>%
#   summarize(Total_Participants = sum(Nb.of.participants, na.rm = TRUE))
# 
# write.csv(table4, file = "../derived/table4.csv", row.names = FALSE)

mytab4 <- data.frame(
  Monolingual_Language = c("Afrikaans", "Arabic (Egyptian or Kuwaiti)", "Basque", "Cantonese", "Catalan", "Cree", "Croatian", "Czech", "Danish", "Dutch", "English", "Estonian", "Farsi", "French", "German", "Greek", "Hebrew", "Hungarian", "Icelandic", "Indonesian", "Irish", "Italian", "Jamaican", "Japanese", "Korean", "Mandarin", "Norwegian", "Nungon", "Polish", "Portuguese (Brazilian or European)", "Romanian", "Russian", "Serbian", "Sesotho", "Slovenian", "Spanish", "Swedish", "Taiwanese", "Tamil", "Thai", "Turkish", "Welsh"),
  Total_Participants = c(2, 80, 46, 8, 17, 1, 3, 6, 2, 23, 210, 32, 5, 46, 46, 6, 122, 8, 1, 8, 7, 8, 2, 148, 4, 157, 11, 5, 4, 9, 6, 2, 8, 4, 20, 75, 9, 4, 1, 18, 1, 475),
  Multilingual_Languages = c("Dutch/English, Dutch/French", "Dutch/Italian", "Spanish/Catalan", "Spanish/Galician", "English/Cantonese", "English/French", "English/Hebrew", "English/Japanese", "English/Japanese/Danish", "English/Mandarin", "English/Mandarin/Cantonese", "English/Russian", "English/Spanish", "French/Russian", "Hungarian/Catalan/Spanish", "Hungarian/Farsi/English", "German/Spanish", "Italian/Japanese", "Italian/German", "Portuguese/Swedish/English", NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA),
  Total_Participants_Multilingual = c(35, 4, 6, 66, 9, 12, 1, 1, 1, 55, 11, 22, 12, 1, 1, 2, 9, 1, 2, 3, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA)
)


apa_table(mytab4,caption="Total number of participants per language,", longtable = T, align=c("lccccccccr"))
#write.csv(mytab4,"output/table3.csv")
```

```{r figure7,echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "Indo-European languages by country."}
lang_fam<- read.csv("../derived/annotations_included_languagefamily.csv") 
# xtabs(~indo_european, lang_fam) indoeuropean 122 / no 58
#xtabs(~language_family, lang_fam)

# Plot this categorywise bar chart
glang_fam <- ggplot(data=lang_fam, aes(country))
# g + geom_bar(aes(fill=Education.ac), width = 0.6, position = "fill") + 
#   theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() +
#   labs(title="Education of caregivers") 
glang_fam + geom_bar(aes(fill=language_family), width = 0.9) + 
  theme(axis.text.x = element_text(vjust=0.4)) +  coord_flip() + 
  #facet_grid(~ scenario_f) +
  theme_minimal()  +
  #ggtitle("Education of caregivers")  + 
  theme(plot.title = element_text(size = 10, face = "bold"),strip.text.x = element_text(size = 10)) +
  coord_flip() + 
  #guides(fill=FALSE ) + 
  scale_fill_manual(breaks=c("indo_european", "other"),labels=c("Indo-European", "Other"), values = c("#CC5700", "#008080" )) +
  scale_x_discrete(name = "Countries or Territories") +
  scale_y_continuous(name = "Number of Indo-European corpora") +
  labs(title="Language Family", fill = "")

```

About a third (32%) of the included corpora that had available data for this variable (N = 110) were not monolingual. It is hard to find reliable estimates of the percentage of the population which is not monolingual in the world or in the countries represented in our sub-sample of CHILDES, but for instance, in Europe in 2016, 65% of adults reported knowing multiple languages (Eurostat, 2022). According to such estimates, even if samples are linguistically diverse, it would appear that input data in our sub-sample of CHILDES under-represents bilingual and multilingual households.



# Discussion
<!-- in 2020, 47,000 people were scientists in the USA (including both in academia and the private sector), 135,000 were professors, 20 million were Masters students, and 3 million were PhD students. Given that the USA’s population is 392 million, -->

\newpage
# References
