---
title: "childes_weird"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://rpubs.com/gloukatou/733656

TODO

- remove bilingualism
- fix %rur
- add family size



## R Markdown

```{r libraries, message=FALSE, warning=FALSE}

read.csv("childes_metrics.csv")->x

#devtools::install_github("unDocUMeantIt/koRpus.lang.fr") # stable release


#library(childesr)
#library(wordbankr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
require(scales)
library(stringr)
library(kableExtra)
library(WDI)


library(RColorBrewer)
myPalette <- brewer.pal(5, "Set2") 


group_count <-function(data_, column_){
data_<- data_ %>%
  group_by({{column_}}) %>%
    count() 
      return(data_)
}

group_select<-function(data_, column_){
data_ <- data_%>%
    group_by(!!column_) %>%
      select(Corpus, !!column_)
data_ <- data_[ order( data_[,3] ),]
return(data_)
}
```

# Load annotations 

```{r load annotations}

#####TODOS first part is the descriptive section, to what extent the full range of populations is represented, what are some codes that are missing often,
##### descriptive section and representation of the population
##########systematic review of features
##################confront eg proportion of child population versus proportion of data by eg continent
#############second goal: point out strengths and limitations of the current CHILDES, explain what is missing

annotations<- read.csv("Table for authors - annotations.csv") 
colnames(annotations)[colnames(annotations)=="X..of.children.with.siblings"]<-"Nb.of.children.with.siblings"
colnames(annotations)[colnames(annotations)=="X..of.children.with.older.siblings"]<-"Nb.of.children.with.older.siblings"
colnames(annotations)=gsub("\\.\\..*","",colnames(annotations))
colnames(annotations)[colnames(annotations)=="Our.coding.of"]<-"Our.coding.of.community.type"

annotations$Number.of.participants[annotations$Number.of.participants=="1 or 5"]<-NA
annotations$Number.of.participants[annotations$Number.of.participants=="1 CHI + MOT/FAT/INV"]<-1
annotations$Number.of.participants[annotations$Number.of.participants=="4 CHI + MOTs, some FATs, some SIBs"]<-4
annotations$Number.of.participants=as.numeric(as.character(annotations$Number.of.participants))

annotations$Nb.of.children.with.siblings=as.numeric(as.character(annotations$Nb.of.children.with.siblings))

annotations$Proportion.With.Siblings=annotations$Nb.of.children.with.siblings/annotations$Number.of.participants

#table(annotations$Household.structure)
annotations$Household.structure=tolower(annotations$Household.structure)
annotations$Household.structure[grep("different|diverse|varied",annotations$Household.structure)] <- NA #turn to NA?
annotations$Household.structure[annotations$Household.structure %in% c("nuclear, extended","extended/nuclear")]<-"varied"
annotations$Household.structure[grep("nuclear|nucear|single",annotations$Household.structure)]<-"nuclear" # including single parent
annotations$Household.structure[grep("extended",annotations$Household.structure)]<-"extended"
annotations$Household.structure[annotations$Household.structure==""]<-NA


# Education & SES cleaning
annotations$Education.Naomi[annotations$Education.Naomi==""]<-NA
annotations$Education.min<-gsub("-.*","",annotations$Education.Naomi)
annotations$Education.max<-gsub(".*-","",annotations$Education.Naomi)
annotations$Education.mid<-as.numeric(as.character(annotations$Education.min))+(as.numeric(as.character(annotations$Education.max))-as.numeric(as.character(annotations$Education.min)))/2

annotations$SES.Naomi[annotations$SES.Naomi==""]<-NA
annotations$SES.min<-gsub("-.*","",annotations$SES.Naomi)
annotations$SES.max<-gsub(".*-","",annotations$SES.Naomi)
annotations$SES.max[grep("upper",annotations$Parental.socioeconomic.status)]<-3

annotations$SES.mid<-as.numeric(as.character(annotations$SES.min))+(as.numeric(as.character(annotations$SES.max))-as.numeric(as.character(annotations$SES.min)))/2


```

`r nrow(annotations)` corpora annotated. 



```{r group_by inclusion}

include<- annotations %>%
  mutate(inclusion=tolower(Inclusion)) %>%
  mutate(inclusion = ifelse(str_detect(inclusion, "yes"), "yes", "no")) %>%
  group_by(inclusion) %>%
  count() 

ggplot(include,aes(y=n, x=inclusion)) +
  geom_bar(stat="identity", width=1, color="white") +
  theme_minimal() + 
  ggtitle("Corpus inclusion") 

annotations <- annotations %>%
  filter(Inclusion=="yes")

```
From the total amount of corpora, `r nrow(annotations)` have been included in this study. The following analyses will continue only on these included corpora.



```{r group_by location_cleaning}

annotations$country <-NA
annotations$continent <-NA


annotations$country[grep("Ireland", annotations$Location)]  <- "Ireland"
annotations$country[grep("England|Arfon area Gwynedd, North Wales|Belfast, Northern Ireland|Nottingham/Manchester, England|England, Brighton|Wales", annotations$Location)]  <- "United Kingdom"
annotations$country[grep("Portugal", annotations$Location)]  <- "Portugal"
annotations$country[grep("Spain|Madrid, Spain ; Tenerife, Canary Islands|Madrid, Spain|Navarra, Spain|Alt penedes, region of catalonia|spain| Lloret de mar|Barcelona| SPAIN|Salamanca|Lugo", annotations$Location)]  <- "Spain"
annotations$country[grep("France|France (Normandy, Marseille, + places visited)", annotations$Location)]  <- "France"
annotations$country[grep("Naples|italy|Roma", annotations$Location)]  <- "Italy"
annotations$country[grep("Switzerland", annotations$Location)]  <- "Switzerland"
annotations$country[grep("Belgium", annotations$Location)]  <- "Belgium"
annotations$country[grep("Germany", annotations$Location)]  <- "Germany"
annotations$country[grep("Netherlands", annotations$Location)]  <- "Netherlands"

#Following the EuroVoc classification https://en.wikipedia.org/wiki/EuroVoc + Italy and Spain and Portugal
annotations$continent[grep("Andorra|Austria|Belgium|France|Germany|Ireland|Italy|Liechtenstein|Luxembourg|Monaco|Netherlands|Portugal|Spain|Switzerland|United Kingdom", annotations$country)]  <- "Western Europe"


annotations$country[grep("Poznań, Poland", annotations$Location)]  <- "Poland"
annotations$country[grep("Estonia|Tartu, Estonia ; Rapla, Estonia|Southern Estonia|Tartu, Estonia", annotations$Location)]  <- "Estonia"
annotations$country[grep("Hungary", annotations$Location)]  <- "Hungary"
annotations$country[grep("Czech Republic", annotations$Location)]  <- "Czech Republic"
annotations$country[grep("Bucharest, Romania|Romania", annotations$Location)]  <- "Romania"
annotations$country[grep("Serbia", annotations$Location)]  <- "Serbia"
annotations$country[grep("Croatia", annotations$Location)]  <- "Croatia"
annotations$country[grep("Slovenia", annotations$Location)]  <- "Slovenia"

annotations$country[grep("Sweden", annotations$Location)]  <- "Sweden"
annotations$country[grep("Iceland", annotations$Location)]  <- "Iceland"
annotations$country[grep("Norway", annotations$Location)]  <- "Norway"
annotations$country[grep("Denmark", annotations$Location)]  <- "Denmark"

annotations$country[grep("Athens, Greece", annotations$Location)]  <- "Greece"

annotations$country[grep("Moscow, Russia", annotations$Location)]  <- "Russia"

annotations$continent[grep("Poland|Estonia|Hungary|Czech|Romania|Serbia|Croatia|Slovenia|Sweden|Iceland|Norway|Denmark|Greece|Russia", annotations$country)]  <- "Non-Western Europe"


annotations$country[grep("Turkey", annotations$Location)]  <- "Turkey"
annotations$country[grep("Iran", annotations$Location)]  <- "Iran"
annotations$country[grep("Israel|Yahud, Israel", annotations$Location)]  <- "Israel"
annotations$country[grep("Kuwait", annotations$Location)]  <- "Kuwait"

annotations$country[grep("Bombay, India", annotations$Location)]  <- "India"

annotations$country[grep("China", annotations$Location)]  <- "China"

annotations$country[grep("Singapore", annotations$Location)]  <- "Singapore"
annotations$country[grep("Indonesia", annotations$Location)]  <- "Indonesia"
annotations$country[grep("Bangkok, Thailand", annotations$Location)]  <- "Thailand"
annotations$country[grep("Osaka|Tokyo|Nagoya|osaka|Kusatsu City, Shiga Pref", annotations$Location)]  <- "Japan"
annotations$country[grep("Korea", annotations$Location)]  <- "Korea" #NOTE: one corpus just says "Korea", data collected in 2009-2011, assuming it's South Korean
annotations$country[grep("Hong-Kong", annotations$Location)]  <- "Hong Kong"
annotations$country[grep("Taiwan", annotations$Location)]  <- "Taiwan"

annotations$continent[grep("Turkey|Iran|Israel|Kuwait|India|China|Singapore|Indonesia|Thailand|Japan|Korea|Hong Kong|Taiwan", annotations$country)]  <- "Asia"


annotations$country[grep("Papua-New Guinea", annotations$Location)]  <- "Papua New Guinea"
annotations$continent[grep("Papua New Guinea", annotations$country)]  <- "Oceania"


annotations$country[grep("Alexandria, Egypt", annotations$Location)]  <- "Egypt"
annotations$country[grep("Mokhotlong, Lesotho", annotations$Location)]  <- "Lesotho"
annotations$country[grep("South Africa", annotations$Location)]  <- "South Africa"
annotations$continent[grep("Egypt|Lesotho|Africa", annotations$country)]  <- "Africa"


annotations$country[grep("Rio Cuarto, Cordoba, Argentina", annotations$Location)]  <- "Argentina"
annotations$country[grep("Sao Paulo", annotations$Location)]  <- "Brazil"
annotations$country[grep("Mexico", annotations$Location)]  <- "Mexico"
annotations$country[grep("Jamaica", annotations$Location)]  <- "Jamaica"
annotations$continent[grep("Argentina|Brazil|Mexico|Jamaica", annotations$country)]  <- "Latin America"

annotations$country[grep("Michigan, USA|USA, Northern Virginia|California, USA|washington dc|United States|USA|Washington|Maryland|San Fran", annotations$Location)]  <- "United States"
annotations$country[grep("Canada", annotations$Location)]  <- "Canada"
annotations$continent[grep("United States|Canada", annotations$country)]  <- "North America"


#Special cases
annotations$country[grep("Sweden ; Portugal", annotations$Location)]  <- "Sweden & Portugal"
annotations$country[grep("Spain (Lloret de Mar), Hungary (Kecskemét)", annotations$Location)]  <- "Spain & Hungary"
#We are leaving continent as NA (because one country is Western & the other Eastern Europe)


#to check if any left
#annotations[is.na(annotations$country),"Location"]
#annotations[is.na(annotations$continent),"Location"]





```

```{r cleaning type}

annotations$Type.of.community.at.the.time.of.the.recordings[is.na(annotations$Type.of.community.at.the.time.of.the.recordings)]<-annotations$Naomi.community.type[is.na(annotations$Type.of.community.at.the.time.of.the.recordings)]

annotations$Type.of.community.at.the.time.of.the.recordings[annotations$Type.of.community.at.the.time.of.the.recordings %in% c("academic","capital city of the Soviet Union","city","industial","industrial ","work-for-pay",
                                                                                                                               "industrial & service &trade activities...","industrial city (3.000.000 inhabitants",
                                                                                                                               "Industrial city (population: around 272,000 inhabitants)",
                                                                                                                               "Mediterranean city of 1.6 million inhabitants",
                                                                                                                               "modern city families, usually both working parents",
                                                                                                                               "Orthodox Jews",
                                                                                                                               "Seaside tourist town",
                                                                                                                               "work-for-pay/industrial","industrial")]<-"urban"

annotations$Type.of.community.at.the.time.of.the.recordings[annotations$Type.of.community.at.the.time.of.the.recordings %in% c("farmer","rural","rural farming village")]<-"rural"

annotations$Type.of.community.at.the.time.of.the.recordings[annotations$Type.of.community.at.the.time.of.the.recordings %in% c("industrial, farmers")]<-"both"
table(annotations$Type.of.community.at.the.time.of.the.recordings)
```



```{r ancillary data}
byPart<- annotations %>%
  select(Language.or.Languages.spoken.in.recordings, Number.of.participants) %>%
   filter(!is.na(Number.of.participants)) %>%
     group_by(Language.or.Languages.spoken.in.recordings) %>%
       summarise(numpar = sum(as.numeric(Number.of.participants))) 


read.csv("country_iso.csv")->iso_lookup
codes=iso_lookup$Code
iso_lookup$Name<-as.character(iso_lookup$Name)
iso_lookup$Name=gsub(", Islamic Republic of","",iso_lookup$Name)
iso_lookup$Name=gsub("Russian Federation","Russia",iso_lookup$Name)
iso_lookup$Name=gsub(", Province of China","",iso_lookup$Name)
iso_lookup$Name=gsub(", Republic of","",iso_lookup$Name)

names(codes)<-iso_lookup$Name

annotations$country_WB<-codes[annotations$country]

annotations$country_WB <-factor(annotations$country_WB)
annotations[is.na(annotations$country_WB),"country"]


annotations$continent <-factor(annotations$continent)
# Indices
# WDIsearch('gdp')
#"GDP per capita, PPP (constant 2005 international $)"
# WDIsearch('Educational attainment, at least')
# Naomi: secondary, some college, university, 
# WDI:  SE.SEC.CUAT.LO.ZS    "Educational attainment, at least completed lower secondary, population 25+, total (%) (cumulative)" -- 62 countries are NOT NA
# WDI:  SE.SEC.CUAT.UP.ZS   Educational attainment, at least completed upper secondary, population 25+, total (%) (cumulative) -- 66 countries are NOT NA
# WDI:  SE.TER.CUAT.BA.ZS   Educational attainment, at least Bachelor's or equivalent, population 25+, total (%) (cumulative) -- only 25 countries are not NA
# note, only 23 countries have all three... 

## ADD
# "SH.FPL.IDLC.Q3" "Mean ideal number of children (per woman): Q3"
# "SP.DYN.CEBN.Q3" "Mean number of children ever born to women aged 40-49: Q3" 


dat_all = WDI(indicator=c('NY.GDP.PCAP.PP.KD','SE.SEC.CUAT.UP.ZS','SP.RUR.TOTL.ZS','5.01.01.01.indust'), start=2012, end=2012)
# flip %rur

dat = WDI(indicator=c('NY.GDP.PCAP.PP.KD','SE.SEC.CUAT.UP.ZS','SP.RUR.TOTL.ZS','5.01.01.01.indust'), country=c(levels(annotations$country_WB)), start=2012, end=2012)

annotations=merge(annotations,dat,by="country",all.x=T)

table(!is.na(annotations$SES.mid),!is.na(annotations$NY.GDP.PCAP.PP.KD))
table(!is.na(annotations$Education.mid),!is.na(annotations$SE.SEC.CUAT.UP.ZS))
table(!is.na(annotations$Type.of.community.at.the.time.of.the.recordings),!is.na(annotations$SP.RUR.TOTL.ZS))
            


```

# At the level of the population

## Western

There are corpora for every continent, although more corpora in Europe than other continents:

```{r tab x cont}
kable(table(annotations$continent))

```


```{r nch x cont}
mytab=aggregate(annotations$Number.of.participants,by=list(annotations$continent),sum)
colnames(mytab)<-c("Continent","Number of children")
kable(mytab)

```
## Educated



```{r}
dens_dat=data.frame(rbind(cbind(dat_all$SE.SEC.CUAT.UP.ZS[!is.na(dat_all$SE.SEC.CUAT.UP.ZS)],0),
               cbind(dat$SE.SEC.CUAT.UP.ZS[!is.na(dat$SE.SEC.CUAT.UP.ZS)],1) ))
colnames(dens_dat)<-c("completed.HS","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

ggplot(dens_dat, aes(x = completed.HS, fill = in.childes)) + geom_density(alpha = 0.5)


```


## Urban versus rural

```{r }
dens_dat=data.frame(rbind(cbind(dat_all$SP.RUR.TOTL.ZS[!is.na(dat_all$SP.RUR.TOTL.ZS)],0),
               cbind(dat$SP.RUR.TOTL.ZS[!is.na(dat$SP.RUR.TOTL.ZS)],1) ))
colnames(dens_dat)<-c("pc.rural","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

ggplot(dens_dat, aes(x = pc.rural, fill = in.childes)) + geom_density(alpha = 0.5)

```


## Rich



```{r}

dens_dat=data.frame(rbind(cbind(dat_all$NY.GDP.PCAP.PP.KD[!is.na(dat_all$NY.GDP.PCAP.PP.KD)],0),
               cbind(dat$NY.GDP.PCAP.PP.KD[!is.na(dat$NY.GDP.PCAP.PP.KD)],1) ))
colnames(dens_dat)<-c("GDP.per.capita","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

ggplot(dens_dat, aes(x = GDP.per.capita, fill = in.childes)) + geom_density(alpha = 0.5)

```


## Monolingualism

```{r group_by bilinguals_community}
annotations$Bilingualism.Multilingualism.in.community[annotations$Bilingualism.Multilingualism.in.community %in% c(" yes (bilingualism Galician-Spanish)","bilingual (Arabic/English)","probably","Yes (Spanish/Catalan)")] <- "yes"
annotations$Bilingualism.Multilingualism.in.community[annotations$Bilingualism.Multilingualism.in.community %in% c("Minimal","slight","yes (but minimal)","only for Jong, other monolinguals")] <- "minimal"
annotations$Bilingualism.Multilingualism.in.community<-tolower(annotations$Bilingualism.Multilingualism.in.community)
annotations$Bilingualism.Multilingualism.in.community[annotations$Bilingualism.Multilingualism.in.community==""]<-NA

byBilinguals<-group_count(annotations, Bilingualism.Multilingualism.in.community)

ggplot(byBilinguals,aes(y=n, x=Bilingualism.Multilingualism.in.community)) +
  geom_bar(stat="identity", width=1, color="white") +
  theme_minimal() + 
  ggtitle("Bilingualism/Multilingualism in the community") 

```


# At the level of the family

## Educated



```{r}
plot(annotations$Education.mid~annotations$SE.SEC.CUAT.UP.ZS,xlab="Proportion of the population finishing high school",ylab="Education level in the CHILDES sample",yaxt="n")
axis(2,at=1:5,labels=c("primary", "secondary","some college","college","postgraduate"),las=2)
```


## Urban versus rural



```{r tab x rur/urb}
kable(table(annotations$Type.of.community.at.the.time.of.the.recordings))
#annotations[!is.na(annotations$Type.of.community.at.the.time.of.the.recordings) & !is.na(annotations$SP.RUR.TOTL.ZS),c("Type.of.community.at.the.time.of.the.recordings","SP.RUR.TOTL.ZS")]

boxplot(annotations$SP.RUR.TOTL.ZS ~ annotations$Type.of.community.at.the.time.of.the.recordings,type='n')
points(annotations$SP.RUR.TOTL.ZS ~ jitter(as.numeric(as.factor(annotations$Type.of.community.at.the.time.of.the.recordings))))
```


```{r nch x rur/urb}
mytab=aggregate(annotations$Number.of.participants,by=list(annotations$Type.of.community.at.the.time.of.the.recordings),sum)
colnames(mytab)<-c("Setting","Number of children")
kable(mytab)

```


## Rich

This graph isn't good, but we don't know exactly how to convey this.


```{r, eval=F}
plot(annotations$SES.mid~annotations$NY.GDP.PCAP.PP.KD,xlab="GDP per capita",ylab="SES in the CHILDES sample",yaxt="n")
axis(2,at=1:3,labels=c("lower", "middle","upper"),las=2)
```

## Household structure

```{r nuclear vs extended}
table(annotations$Household.structure)
```


## Siblings

- change to histogram
- show proportion of corpora that have at least one child with sibling

```{r prop kids with sibs}

summary(annotations$Proportion.With.Siblings)
plot(density(annotations$Proportion.With.Siblings[!is.na(annotations$Proportion.With.Siblings)]))
```



```{r group_by siblings, warning=FALSE, message=FALSE}

bySiblnum<-annotations %>%
    group_by(Average.number.of.siblings) %>%
      count()

ggplot(bySiblnum, mapping = aes(x=Average.number.of.siblings, y=n)) + 
  geom_point() + 
    ggtitle("Average number of siblings per corpus") + 
      geom_col()

#ggplot(bySiblnum[!is.na(bySiblnum$Average.number.of.siblings),], mapping = aes(x=Average.number.of.siblings, y=n)) 

annotations$Average.number.of.siblings <- as.numeric(as.character(annotations$Average.number.of.siblings))
head(bySiblnum)
```

The mean number of  siblings has been left NA for `r bySiblnum[is.na(bySiblnum$Average.number.of.siblings),]$n` out of `r sum(bySiblnum$n)` corpora. In the remaining corpora, the mean number of siblings per corpus is `r mean(annotations[!is.na(annotations$Average.number.of.siblings),]$Average.number.of.siblings, na.rm=TRUE)` with min `r min(annotations[!is.na(annotations$Average.number.of.siblings),]$Average.number.of.siblings)` and max `r max(annotations[!is.na(annotations$Average.number.of.siblings),]$Average.number.of.siblings)`.


## Bilingualism/Multilingualism

```{r group_by bilinguals}

annotations$Bilingualism.Multilingualism.in.family[annotations$Bilingualism.Multilingualism.in.family %in% c("1 Portuguese-French; 1 Russian-French","participants were included as en or fr when their language exposure across settings consisted of 75% or more of the respective language by parent report")]<-"all"
annotations$Bilingualism.Multilingualism.in.family[annotations$Bilingualism.Multilingualism.in.family %in% c("only for Jong, other monolinguals","1")]<-"some"
annotations$Bilingualism.Multilingualism.in.family<-tolower(annotations$Bilingualism.Multilingualism.in.family)
annotations$Bilingualism.Multilingualism.in.family[grep("yes",annotations$Bilingualism.Multilingualism.in.family)]<-"all"
annotations$Bilingualism.Multilingualism.in.family[grep("bilingual",annotations$Bilingualism.Multilingualism.in.family)]<-"all"
annotations$Bilingualism.Multilingualism.in.family[grep("no",annotations$Bilingualism.Multilingualism.in.family)]<-"none"
annotations$Bilingualism.Multilingualism.in.family[annotations$Bilingualism.Multilingualism.in.family=="parents are linguists; fluent in spanish, do field work on turkish, mayan languages, possibly more."]<-"none"
annotations$Bilingualism.Multilingualism.in.family[annotations$Bilingualism.Multilingualism.in.family %in% c("","unknown","?")]<-NA

byBilinguals<-group_count(annotations, Bilingualism.Multilingualism.in.family)

ggplot(byBilinguals,aes(y=n, x=Bilingualism.Multilingualism.in.family)) +
  geom_bar(stat="identity", width=1, color="white") +
  theme_minimal() + 
  ggtitle("Bilingualism/Multilingualism in family") 


```





# Old information



## Based on "Fertility.rate.of.community.at.the.time.of.the.recordings, Interbirth intervals, Schooling, Community, Healty, Culture, Mortality, Household..." annotation columns
### Mostly NAs. For now just prints column 

```{r group_by weird_var, warning=FALSE, message=FALSE}

byFertility <- group_select(annotations, "Fertility.rate.of.community.at.the.time.of.the.recordings")
byFertility <- byFertility %>% group_by(Fertility.rate.of.community.at.the.time.of.the.recordings) %>%  summarise(n = n())
byFertility
```

```{r group_by weird_var2, warning=FALSE, message=FALSE}

byInterbirth <- group_select(annotations, "Interbirth.intervals.in.community.at.the.time.of.the.recordings")
byInterbirth <- byInterbirth %>% group_by(Interbirth.intervals.in.community.at.the.time.of.the.recordings) %>%  summarise(n = n())
byInterbirth

bySchooling <- group_select(annotations, "Access.to.schooling.for.recorded.children")
bySchooling <- bySchooling %>% group_by(Access.to.schooling.for.recorded.children) %>%  summarise(n = n())
bySchooling

byCommunity <- group_select(annotations, "Type.of.community.at.the.time.of.the.recordings")
byCommunity <- byCommunity %>% group_by(Type.of.community.at.the.time.of.the.recordings) %>%  summarise(n = n())
byCommunity

byHealth <- group_select(annotations, "Access.to.health.service.for.recorded.children")
byHealth <- byHealth %>% group_by(Access.to.health.service.for.recorded.children) %>%  summarise(n = n())
byHealth 

#byCulture <- group_select(annotations, "type_culture")
#byCulture <- byCulture %>% group_by(Type.of.culture..minority..majority.) %>%  summarise(n = n())
#byCulture

#byMortality<- group_select(annotations, "Mortality.rate.of.children.in.the.community")
#byMortality<- byMortality %>% group_by(Mortality.rate.of.children.in.the.community) %>%  summarise(n = n())
#byMortality

byLangSpeakers<- group_select(annotations, "Number.of.speakers.of.the.language")
byLangSpeakers<- byLangSpeakers %>% group_by(Number.of.speakers.of.the.language) %>%  summarise(n = n())
byLangSpeakers

byHousehold<- group_select(annotations, "Household.structure")
byhousehold<- byHousehold %>% group_by(Household.structure) %>%  summarise(n = n())
byhousehold

#byTimespent<- group_select(annotations, #"Where.children.spend.their.time..home..nursery..playing.by.the.river....")
#byTimespent<- byTimespent %>% #group_by(Where.children.spend.their.time..home..nursery..playing.by.the.river....) %>%  summarise(n = n())
#byTimespent

```






### Corpora by countries and continents
```{r group_by Location}

byLocation <- annotations %>%
  select(country, continent) %>%
    group_by(country, continent) %>%
      count()

ggplot(byLocation, aes(x=n, y=country)) +
  geom_bar(stat="identity", width=1, color="white") +
    ggtitle("Corpora by Countries and Continents") +
      facet_wrap(~continent, scale="free") 

head(byLocation)

```

`r length(unique(byLocation$country))` countries are represented in childes. `r length(unique(byLocation$continent))-1` continents. 
`r nrow(annotations[annotations$continent=="Europe",])` of the `r nrow(annotations)` corpora are European, and
`r nrow(annotations[annotations$continent=="North America",])` of the corpora are North American. 

## Based on "Location" annotation column
### Corpora by countries and continents
```{r group_by location1}
byLocation1 <- group_count(annotations, continent)
lbls <- paste(byLocation1$continent, byLocation1$n, sep=" ") # add numbers to labels
pie(byLocation1$n, labels=lbls, border="white", col=myPalette, main="Corpora by continent")

head(byLocation1)

```


## Recording age
### Based on "Mean.child.age.at.beginning.of.recordings, Mean.child.age.at.end.of.recordings" annotation columns

This chunk not working because it uses old column naming convention, and Alex thought it should be replaced with something that draws age from the .cha, rather than human inputting.

```{r group_by age, warning=FALSE, message=FALSE, eval=F}
byAge <- annotations %>% 
 select(first_age, last_age, country, continent) %>%
  group_by(country, continent) %>%
    summarize(min = min(first_age), max = max(last_age)) %>%
       arrange(desc(min)) 


ggplot(byAge, aes(y=country))+
  geom_linerange(aes(xmin=min,xmax=max),linetype=2,color="blue")+
    geom_point(aes(x=min),size=3,color="red")+
      geom_point(aes(x=max),size=3,color="red")+
        ggtitle("Nim and Max age per country") +
          theme_bw()+
             facet_wrap(~continent, scale="free") 
            
```





## Number of children in CHILDES
### Based on "Number.of.participants" annotation column
#### Number of target children per language

```{r group_by participants}

byPart<- annotations %>%
  select(Language.or.Languages.spoken.in.recordings, Number.of.participants, Bilingualism.Multilingualism.in.corpus) %>%
   filter(!is.na(Number.of.participants)) %>%
     group_by(Language.or.Languages.spoken.in.recordings, Bilingualism.Multilingualism.in.corpus) %>%
       summarise(numpar = sum(as.numeric(Number.of.participants))) 
#     p

byPart %>%      
  ggplot(mapping = aes(y = reorder(Language.or.Languages.spoken.in.recordings, numpar), numpar)) + 
    geom_point() + 
      ggtitle("Number of target children per language") + 
        geom_col()+ 
          facet_wrap(~Bilingualism.Multilingualism.in.corpus, scale="free") +
            labs(y= "Languages", x = "Number of target children")

byPart_nona <- byPart[!is.na(byPart$Bilingualism.Multilingualism.in.corpus),]
byPart_mono <- byPart_nona %>%
  filter(Bilingualism.Multilingualism.in.corpus=="Monolinguals")
byPart_bili <- byPart_nona %>%
  filter(Bilingualism.Multilingualism.in.corpus=="Bilinguals")

head(byPart)
```

`r nrow(annotations[is.na(annotations$Number.of.participants), ])` corpora have NA in "Number of participants". 
There is data from `r sum(byPart_mono$numpar)` target children in Monolingual corpora.
There is data from `r sum(byPart[is.na(byPart$Bilingualism.Multilingualism.in.corpus),]$numpar)` target children in corpora which have not been classified as Monolingual/Bilingual.
There is data from `r sum(byPart_bili$numpar)` target children in Bilingual corpora.
Overall, the mean number of target children per language is `r mean(byPart$numpar)` with min `r min(byPart$numpar)` and max `r max(byPart$numpar)`.

### Based on "Number.of.participants" annotation column
#### Number of target children per continent

```{r group_by participants_continent}

byPartCont <- annotations %>%
    group_by(continent) %>%
      select(Number.of.participants, continent) %>%
      summarise(numpar = sum(as.numeric(Number.of.participants), na.rm=TRUE)) 

byPartCont %>%      
  ggplot(mapping = aes(y = reorder(continent, numpar), numpar)) + 
      ggtitle("Number of target children per continent") + 
        geom_col()+
           labs(y= "Continents", x = "Number of target children")
head(byPartCont)
```

`r byPartCont[byPartCont$continent=="Europe",]$numpar[1]` target children in Europe, `r byPartCont[byPartCont$continent=="Africa",]$numpar[1]` in Africa,
`r byPartCont[byPartCont$continent=="North America",]$numpar[1]` in North America, `r byPartCont[byPartCont$continent=="Asia",]$numpar[1]`in Asia, `r byPartCont[byPartCont$continent=="Central and South America",]$numpar[1]` in Central and South America.

## Optional chunk merging with childesr for extra information

```{r load_join childes, eval=FALSE}
#d_transcripts <- get_transcripts()
#d_statistics <- get_speaker_statistics()

#mergedChiAnnot <- annotations %>% left_join(d_transcripts)
#mergedChiAnnot <- mergedChiAnnot %>% left_join(n_utt_d)
#mergedChiAnnot <- mergedChiAnnot %>% left_join(n_utt_speaker)
#mergedChiAnnot<-mergedChiAnnot[!(is.na(mergedChiAnnot$n_utt) | mergedChiAnnot$n_utt=="" ), ]

#if ( length(unique(mergedChiAnnot$Corpus)) < nrow(annotations) ) {
#annotations$Corpus[!annotations$Corpus %in% mergedChiAnnot$Corpus]
#  } #print non merged corpora
```

### bilingualism

```{r}
annotations$Bilingualism.Multilingualism.in.corpus[annotations$Bilingualism.Multilingualism.in.corpus %in% c("trilinguism")]<-"yes"
annotations$Bilingualism.Multilingualism.in.corpus<-tolower(annotations$Bilingualism.Multilingualism.in.corpus)
annotations$Bilingualism.Multilingualism.in.corpus[grep("yes",annotations$Bilingualism.Multilingualism.in.corpus)]<-"yes"
annotations$Bilingualism.Multilingualism.in.corpus[grep("no",annotations$Bilingualism.Multilingualism.in.corpus)]<-"no"
annotations$Bilingualism.Multilingualism.in.corpus[annotations$Bilingualism.Multilingualism.in.corpus==""]<-NA

byBilinguals<-group_count(annotations, Bilingualism.Multilingualism.in.corpus)

ggplot(byBilinguals,aes(y=n, x=Bilingualism.Multilingualism.in.corpus)) +
  geom_bar(stat="identity", width=1, color="white") +
  theme_minimal() + 
  ggtitle("Bilingualism in corpus") 
```

