---
title: "childes_weird"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
```



```{r libraries, message=FALSE, warning=FALSE}


library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
require(scales)
library(stringr)
library(kableExtra)
library(ggpubr)
library(plotly)
library(GGally)
library(ggthemes)
library(rjson)


library(RColorBrewer)
myPalette <- brewer.pal(5, "Set2") 


group_count <-function(data_, column_){
data_<- data_ %>%
  group_by({{column_}}) %>%
    count() 
      return(data_)
}

group_select<-function(data_, column_){
data_ <- data_%>%
    group_by(!!column_) %>%
      select(Corpus, !!column_)
data_ <- data_[ order( data_[,3] ),]
return(data_)
}
```

# Load annotations 

```{r load annotations}

annotations<- read.csv("Table for authors - annotations.csv") 
colnames(annotations)[colnames(annotations)=="X..of.children.with.siblings"]<-"Nb.of.children.with.siblings"
colnames(annotations)[colnames(annotations)=="X..of.children.with.older.siblings"]<-"Nb.of.children.with.older.siblings"
colnames(annotations)=gsub("\\.\\..*","",colnames(annotations))
colnames(annotations)[colnames(annotations)=="Our.coding.of"]<-"Our.coding.of.community.type"

annotations$Number.of.participants[annotations$Number.of.participants=="1 or 5"]<-NA
annotations$Number.of.participants[annotations$Number.of.participants=="1 CHI + MOT/FAT/INV"]<-1
annotations$Number.of.participants[annotations$Number.of.participants=="4 CHI + MOTs, some FATs, some SIBs"]<-4
annotations$Number.of.participants=as.numeric(as.character(annotations$Number.of.participants))

annotations$Nb.of.children.with.siblings=as.numeric(as.character(annotations$Nb.of.children.with.siblings))

annotations$Proportion.With.Siblings=annotations$Nb.of.children.with.siblings/annotations$Number.of.participants

#table(annotations$Household.structure)
annotations$Household.structure=tolower(annotations$Household.structure)
annotations$Household.structure[grep("different|diverse|varied",annotations$Household.structure)] <- NA #turn to NA?
annotations$Household.structure[annotations$Household.structure %in% c("nuclear, extended","extended/nuclear")]<-"varied"
annotations$Household.structure[grep("nuclear|nucear|single",annotations$Household.structure)]<-"nuclear" # including single parent
annotations$Household.structure[grep("extended",annotations$Household.structure)]<-"extended"
annotations$Household.structure[annotations$Household.structure==""]<-NA


# Education & SES cleaning
annotations$Education.Naomi[annotations$Education.Naomi==""]<-NA
annotations$Education.Naomi[annotations$Education.Naomi=="4-6"]<-"4-5" #fix typo
annotations$Education.min<-as.numeric(as.character(gsub("-.*","",annotations$Education.Naomi)))
annotations$Education.max<-as.numeric(as.character(gsub(".*-","",annotations$Education.Naomi)))
annotations$Education.mid<-annotations$Education.min+(annotations$Education.max-annotations$Education.min)/2
# 1 = primary
# 2 = secondary school
# 3 = some college
# 4 = university
# 5 = postgraduate

annotations$Education.ac=NA
annotations$Education.ac[annotations$Education.min==1 & !is.na(annotations$Education.min)]<-"Some primary"
annotations$Education.ac[annotations$Education.min==2 & !is.na(annotations$Education.min)]<-"Some secondary"
annotations$Education.ac[annotations$Education.min==3 & !is.na(annotations$Education.min)]<-"Some college"
annotations$Education.ac[annotations$Education.min>3 & !is.na(annotations$Education.min)]<-"College and above"

annotations$SES.Naomi[annotations$SES.Naomi==""]<-NA
annotations$SES.Naomi[annotations$SES.Naomi=="middle class"]<-2


```


```{r group_by inclusion}

# include<- annotations %>%
#   mutate(inclusion=tolower(Inclusion)) %>%
#   mutate(inclusion = ifelse(str_detect(inclusion, "yes"), "yes", "no")) %>%
#   group_by(inclusion) %>%
#   count() 

# ggplot(include,aes(y=n, x=inclusion)) +
#   geom_bar(stat="identity", width=1, color="white") +
#   theme_minimal() + 
#   ggtitle("Corpus inclusion") 

all_annotations <- annotations

annotations <- all_annotations %>%
  filter(Inclusion=="yes")

table(all_annotations$why_exclude)->excl

```






```{r group_by location_cleaning}

annotations$country <-NA
annotations$continent <-NA


annotations$country[grep("Ireland", annotations$Location)]  <- "Ireland"
annotations$country[grep("England|Arfon area Gwynedd, North Wales|Belfast, Northern Ireland|Nottingham/Manchester, England|England, Brighton|Wales", annotations$Location)]  <- "United Kingdom"
annotations$country[grep("Portugal", annotations$Location)]  <- "Portugal"
annotations$country[grep("Spain|Madrid, Spain ; Tenerife, Canary Islands|Madrid, Spain|Navarra, Spain|Alt penedes, region of catalonia|spain| Lloret de mar|Barcelona| SPAIN|Salamanca|Lugo", annotations$Location)]  <- "Spain"
annotations$country[grep("France|France (Normandy, Marseille, + places visited)", annotations$Location)]  <- "France"
annotations$country[grep("Naples|italy|Roma", annotations$Location)]  <- "Italy"
annotations$country[grep("Switzerland", annotations$Location)]  <- "Switzerland"
annotations$country[grep("Belgium", annotations$Location)]  <- "Belgium"
annotations$country[grep("Germany", annotations$Location)]  <- "Germany"
annotations$country[grep("Netherlands", annotations$Location)]  <- "Netherlands"

#Following the EuroVoc classification https://en.wikipedia.org/wiki/EuroVoc + Italy and Spain and Portugal
annotations$continent[grep("Andorra|Austria|Belgium|France|Germany|Ireland|Italy|Liechtenstein|Luxembourg|Monaco|Netherlands|Portugal|Spain|Switzerland|United Kingdom", annotations$country)]  <- "Western Europe"


annotations$country[grep("Poznań, Poland", annotations$Location)]  <- "Poland"
annotations$country[grep("Estonia|Tartu, Estonia ; Rapla, Estonia|Southern Estonia|Tartu, Estonia", annotations$Location)]  <- "Estonia"
annotations$country[grep("Hungary", annotations$Location)]  <- "Hungary"
annotations$country[grep("Czech Republic", annotations$Location)]  <- "Czech Republic"
annotations$country[grep("Bucharest, Romania|Romania", annotations$Location)]  <- "Romania"
annotations$country[grep("Serbia", annotations$Location)]  <- "Serbia"
annotations$country[grep("Croatia", annotations$Location)]  <- "Croatia"
annotations$country[grep("Slovenia", annotations$Location)]  <- "Slovenia"

annotations$country[grep("Sweden", annotations$Location)]  <- "Sweden"
annotations$country[grep("Iceland", annotations$Location)]  <- "Iceland"
annotations$country[grep("Norway", annotations$Location)]  <- "Norway"
annotations$country[grep("Denmark", annotations$Location)]  <- "Denmark"

annotations$country[grep("Athens, Greece", annotations$Location)]  <- "Greece"

annotations$country[grep("Moscow, Russia", annotations$Location)]  <- "Russia"

annotations$continent[grep("Poland|Estonia|Hungary|Czech|Romania|Serbia|Croatia|Slovenia|Sweden|Iceland|Norway|Denmark|Greece|Russia", annotations$country)]  <- "Non-Western Europe"


annotations$country[grep("Turkey", annotations$Location)]  <- "Turkey"
annotations$country[grep("Iran", annotations$Location)]  <- "Iran"
annotations$country[grep("Israel|Yahud, Israel", annotations$Location)]  <- "Israel"
annotations$country[grep("Kuwait", annotations$Location)]  <- "Kuwait"

annotations$country[grep("Bombay, India", annotations$Location)]  <- "India"

annotations$country[grep("China", annotations$Location)]  <- "China"

annotations$country[grep("Singapore", annotations$Location)]  <- "Singapore"
annotations$country[grep("Indonesia", annotations$Location)]  <- "Indonesia"
annotations$country[grep("Bangkok, Thailand", annotations$Location)]  <- "Thailand"
annotations$country[grep("Osaka|Tokyo|Nagoya|osaka|Kusatsu City, Shiga Pref", annotations$Location)]  <- "Japan"
annotations$country[grep("Korea", annotations$Location)]  <- "Korea" #NOTE: one corpus just says "Korea", data collected in 2009-2011, assuming it's South Korean
annotations$country[grep("Hong-Kong", annotations$Location)]  <- "Hong Kong"
annotations$country[grep("Taiwan", annotations$Location)]  <- "Taiwan"

annotations$continent[grep("Turkey|Iran|Israel|Kuwait|India|China|Singapore|Indonesia|Thailand|Japan|Korea|Hong Kong|Taiwan", annotations$country)]  <- "Asia"


annotations$country[grep("Papua-New Guinea", annotations$Location)]  <- "Papua New Guinea"
annotations$continent[grep("Papua New Guinea", annotations$country)]  <- "Oceania"


annotations$country[grep("Alexandria, Egypt", annotations$Location)]  <- "Egypt"
annotations$country[grep("Mokhotlong, Lesotho", annotations$Location)]  <- "Lesotho"
annotations$country[grep("South Africa", annotations$Location)]  <- "South Africa"
annotations$continent[grep("Egypt|Lesotho|Africa", annotations$country)]  <- "Africa"


annotations$country[grep("Rio Cuarto, Cordoba, Argentina", annotations$Location)]  <- "Argentina"
annotations$country[grep("Sao Paulo", annotations$Location)]  <- "Brazil"
annotations$country[grep("Mexico", annotations$Location)]  <- "Mexico"
annotations$country[grep("Jamaica", annotations$Location)]  <- "Jamaica"
annotations$continent[grep("Argentina|Brazil|Mexico|Jamaica", annotations$country)]  <- "Latin America"

annotations$country[grep("Michigan, USA|USA, Northern Virginia|California, USA|washington dc|United States|USA|Washington|Maryland|San Fran", annotations$Location)]  <- "United States"
annotations$country[grep("Canada", annotations$Location)]  <- "Canada"
annotations$continent[grep("United States|Canada", annotations$country)]  <- "North America"


#Special cases
annotations$country[grep("Sweden ; Portugal", annotations$Location)]  <- "Sweden & Portugal"
annotations$country[grep("Spain (Lloret de Mar), Hungary (Kecskemét)", annotations$Location)]  <- "Spain & Hungary"
#We are leaving continent as NA (because one country is Western & the other Eastern Europe)


#to check if any left
#annotations[is.na(annotations$country),"Location"]
#annotations[is.na(annotations$continent),"Location"]





```

```{r cleaning language}
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="British English"]<-"English"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Catalan (and the variant of Spanish spoken in Barcelona)"]<-"Catalan/Spanish"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Japanese (Kyoto dialect)"]<-"Japanese"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Mandarin Chinese, influenced by the dialects of Jianghuai Mandarin and Cantonese to a minor degree"]<-"Mandarin"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Norwegian (Children: mostly trøndersk, parents/care-takers: trøndersk, vestnorsk, nordnorsk, austnorsk)"]<-"Norwegian"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Spanish Spain"]<-"Spanish"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings %in% c("European Portuguese","Brazilian Portuguese")]<-"Portuguese (Brazilian or European)"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings %in% c("Kuwaiti Arabic","Egyptian Arabic")]<-"Arabic (Egyptian or Kuwaiti)"

lang_temp=levels(factor(annotations$Language.or.Languages.spoken.in.recordings))

#split up &
lang_temp2=NULL
for(x in lang_temp) lang_temp2=c(lang_temp2,unlist(strsplit(x," & ")))
lang_temp2=levels(factor(lang_temp2))

#put mono separately from multi
mono=lang_temp2[grep("/",lang_temp2, invert=T)]
multi=lang_temp2[grep("/",lang_temp2)]

mono_langs=mono[1]
for(x in 2:length(mono)) mono_langs=paste0(mono_langs,", ",mono[x])

multi_langs=multi[1]
for(x in 2:length(multi)) multi_langs=paste0(multi_langs,", ",multi[x])

annotations$Bilingualism.Multilingualism.in.corpus<-tolower(annotations$Bilingualism.Multilingualism.in.corpus)
annotations$Bilingualism.Multilingualism.in.corpus[grep("yes",annotations$Bilingualism.Multilingualism.in.corpus)]<-"yes"
annotations$Bilingualism.Multilingualism.in.corpus[grep("trilinguism",annotations$Bilingualism.Multilingualism.in.corpus)]<-"yes"


```


```{r cleaning type}

annotations$Type.of.community.at.the.time.of.the.recordings[is.na(annotations$Type.of.community.at.the.time.of.the.recordings)]<-annotations$Naomi.community.type[is.na(annotations$Type.of.community.at.the.time.of.the.recordings)]

annotations$Type.of.community.at.the.time.of.the.recordings[annotations$Type.of.community.at.the.time.of.the.recordings %in% c("academic","capital city of the Soviet Union","city","industial","industrial ","work-for-pay",
                                                                                                                               "industrial & service &trade activities...","industrial city (3.000.000 inhabitants",
                                                                                                                               "Industrial city (population: around 272,000 inhabitants)",
                                                                                                                               "Mediterranean city of 1.6 million inhabitants",
                                                                                                                               "modern city families, usually both working parents",
                                                                                                                               "Orthodox Jews",
                                                                                                                               "Seaside tourist town",
                                                                                                                               "work-for-pay/industrial","industrial")]<-"urban"

annotations$Type.of.community.at.the.time.of.the.recordings[annotations$Type.of.community.at.the.time.of.the.recordings %in% c("farmer","rural","rural farming village")]<-"rural"

annotations$Type.of.community.at.the.time.of.the.recordings[annotations$Type.of.community.at.the.time.of.the.recordings %in% c("industrial, farmers")]<-"both"
#table(annotations$Type.of.community.at.the.time.of.the.recordings)
```



```{r continent}
byPart<- annotations %>%
  select(Language.or.Languages.spoken.in.recordings, Number.of.participants) %>%
   filter(!is.na(Number.of.participants)) %>%
     group_by(Language.or.Languages.spoken.in.recordings) %>%
       summarise(numpar = sum(as.numeric(Number.of.participants))) 

read.csv("country_iso.csv")->iso_lookup
codes=iso_lookup$Code
iso_lookup$Name<-as.character(iso_lookup$Name)
iso_lookup$Name=gsub(", Islamic Republic of","",iso_lookup$Name)
iso_lookup$Name=gsub("Russian Federation","Russia",iso_lookup$Name)
iso_lookup$Name=gsub(", Province of China","",iso_lookup$Name)
iso_lookup$Name=gsub(", Republic of","",iso_lookup$Name)
iso_lookup$Name=gsub("Congo, the Democratic Republic of the","Democratic Republic of Congo",iso_lookup$Name)

names(codes)<-iso_lookup$Name


annotations$country_WB<-codes[annotations$country]

annotations$country_WB <-factor(annotations$country_WB)
annotations[is.na(annotations$country_WB),"country"]


annotations$continent <-factor(annotations$continent)

```

```{r ancillary data: wdi}
read.csv("wdi-data.csv")->wdi_all

#remove areas
wdi_all=wdi_all[!(wdi_all$country %in% c( "Africa Eastern and Southern" ,"Africa Western and Central",
                                   "Arab World", "Caribbean small states", "Central Europe and the Baltics" ,
                                  "European Union", "Fragile and conflict affected situations",
                                  "OECD members", "West Bank and Gaza",
                                 "Small states"  , "Pacific island small states" 
                                  , "Caribbean small states" , "Other small states",
                                  "Latin America & the Caribbean (IDA & IBRD countries)", "Middle East & North Africa (IDA & IBRD countries)"  ,
 "East Asia & Pacific (IDA & IBRD countries)"     ,      "South Asia (IDA & IBRD)"    ,                        
 "Sub-Saharan Africa (IDA & IBRD countries)"    ,        "Europe & Central Asia (IDA & IBRD countries)" ,
                                  "Turks and Caicos Islands", 
                                  "Euro area"      ,                                     
 "High income"                               ,           "Heavily indebted poor countries (HIPC)"              
, "IBRD only"                                ,            "IDA total"                                           
, "IDA blend"                               ,            "IDA only"                                            
, "Latin America & Caribbean (excluding high income)",
                          "Middle East & North Africa"                          
, "IDA & IBRD total"  ,
                                  "Europe & Central Asia"  ,  "Sub-Saharan Africa (excluding high income)"  ,"Sub-Saharan Africa"    ,                               "Africa Eastern and Southern",
                                  "Least developed countries: UN classification"   ,      "Low income"                                          
, "Lower middle income"      ,                            "Low & middle income"                                 
, "Middle income"          ,                              "Middle East & North Africa (excluding high income)"  
, "Upper middle income"     ,                             "North America"                                       
, "Not classified"                                             
, "East Asia & Pacific", "Pre-demographic dividend" , "Early-demographic dividend" ,"Late-demographic dividend"       , "Post-demographic dividend" , "World","East Asia & Pacific (excluding high income)"  , "Europe & Central Asia (excluding high income)","Latin America & Caribbean","Western Sahara","World" ,"South Asia"
                                  )),]
#cleanup
wdi_all$country[wdi_all$country=="Congo, Dem. Rep."]<-"Democratic Republic of Congo"
wdi_all$country[wdi_all$country=="Congo, Rep."]<-"Congo"
wdi_all$country[wdi_all$country=="Lao PDR"]<-"Laos"
wdi_all$country[wdi_all$country=="Virgin Islands (U.S.)"]<-"United States Virgin Islands"
wdi_all$country<-gsub(", The","",wdi_all$country)
wdi_all$country<-gsub(", RB","",wdi_all$country)
wdi_all$country<-gsub(", Arab Rep.","",wdi_all$country)
wdi_all$country<-gsub(", Islamic Rep.","",wdi_all$country)
wdi_all$country<-gsub(" SAR, China","",wdi_all$country)
wdi_all$country<-gsub("St.","Saint",wdi_all$country,fixed=T)
wdi_all$country<-gsub(", Fed. Sts.","",wdi_all$country,fixed=T)
wdi_all$country<-gsub("Kyrgyz Republic","Kyrgyzstan",wdi_all$country,fixed=T)
wdi_all$country<-gsub(" Darussalam","",wdi_all$country,fixed=T)
wdi_all$country<-gsub("Cabo Verde","Cape Verde",wdi_all$country,fixed=T)
wdi_all$country<-gsub("Czech Republic","Czechia",wdi_all$country,fixed=T)
wdi_all$country<-gsub("Korea, Dem. People's Rep.","North Korea",wdi_all$country,fixed=T)
wdi_all$country<-gsub("Korea, Rep.","South Korea",wdi_all$country,fixed=T)
wdi_all$country<-gsub("Yemen, Rep.","Yemen",wdi_all$country,fixed=T)
wdi_all$country<-gsub("Syrian Arab Republic","Syria",wdi_all$country,fixed=T)
wdi_all$country<-gsub("Russian Federation","Russia",wdi_all$country,fixed=T)
wdi_all$country<-gsub("Slovak Republic","Slovakia",wdi_all$country,fixed=T)

```

```{r ancillary data: languages, eval=F}
#beginning to think this is too complicated for what we want
#https://raw.githubusercontent.com/mledoze/countries/master/dist/countries.json
jsonfile <- "https://raw.githubusercontent.com/mledoze/countries/master/dist/countries.json"
langs=fromJSON(paste(readLines(jsonfile), collapse=""))

```



```{r ancillary data: owid}
# https://ourworldindata.org/grapher/economic-inequality-gini-index
read.csv("economic-inequality-gini-index-OWID-20220215.csv")->gini
gini[gini$Year==2011,]->gini

#https://ourworldindata.org/grapher/political-regimes
read.csv("political-regimes-OWID-20220215.csv")->democr
democr[democr$Year==2011,]->democr

#https://ourworldindata.org/grapher/population-past-future
read.csv("population-past-future-OWID-20220217.csv")->pop
pop[pop$Year==2011,]->pop
colnames(pop)[colnames(pop)=="Population..historical.estimates.and.future.projections."]<-"Population"

#https://ourworldindata.org/tertiary-education
read.csv("share-of-the-population-with-completed-tertiary-education-OWID-20220217.csv")->ed
ed[ed$Year==2010,]->ed #note, no data for 2011, data only every 10 years
colnames(ed)[colnames(ed)=="Barro.Lee..Percentage.of.population.age.15..with.tertiary.schooling..Completed.Tertiary"]<-"College"




read.csv("children-born-per-woman-OWID-20220112.csv")->cpw
cpw[cpw$Year==2011,]->cpw
colnames(cpw)[colnames(cpw)=="Fertility.rate..Select.Gapminder..v12...2017."]<-"Fertility"

merge(gini,democr[,c("Entity","Political.regime")],all=T)->owid
merge(owid,pop[,c("Entity","Population")],all=T)->owid
merge(owid,ed[,c("Entity","College")],all=T)->owid
merge(owid,cpw[,c("Entity","Fertility")],all=T)->owid

#clean up
owid=owid[grep("(Urban)",owid$Entity,invert=T),]
owid=owid[grep("(Rural)",owid$Entity,invert=T),]
owid=owid[!(owid$Entity%in% c("World","Western Sahara")),]
# note: ,"Palestine/Gaza","Palestine/West Bank" -- we are leaving them in and not matching them with the World Bank's data, which has just Palestine
owid$Political.regime_t<-NA
owid$Political.regime_t[owid$Political.regime %in% c(0,1)]<-"autocracies"
#owid$Political.regime_t[owid$Political.regime==1]<-"electoral \n autocracies"
owid$Political.regime_t[owid$Political.regime%in% c(2,3)]<-"democracies"
#owid$Political.regime_t[owid$Political.regime==3]<-"liberal \n democracies"
  
  
# check no missing countries across WB & OWID merge
owid$Entity[!(owid$Entity %in% levels(factor(wdi_all$country)))]
#2 data points lost: Channel Islands & Kosovo
wdi_all$country[!(wdi_all$country %in% levels(factor(owid$Entity)))]


ind_all = merge(wdi_all,owid,by.x="country",by.y="Entity",all=T)
#ind_all$country

#https://github.com/lukes/ISO-3166-Countries-with-Regional-Codes/blob/master/all/all.csv
read.csv("ISO-3166-Countries-with-Regional-Codes.csv")->regions

ind_all=merge(ind_all,regions,by.x="iso2c",by.y="alpha.2")

ind_all$simple_region=ind_all$sub.region
ind_all$simple_region[ind_all$sub.region%in% c("Australia and New Zealand","Melanesia","Micronesia","Polynesia")]<-"Oceania"

ind_all$simple_region[grep("Asia",ind_all$sub.region)]<-"Asia"
ind_all$simple_region[grep("Africa",ind_all$sub.region)]<-"Africa"

ind_all$simple_region[ind_all$sub.region%in% c("Southern Europe","Eastern Europe","Northern Europe")]<-"Non-Western Europe"

ind_all$simple_region[ind_all$country%in% c("Spain","Italy","Portugal")]<-"Western Europe"

ind_all$log_gdp=log(ind_all$NY.GDP.PCAP.PP.KD,10)
ind_all$log2_fert=log(ind_all$Fertility,2)
ind_all$log_pop=log(ind_all$Population,10)

ind_all$western<-"non-Western"
ind_all$western[ind_all$simple_region %in% c('Western Europe',"Northern America")]<-"Western"
ind_all$western[ind_all$country %in% c('Norway',"Australia","New Zealand","Iceland","Sweden","Denmark")]<-"Western"
ind_all$western[ind_all$country %in% c('Bermuda',"Greenland")]<-"non-Western"
ind_all$western=factor(ind_all$western)
ind = ind_all[ind_all$iso2c %in% levels(annotations$country_WB),]

#final check
#levels(factor(ind_all$country))


annotations=merge(annotations,ind_all,by="country",all.x=T)

```


```{r oecd-kidsNhousehold-data}
read.csv("Households-by-number-of-children-2015-OECD-20220217.csv",sep=";")->nkids
for(i in 2:dim(nkids)[2]) nkids[,c(i)]<-as.numeric(as.character(nkids[,c(i)]))
nkids[grep("average",nkids$country,invert=T),]->nkids
nkids$anychildren<-rowSums(nkids[,c("X1.child","X2.children","X3.or.more.children")],na.rm=T)
nkids$total=nkids$X0.children+nkids$anychildren # this mostly checks out -- total is 100%
#% of single kids' household, out of households with any children
nkids$single=(nkids$X1.child/nkids$anychildren)*100
summary(nkids)

```

## Methods info

We had WDI information for `r dim(ind_all)[1]` countries, of which  `r dim(ind)[1]` were present in CHILDES. GDP data was available for `r sum(!is.na(ind_all$NY.GDP.PCAP.PP.KD))` countries overall, and `r sum(!is.na(ind$NY.GDP.PCAP.PP.KD))` countries in CHILDES. Education information was available for `r sum(!is.na(ind_all$SE.SEC.CUAT.UP.ZS))` countries overall, and `r sum(!is.na(ind$SE.SEC.CUAT.UP.ZS))` countries in CHILDES.  Information on the proportion of the population that was urban was available for `r sum(!is.na(ind_all$SP.URB.TOTL.ZS))` countries overall, and `r sum(!is.na(ind$SP.URB.TOTL.ZS))` countries in CHILDES. 

Our World in Data was our source of fertility data and type of political regime, with data available for `r sum(!is.na(cpw$Fertility))` countries overall for the selected year,  `r sum(!is.na(ind$Fertility))` of which were represented in CHILDES. We also drew from Our World in Data information about income inequality within countries via the GINI coefficient, containing `r sum(!is.na(gini$Gini.index))` countries overall for the selected year,  `r sum(!is.na(ind$Gini.index))` of which were represented in CHILDES.

## Introduction: Confounds

> As we further discuss below, some of these factors are correlated, such that it is not obvious what each separate letter contributes (see fig. 1). Others show such large within-variable variance, that they might not be a useful addition (see fig. 2).



```{r scatterplot matrix, fig.cap="Evidence that values in the multidimensional WEIRD complex are only partially correlated, with a focus on continuous variables. The diagonal shows density of the distribution of each of the variables. Panels below the diagonal show the scatter plot for the two variables involved (e.g., GDP and proportion completed highschool for the second row, first column). Those above the diagonal show the Pearson correlation for the two variables involved. Education is represented by proportion of the population completing high school; industrialization by proportion of the population living in urban (as opposed to rural) sites; richness by GDP per capita. In addition, we show an index of within-country wealth variation (the Gini coefficient: higher gini means greater wealth inequality) and women's average total fertility."}
# W E I R D
# p1 <- ggpairs(ind_all[,c('SE.SEC.CUAT.UP.ZS','SP.URB.TOTL.ZS','log_gdp','Gini.index','log2_fert')], title="Association across vars: all countries",aes(colour = ind_all$western, fill = ind_all$western, alpha = 0.4),columnLabels=c("% compl. high school","% urban","GDP (log 10)","Gini","Fert. (log 2)"), 
#                lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.7))) +
#   theme_bw()
# p1

p <- ggpairs(ind_all[,c('SE.SEC.CUAT.UP.ZS','SP.URB.TOTL.ZS','log_gdp','Gini.index','log2_fert')], title="Association across vars: all countries",columnLabels=c("% compl. high school","% urban","GDP (log 10)","Gini","Fert. (log 2)"), 
               lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.7))) +
  theme_bw()
p


```

```{r western and demo, echo=FALSE, out.height="200%", fig.cap="Evidence that values in the multidimensional WEIRD complex are only partially correlated (continued). Violin plots of continuous variables as a function of the two discrete variables: Western and type of political regime. Education is represented by proportion of the population completing high school; industrialization by percentage of the population living in urban (as opposed to rural) sites; richness by GDP per capita. In addition, we show an index of within-country wealth variation (the Gini coefficient: higher gini means greater wealth inequality) and women's average total fertility."}


ed <- ggplot(ind_all, aes(western, SE.SEC.CUAT.UP.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% compl. HS") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

urb <- ggplot(ind_all, aes(western, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% urban") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

rich <- ggplot(ind_all, aes(western, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("GDP (log 10)") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

gini <- ggplot(ind_all, aes(western, Gini.index)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("Gini.index") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

fert <- ggplot(ind_all, aes(western, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Western") +  ylab("Fert. (log 2)") 

# ggarrange( ed, urb, rich, gini, fert,  
#           labels = c("A", "B", "C", "D", "E"),
#           ncol = 1, nrow = 5)


ed_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t),], aes(Political.regime_t, SE.SEC.CUAT.UP.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

urb_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t),], aes(Political.regime_t, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

rich_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t),], aes(Political.regime_t, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)   +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

gini_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t),], aes(Political.regime_t, Gini.index)) + geom_violin() + geom_jitter(height = 0, width = 0.1)   +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

fert_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t),], aes(Political.regime_t, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Political regime") +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggarrange( ed, ed_d,  rich,rich_d,   
          urb, urb_d,  gini, gini_d, fert, fert_d,
           ncol = 2, nrow = 5)

```


```{r confounds2, echo=FALSE, out.height="200%", fig.cap="The same, but removing areas with fewer than 1 million"}
p <- ggpairs(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),c('SE.SEC.CUAT.UP.ZS','SP.URB.TOTL.ZS','log_gdp','Gini.index','log2_fert')], title="Association across vars: all countries",columnLabels=c("% compl. high school","% urban","GDP (log 10)","Gini","Fert. (log 2)"), 
               lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.7))) +
  theme_bw()
p

ed <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, SE.SEC.CUAT.UP.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% compl. HS") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

urb <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% urban") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

rich <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("GDP (log 10)") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

gini <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, Gini.index)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("Gini.index") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

fert <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Western") +  ylab("Fert. (log 2)") 

 # ggarrange( ed, urb, rich, gini, fert,  
 #           labels = c("A", "B", "C", "D", "E"),
 #           ncol = 1, nrow = 5)


ed_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t) & ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, SE.SEC.CUAT.UP.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

urb_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

rich_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)   +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

gini_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, Gini.index)) + geom_violin() + geom_jitter(height = 0, width = 0.1)   +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

fert_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Political regime") +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggarrange( ed, ed_d,  rich,rich_d,   
          urb, urb_d,  gini, gini_d, fert, fert_d,
           ncol = 2, nrow = 5)

```

## Results

### Early remarks

A total of `r nrow(all_annotations)` corpora were initially considered. We excluded  `r excl["Clinical"]` because at least some of the children were not typically developing;  `r excl["In lab/school"]` because data was collected in the lab or school; `r excl["Not available"]` because data was not available; `r excl["Only child transcribed"]` because only the child was transcribed; `r excl["Diary"]` because they were diary studies; `r excl["Task-oriented"]` because speech was triggered by a task (elicitation, story-telling, etc.); `r excl["Unfamiliar adult"]` because the conversation involved exclusively unfamiliar adults. After these exclusions, `r nrow(annotations)` remained. The following analyses will continue only on these included corpora.



### Comparison between corpora and world data in terms of education, income, urbanity and family size


The samples are varied in geographic terms, with corpora for every populated continent. Specifically, `r sum(annotations$continent=="Africa")` corpora were collected in Africa; `r sum(annotations$continent=="Asia")` in Asia; `r sum(annotations$continent=="Western Europe")` in Western Europe, and a further `r sum(annotations$continent=="Non-Western Europe")` in Non-Western Europe; `r sum(annotations$continent=="North America")` in North America and `r sum(annotations$continent=="Western Europe")` in Latin America. Only `r sum(annotations$continent=="Oceania")` was collected in Oceania.

The samples, however, over-represent richer countries, with relatively educated populations, living mostly in urban settings, and having relatively small families (Figure 3).




```{r figcomp, fig.cap="Figure 3: Density plots showing the distribution of country-level descriptors for all countries (dark gray) versus countries represented in CHILDES (orange). GDP stands for Gross Domestic Product. Percent completed high school is the percent of the country's population completing high school. Percent urban is the percent of the country's population residing in urban (as opposed to rural) locations. Total fertility rate is the average number of children a woman has over her whole reproductive period."}



#  rich
dens_dat=data.frame(rbind(cbind(ind_all$log_gdp[!is.na(ind_all$log_gdp)],0),
               cbind(ind$log_gdp[!is.na(ind$log_gdp)],1) ))
colnames(dens_dat)<-c("GDP.per.capita","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

rich <- ggplot(dens_dat, aes(x = GDP.per.capita, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "GDP per capita (log)")+ scale_fill_colorblind()


#Educated
dens_dat=data.frame(rbind(cbind(ind_all$SE.SEC.CUAT.UP.ZS[!is.na(ind_all$SE.SEC.CUAT.UP.ZS)],0),
               cbind(ind$SE.SEC.CUAT.UP.ZS[!is.na(ind$SE.SEC.CUAT.UP.ZS)],1) ))
colnames(dens_dat)<-c("completed.HS","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

ed <- ggplot(dens_dat, aes(x = completed.HS, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = c(0.25, 0.8), axis.title.y=element_blank() , legend.title = element_blank())+labs( x = "Percent completed high school")+ scale_fill_colorblind()



#urban
dens_dat=data.frame(rbind(cbind(ind_all$SP.URB.TOTL.ZS[!is.na(ind_all$SP.URB.TOTL.ZS)],0),
               cbind(ind$SP.URB.TOTL.ZS[!is.na(ind$SP.URB.TOTL.ZS)],1) ))
colnames(dens_dat)<-c("pc.urban","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

urb <- ggplot(dens_dat, aes(x = pc.urban, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() )+labs( x = "Percent urban")+ scale_fill_colorblind()

#Fertility
dens_dat=data.frame(rbind(cbind(ind_all$Fertility[!is.na(ind_all$Fertility)],0),
               cbind(ind$Fertility[!is.na(ind$Fertility)],1) ))
colnames(dens_dat)<-c("N.children.woman","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

fert <- ggplot(dens_dat, aes(x = N.children.woman, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "Total fertility rate")+ scale_fill_colorblind()

# combined

#W E I R D
ggarrange(ed, urb, rich, fert,  
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2)
```



### Description of included families

We then investigated the extent to which samples themselves captured diverse families. We started with how varied language backgrounds were. All corpora could be coded in terms of which language or languages were spoken in the recordings.  A total of `r length(lang_temp)` different languages or language combinations (for bilingual and multilingual children) were reportedly spoken in the corpora. The samples in which only one language was reported spoke `r mono_langs`, and the  samples in which multiple languages were reported spoke `r multi_langs`. That said, for `r sum(is.na(annotations$Bilingualism.Multilingualism.in.corpus))` corpora, we could not establish based on the description whether the corpora and/or the community from which the data were collected were monolingual mostly or not. Among the `r sum(!is.na(annotations$Bilingualism.Multilingualism.in.corpus))` corpora for which this could be established, `r round(sum(annotations$Bilingualism.Multilingualism.in.corpus=="yes" & !is.na(annotations$Bilingualism.Multilingualism.in.corpus)) /sum(!is.na(annotations$Bilingualism.Multilingualism.in.corpus))*100)`% were multilingual. It is hard to find reliable  estimates of the percentage of the population who are multilingual in the world (which some estimate at 57%) or the countries represented in CHILDES, but for instance, in Europe in 2016, 65% of adults reported knowing multiple languages (Eurostat, 2022). According to such estimates, even if samples are linguistically diverse, it would appear that input data in CHILDES under-represents bilinguals and multilinguals.

<!-- https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Foreign_language_skills_statistics -->

For education, there
 were `r sum(is.na(annotations$Education.Naomi))` missing values (`r round(sum(is.na(annotations$Education.Naomi))/length(annotations$Education.Naomi)*100)`%); and `r sum(annotations$Education.Naomi=="diverse",na.rm=T)` were described as diverse, without clarifying the range of education covered. Of the remaining `r sum(!is.na(annotations$Education.min))` samples, `r sum(annotations$Education.min==1 & !is.na(annotations$Education.min))` had at least some parents with primary-level education; `r sum(annotations$Education.min==2 & !is.na(annotations$Education.min))` had parents with secondary school education as the lower bound of the education range, and a further `r sum(annotations$Education.min==3 & !is.na(annotations$Education.min))` had some college as the lower bound. Thus, the majority of the samples (N = `r sum(annotations$Education.min>3 & !is.na(annotations$Education.min))`, `r round(sum(annotations$Education.min>3 & !is.na(annotations$Education.min))/sum(!is.na(annotations$Education.min))*100)`%) portrayed children whose parents had at least a graduate, if not a postgraduate, degree. In those same countries, the proportion of the population with tertiary education was `r round(mean(annotations$College, na.rm=T))`%) in 2010, suggesting that samples were not representative in terms of education.

For socioeconomic status, there
 were `r sum(is.na(annotations$SES.Naomi))` missing values (`r round(sum(is.na(annotations$SES.Naomi))/length(annotations$SES.Naomi)*100)`%). Of the remaining `r sum(!is.na(annotations$SES.Naomi))` samples, `r sum(annotations$SES.Naomi==1 & !is.na(annotations$SES.Naomi))` were described as having low SES; `r sum(annotations$SES.Naomi=="1-2" & !is.na(annotations$SES.Naomi))` were described as spanning both lower and middle or higher SES; and `r sum(annotations$SES.Naomi==2 & !is.na(annotations$SES.Naomi))` were described as middle or higher SES exclusively. Given that most countries in CHILDES are in the OECD, we can compare this proportion with the proportion of the population in these countries that are middle class. According to a 2016 report, "Almost two-thirds of people live in middle-income households in OECD countries", for whom "household net income [is] between 0.75 and 2 times the median". Thus, middle and higher class participants appear to be over-represented in CHILDES data, composing a majority of available data.
<!-- https://www.oecd.org/inclusive-growth/about/centre-for-opportunity-and-equality/Issues-note-Middle-Class-squeeze.pdf -->
 

Samples were also very homogeneous in terms of the type of community represented. Setting aside `r sum(is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` missing values (`r round(sum(is.na(annotations$Type.of.community.at.the.time.of.the.recordings))/length(annotations$Type.of.community.at.the.time.of.the.recordings)*100)`%), and focusing on the remaining `r sum(!is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` samples,  `r sum(annotations$Type.of.community.at.the.time.of.the.recordings=="urban" & !is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` (`r round(sum(annotations$Type.of.community.at.the.time.of.the.recordings=="urban" & !is.na(annotations$Type.of.community.at.the.time.of.the.recordings))/sum(!is.na(annotations$Type.of.community.at.the.time.of.the.recordings))*100)`% of the samples for which this variable was available) were described as industrialized or urban, and one additional one as both rural and urban. Only `r sum(annotations$Type.of.community.at.the.time.of.the.recordings!="urban" & !is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` samples were described as farming or rural. In those same countries, the proportion of the population residing in urban settings was `r round(mean(annotations$SP.URB.TOTL.ZS, na.rm=T))`%) in 2011, suggesting that samples were not representative of their countries in terms of rural versus urban settings either.




As for household structure, there was a great deal of missing data, with only `r  sum(!is.na(annotations$Household.structure))` (`r round(sum( !is.na(annotations$Household.structure))/length(annotations$Household.structure)*100)`%) of the samples specified. Among these, `r  sum(annotations$Household.structure=="nuclear" & !is.na(annotations$Household.structure))` (`r round(sum(annotations$Household.structure=="nuclear" & !is.na(annotations$Household.structure))/sum(!is.na(annotations$Household.structure))*100)`%) were nuclear; in `r  sum(annotations$Household.structure=="extended" & !is.na(annotations$Household.structure))` (`r round(sum(annotations$Household.structure=="extended" & !is.na(annotations$Household.structure))/sum(!is.na(annotations$Household.structure))*100)`%) were extended; and in one sample the structure was varied. We do not know of a country-level index that would allow us to check whether CHILDES corpora are representative of their countries for this variable.



Whether the children had siblings or not was an information that was missing for `r sum(is.na(annotations$Proportion.With.Siblings))` samples (`r round(sum(is.na(annotations$Proportion.With.Siblings))/length(annotations$Proportion.With.Siblings)*100)`% of included samples). For the remaining `r sum(!is.na(annotations$Proportion.With.Siblings))` samples, most contained data from multi-child families. In fact, only `r sum(annotations$Proportion.With.Siblings==0 & !is.na(annotations$Proportion.With.Siblings))` samples (`r round(sum(annotations$Proportion.With.Siblings==0 & !is.na(annotations$Proportion.With.Siblings))/sum(!is.na(annotations$Proportion.With.Siblings))*100)`% of samples having information on siblings) were constituted exclusively by children with no siblings, and the remaining had at least one sibling. As a comparison point, `r round(mean(nkids$single,na.rm=T))`% of children had no siblings in OECD countries according to 2015 data. In this sense, children with multiple siblings appear to be over-represented in CHILDES. Note that this could emerge from a bias in reporting, with corpora creators more often reporting the number of siblings when there are some than when there are none.



<!-- https://www.oecd.org/els/family/database.htm -->


```{r, eval=F}
table(annotations$Education.ac,annotations$SES.Naomi)
#,annotations$Proportion.With.Siblings==0 & !is.na(annotations$Proportion.With.Siblings)
#annotations$Type.of.community.at.the.time.of.the.recordings,annotations$Household.structure,


table(is.na(annotations$Education.ac),is.na(annotations$Household.structure))
```


