---
title: "childes_weird"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
```



```{r libraries, message=FALSE, warning=FALSE}


library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
require(scales)
library(stringr)
library(kableExtra)
library(ggpubr)
library(plotly)
library(GGally)

library(RColorBrewer)
myPalette <- brewer.pal(5, "Set2") 


group_count <-function(data_, column_){
data_<- data_ %>%
  group_by({{column_}}) %>%
    count() 
      return(data_)
}

group_select<-function(data_, column_){
data_ <- data_%>%
    group_by(!!column_) %>%
      select(Corpus, !!column_)
data_ <- data_[ order( data_[,3] ),]
return(data_)
}
```

# Load annotations 

```{r load annotations}

annotations<- read.csv("Table for authors - annotations.csv") 
colnames(annotations)[colnames(annotations)=="X..of.children.with.siblings"]<-"Nb.of.children.with.siblings"
colnames(annotations)[colnames(annotations)=="X..of.children.with.older.siblings"]<-"Nb.of.children.with.older.siblings"
colnames(annotations)=gsub("\\.\\..*","",colnames(annotations))
colnames(annotations)[colnames(annotations)=="Our.coding.of"]<-"Our.coding.of.community.type"

annotations$Number.of.participants[annotations$Number.of.participants=="1 or 5"]<-NA
annotations$Number.of.participants[annotations$Number.of.participants=="1 CHI + MOT/FAT/INV"]<-1
annotations$Number.of.participants[annotations$Number.of.participants=="4 CHI + MOTs, some FATs, some SIBs"]<-4
annotations$Number.of.participants=as.numeric(as.character(annotations$Number.of.participants))

annotations$Nb.of.children.with.siblings=as.numeric(as.character(annotations$Nb.of.children.with.siblings))

annotations$Proportion.With.Siblings=annotations$Nb.of.children.with.siblings/annotations$Number.of.participants

#table(annotations$Household.structure)
annotations$Household.structure=tolower(annotations$Household.structure)
annotations$Household.structure[grep("different|diverse|varied",annotations$Household.structure)] <- NA #turn to NA?
annotations$Household.structure[annotations$Household.structure %in% c("nuclear, extended","extended/nuclear")]<-"varied"
annotations$Household.structure[grep("nuclear|nucear|single",annotations$Household.structure)]<-"nuclear" # including single parent
annotations$Household.structure[grep("extended",annotations$Household.structure)]<-"extended"
annotations$Household.structure[annotations$Household.structure==""]<-NA


# Education & SES cleaning
annotations$Education.Naomi[annotations$Education.Naomi==""]<-NA
annotations$Education.Naomi[annotations$Education.Naomi=="4-6"]<-"4-5" #fix typo
annotations$Education.min<-as.numeric(as.character(gsub("-.*","",annotations$Education.Naomi)))
annotations$Education.max<-as.numeric(as.character(gsub(".*-","",annotations$Education.Naomi)))
annotations$Education.mid<-annotations$Education.min+(annotations$Education.max-annotations$Education.min)/2
# 1 = primary
# 2 = secondary school
# 3 = some college
# 4 = university
# 5 = postgraduate
annotations$SES.Naomi[annotations$SES.Naomi==""]<-NA
annotations$SES.min<-gsub("-.*","",annotations$SES.Naomi)
annotations$SES.max<-gsub(".*-","",annotations$SES.Naomi)
annotations$SES.max[grep("upper",annotations$Parental.socioeconomic.status)]<-3




annotations$SES.mid<-as.numeric(as.character(annotations$SES.min))+(as.numeric(as.character(annotations$SES.max))-as.numeric(as.character(annotations$SES.min)))/2


```







```{r group_by inclusion}

# include<- annotations %>%
#   mutate(inclusion=tolower(Inclusion)) %>%
#   mutate(inclusion = ifelse(str_detect(inclusion, "yes"), "yes", "no")) %>%
#   group_by(inclusion) %>%
#   count() 

# ggplot(include,aes(y=n, x=inclusion)) +
#   geom_bar(stat="identity", width=1, color="white") +
#   theme_minimal() + 
#   ggtitle("Corpus inclusion") 

all_annotations <- annotations

annotations <- all_annotations %>%
  filter(Inclusion=="yes")

table(all_annotations$why_exclude)->excl

```


A total of `r nrow(all_annotations)` corpora were initially considered. We excluded  `r excl["Clinical"]` because at least some of the children were not typically developing; `r excl["Diary"]` because they were diary studies;  `r excl["In lab/school"]` because data was collected in the lab or school; `r excl["Not available"]` because data was not available; `r excl["Only one transcribed"]` because only the child or the adult were transcribed; `r excl["Task-oriented"]` because speech was triggered by a task (elicitation, story-telling, etc.); `r excl["Unfamiliar adult"]` because the conversation involved exclusively unfamiliar adults.

After these exclusions, `r nrow(annotations)` remained. The following analyses will continue only on these included corpora.



```{r group_by location_cleaning}

annotations$country <-NA
annotations$continent <-NA


annotations$country[grep("Ireland", annotations$Location)]  <- "Ireland"
annotations$country[grep("England|Arfon area Gwynedd, North Wales|Belfast, Northern Ireland|Nottingham/Manchester, England|England, Brighton|Wales", annotations$Location)]  <- "United Kingdom"
annotations$country[grep("Portugal", annotations$Location)]  <- "Portugal"
annotations$country[grep("Spain|Madrid, Spain ; Tenerife, Canary Islands|Madrid, Spain|Navarra, Spain|Alt penedes, region of catalonia|spain| Lloret de mar|Barcelona| SPAIN|Salamanca|Lugo", annotations$Location)]  <- "Spain"
annotations$country[grep("France|France (Normandy, Marseille, + places visited)", annotations$Location)]  <- "France"
annotations$country[grep("Naples|italy|Roma", annotations$Location)]  <- "Italy"
annotations$country[grep("Switzerland", annotations$Location)]  <- "Switzerland"
annotations$country[grep("Belgium", annotations$Location)]  <- "Belgium"
annotations$country[grep("Germany", annotations$Location)]  <- "Germany"
annotations$country[grep("Netherlands", annotations$Location)]  <- "Netherlands"

#Following the EuroVoc classification https://en.wikipedia.org/wiki/EuroVoc + Italy and Spain and Portugal
annotations$continent[grep("Andorra|Austria|Belgium|France|Germany|Ireland|Italy|Liechtenstein|Luxembourg|Monaco|Netherlands|Portugal|Spain|Switzerland|United Kingdom", annotations$country)]  <- "Western Europe"


annotations$country[grep("Poznań, Poland", annotations$Location)]  <- "Poland"
annotations$country[grep("Estonia|Tartu, Estonia ; Rapla, Estonia|Southern Estonia|Tartu, Estonia", annotations$Location)]  <- "Estonia"
annotations$country[grep("Hungary", annotations$Location)]  <- "Hungary"
annotations$country[grep("Czech Republic", annotations$Location)]  <- "Czech Republic"
annotations$country[grep("Bucharest, Romania|Romania", annotations$Location)]  <- "Romania"
annotations$country[grep("Serbia", annotations$Location)]  <- "Serbia"
annotations$country[grep("Croatia", annotations$Location)]  <- "Croatia"
annotations$country[grep("Slovenia", annotations$Location)]  <- "Slovenia"

annotations$country[grep("Sweden", annotations$Location)]  <- "Sweden"
annotations$country[grep("Iceland", annotations$Location)]  <- "Iceland"
annotations$country[grep("Norway", annotations$Location)]  <- "Norway"
annotations$country[grep("Denmark", annotations$Location)]  <- "Denmark"

annotations$country[grep("Athens, Greece", annotations$Location)]  <- "Greece"

annotations$country[grep("Moscow, Russia", annotations$Location)]  <- "Russia"

annotations$continent[grep("Poland|Estonia|Hungary|Czech|Romania|Serbia|Croatia|Slovenia|Sweden|Iceland|Norway|Denmark|Greece|Russia", annotations$country)]  <- "Non-Western Europe"


annotations$country[grep("Turkey", annotations$Location)]  <- "Turkey"
annotations$country[grep("Iran", annotations$Location)]  <- "Iran"
annotations$country[grep("Israel|Yahud, Israel", annotations$Location)]  <- "Israel"
annotations$country[grep("Kuwait", annotations$Location)]  <- "Kuwait"

annotations$country[grep("Bombay, India", annotations$Location)]  <- "India"

annotations$country[grep("China", annotations$Location)]  <- "China"

annotations$country[grep("Singapore", annotations$Location)]  <- "Singapore"
annotations$country[grep("Indonesia", annotations$Location)]  <- "Indonesia"
annotations$country[grep("Bangkok, Thailand", annotations$Location)]  <- "Thailand"
annotations$country[grep("Osaka|Tokyo|Nagoya|osaka|Kusatsu City, Shiga Pref", annotations$Location)]  <- "Japan"
annotations$country[grep("Korea", annotations$Location)]  <- "Korea" #NOTE: one corpus just says "Korea", data collected in 2009-2011, assuming it's South Korean
annotations$country[grep("Hong-Kong", annotations$Location)]  <- "Hong Kong"
annotations$country[grep("Taiwan", annotations$Location)]  <- "Taiwan"

annotations$continent[grep("Turkey|Iran|Israel|Kuwait|India|China|Singapore|Indonesia|Thailand|Japan|Korea|Hong Kong|Taiwan", annotations$country)]  <- "Asia"


annotations$country[grep("Papua-New Guinea", annotations$Location)]  <- "Papua New Guinea"
annotations$continent[grep("Papua New Guinea", annotations$country)]  <- "Oceania"


annotations$country[grep("Alexandria, Egypt", annotations$Location)]  <- "Egypt"
annotations$country[grep("Mokhotlong, Lesotho", annotations$Location)]  <- "Lesotho"
annotations$country[grep("South Africa", annotations$Location)]  <- "South Africa"
annotations$continent[grep("Egypt|Lesotho|Africa", annotations$country)]  <- "Africa"


annotations$country[grep("Rio Cuarto, Cordoba, Argentina", annotations$Location)]  <- "Argentina"
annotations$country[grep("Sao Paulo", annotations$Location)]  <- "Brazil"
annotations$country[grep("Mexico", annotations$Location)]  <- "Mexico"
annotations$country[grep("Jamaica", annotations$Location)]  <- "Jamaica"
annotations$continent[grep("Argentina|Brazil|Mexico|Jamaica", annotations$country)]  <- "Latin America"

annotations$country[grep("Michigan, USA|USA, Northern Virginia|California, USA|washington dc|United States|USA|Washington|Maryland|San Fran", annotations$Location)]  <- "United States"
annotations$country[grep("Canada", annotations$Location)]  <- "Canada"
annotations$continent[grep("United States|Canada", annotations$country)]  <- "North America"


#Special cases
annotations$country[grep("Sweden ; Portugal", annotations$Location)]  <- "Sweden & Portugal"
annotations$country[grep("Spain (Lloret de Mar), Hungary (Kecskemét)", annotations$Location)]  <- "Spain & Hungary"
#We are leaving continent as NA (because one country is Western & the other Eastern Europe)


#to check if any left
#annotations[is.na(annotations$country),"Location"]
#annotations[is.na(annotations$continent),"Location"]





```

```{r cleaning language}
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Catalan (and the variant of Spanish spoken in Barcelona)"]<-"Catalan/Spanish"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Japanese (Kyoto dialect)"]<-"Japanese"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Mandarin Chinese, influenced by the dialects of Jianghuai Mandarin and Cantonese to a minor degree"]<-"Mandarin"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Norwegian (Children: mostly trøndersk, parents/care-takers: trøndersk, vestnorsk, nordnorsk, austnorsk)"]<-"Norwegian"
annotations$Language.or.Languages.spoken.in.recordings[annotations$Language.or.Languages.spoken.in.recordings=="Spanish Spain"]<-"Spanish"

lang_temp=levels(factor(annotations$Language.or.Languages.spoken.in.recordings))
langs=lang_temp[1]
for(x in 2:length(lang_temp)) langs=paste0(langs,", ",lang_temp[x])

```


```{r cleaning type}

annotations$Type.of.community.at.the.time.of.the.recordings[is.na(annotations$Type.of.community.at.the.time.of.the.recordings)]<-annotations$Naomi.community.type[is.na(annotations$Type.of.community.at.the.time.of.the.recordings)]

annotations$Type.of.community.at.the.time.of.the.recordings[annotations$Type.of.community.at.the.time.of.the.recordings %in% c("academic","capital city of the Soviet Union","city","industial","industrial ","work-for-pay",
                                                                                                                               "industrial & service &trade activities...","industrial city (3.000.000 inhabitants",
                                                                                                                               "Industrial city (population: around 272,000 inhabitants)",
                                                                                                                               "Mediterranean city of 1.6 million inhabitants",
                                                                                                                               "modern city families, usually both working parents",
                                                                                                                               "Orthodox Jews",
                                                                                                                               "Seaside tourist town",
                                                                                                                               "work-for-pay/industrial","industrial")]<-"urban"

annotations$Type.of.community.at.the.time.of.the.recordings[annotations$Type.of.community.at.the.time.of.the.recordings %in% c("farmer","rural","rural farming village")]<-"rural"

annotations$Type.of.community.at.the.time.of.the.recordings[annotations$Type.of.community.at.the.time.of.the.recordings %in% c("industrial, farmers")]<-"both"
#table(annotations$Type.of.community.at.the.time.of.the.recordings)
```



```{r ancillary data}
byPart<- annotations %>%
  select(Language.or.Languages.spoken.in.recordings, Number.of.participants) %>%
   filter(!is.na(Number.of.participants)) %>%
     group_by(Language.or.Languages.spoken.in.recordings) %>%
       summarise(numpar = sum(as.numeric(Number.of.participants))) 

read.csv("country_iso.csv")->iso_lookup
codes=iso_lookup$Code
iso_lookup$Name<-as.character(iso_lookup$Name)
iso_lookup$Name=gsub(", Islamic Republic of","",iso_lookup$Name)
iso_lookup$Name=gsub("Russian Federation","Russia",iso_lookup$Name)
iso_lookup$Name=gsub(", Province of China","",iso_lookup$Name)
iso_lookup$Name=gsub(", Republic of","",iso_lookup$Name)
iso_lookup$Name=gsub("Congo, the Democratic Republic of the","Democratic Republic of Congo",iso_lookup$Name)

names(codes)<-iso_lookup$Name


annotations$country_WB<-codes[annotations$country]

annotations$country_WB <-factor(annotations$country_WB)
annotations[is.na(annotations$country_WB),"country"]


annotations$continent <-factor(annotations$continent)


read.csv("wdi-data.csv")->dat_all

#remove areas
dat_all=dat_all[!(dat_all$country %in% c( "Africa Eastern and Southern" ,"Africa Western and Central",
                                   "Arab World", "Caribbean small states", "Central Europe and the Baltics" ,
                                  "European Union", "Fragile and conflict affected situations",
                                  "OECD members", "West Bank and Gaza",
                                 "Small states"  , "Pacific island small states" 
                                  , "Caribbean small states" , "Other small states",
                                  "Latin America & the Caribbean (IDA & IBRD countries)", "Middle East & North Africa (IDA & IBRD countries)"  ,
 "East Asia & Pacific (IDA & IBRD countries)"     ,      "South Asia (IDA & IBRD)"    ,                        
 "Sub-Saharan Africa (IDA & IBRD countries)"    ,        "Europe & Central Asia (IDA & IBRD countries)" ,
                                  "Turks and Caicos Islands", 
                                  "Euro area"      ,                                     
 "High income"                               ,           "Heavily indebted poor countries (HIPC)"              
, "IBRD only"                                ,            "IDA total"                                           
, "IDA blend"                               ,            "IDA only"                                            
, "Latin America & Caribbean (excluding high income)",
                          "Middle East & North Africa"                          
, "IDA & IBRD total"  ,
                                  "Europe & Central Asia"  ,  "Sub-Saharan Africa (excluding high income)"  ,"Sub-Saharan Africa"    ,                               "Africa Eastern and Southern",
                                  "Least developed countries: UN classification"   ,      "Low income"                                          
, "Lower middle income"      ,                            "Low & middle income"                                 
, "Middle income"          ,                              "Middle East & North Africa (excluding high income)"  
, "Upper middle income"     ,                             "North America"                                       
, "Not classified"                                             
, "East Asia & Pacific", "Pre-demographic dividend" , "Early-demographic dividend" ,"Late-demographic dividend"       , "Post-demographic dividend" , "World","East Asia & Pacific (excluding high income)"  , "Europe & Central Asia (excluding high income)","Latin America & Caribbean","Western Sahara","World" ,"South Asia"
                                  )),]
#cleanup
dat_all$country[dat_all$country=="Congo, Dem. Rep."]<-"Democratic Republic of Congo"
dat_all$country[dat_all$country=="Congo, Rep."]<-"Congo"
dat_all$country[dat_all$country=="Lao PDR"]<-"Laos"
dat_all$country[dat_all$country=="Virgin Islands (U.S.)"]<-"United States Virgin Islands"
dat_all$country<-gsub(", The","",dat_all$country)
dat_all$country<-gsub(", RB","",dat_all$country)
dat_all$country<-gsub(", Arab Rep.","",dat_all$country)
dat_all$country<-gsub(", Islamic Rep.","",dat_all$country)
dat_all$country<-gsub(" SAR, China","",dat_all$country)
dat_all$country<-gsub("St.","Saint",dat_all$country,fixed=T)
dat_all$country<-gsub(", Fed. Sts.","",dat_all$country,fixed=T)
dat_all$country<-gsub("Kyrgyz Republic","Kyrgyzstan",dat_all$country,fixed=T)

read.csv("children-born-per-woman-OWID-20220112.csv")->cpw

cpw[cpw$Year==2011,]->rcpw_all
# rcpw_all <- cpw %>% #old version, I thought data may be missing one particular year for a given country, but it's not the case
#              group_by(Entity) %>%
#              filter(Year == max(Year)) 
colnames(rcpw_all)[colnames(rcpw_all)=="Fertility.rate..Select.Gapminder..v12...2017."]<-"fertility"


ind_all = merge(dat_all,rcpw_all,by.x="country",by.y="Entity",all=T)
#ind_all$country

ind = ind_all[ind_all$iso2c %in% levels(annotations$country_WB),]


annotations=merge(annotations,ind_all,by="country",all.x=T)

# table(!is.na(annotations$SES.mid),!is.na(annotations$NY.GDP.PCAP.PP.KD))
# table(!is.na(annotations$Education.mid),!is.na(annotations$SE.SEC.CUAT.UP.ZS))
# table(!is.na(annotations$Type.of.community.at.the.time.of.the.recordings),!is.na(annotations$SP.URB.TOTL.ZS))
#             
# table(!is.na(annotations$Average.number.of.siblings),!is.na(annotations$SP.DYN.CEBN.Q3))



#annotations=merge(annotations,rcpw_all,by.x="country",by.y="Entity",all.x=T)

#rcpw=rcpw_all[rcpw_all$Entity %in% levels(factor(annotations$country)),]

#TODO check no missing countries across WB & OWID merge
```

We had WDI information for `r dim(ind_all)[1]` countries, of which  `r dim(ind)[1]` were present in CHILDES. GDP data was available for `r sum(!is.na(ind_all$NY.GDP.PCAP.PP.KD))` countries overall, and `r sum(!is.na(ind$NY.GDP.PCAP.PP.KD))` countries in CHILDES. Education information was available for `r sum(!is.na(ind_all$SE.SEC.CUAT.UP.ZS))` countries overall, and `r sum(!is.na(ind$SE.SEC.CUAT.UP.ZS))` countries in CHILDES.  Information on the proportion of the population that was urban was available for `r sum(!is.na(ind_all$SP.URB.TOTL.ZS))` countries overall, and `r sum(!is.na(ind$SP.URB.TOTL.ZS))` countries in CHILDES. 

Fertility rate was obtained from Our World in Data, which had `r sum(!is.na(rcpw_all$fertility))` countries overall for the selected year, and `r sum(!is.na(ind$fertility))` of them were represented in CHILDES. 


## At the level of the population

The samples are varied in geographic terms, with corpora for every peopled continent. Specifically, `r sum(annotations$continent=="Africa")` corpora were collected in Africa; `r sum(annotations$continent=="Asia")` in Asia; `r sum(annotations$continent=="Western Europe")` in Western Europe, and a further `r sum(annotations$continent=="Non-Western Europe")` in Non-Western Europe; `r sum(annotations$continent=="North America")` in North America and `r sum(annotations$continent=="Western Europe")` in Latin America. Only `r sum(annotations$continent=="Oceania")` was collected in Oceania.

The samples, however, over-represent richer countries, with relatively educated populations, living mostly in urban settings, and having relatively small families (Figure XX).




```{r rich}

dens_dat=data.frame(rbind(cbind(ind_all$NY.GDP.PCAP.PP.KD[!is.na(ind_all$NY.GDP.PCAP.PP.KD)],0),
               cbind(ind$NY.GDP.PCAP.PP.KD[!is.na(ind$NY.GDP.PCAP.PP.KD)],1) ))
colnames(dens_dat)<-c("GDP.per.capita","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

rich <- ggplot(dens_dat, aes(x = GDP.per.capita, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "GDP per capita")

```



```{r Educated}
dens_dat=data.frame(rbind(cbind(ind_all$SE.SEC.CUAT.UP.ZS[!is.na(ind_all$SE.SEC.CUAT.UP.ZS)],0),
               cbind(ind$SE.SEC.CUAT.UP.ZS[!is.na(ind$SE.SEC.CUAT.UP.ZS)],1) ))
colnames(dens_dat)<-c("completed.HS","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

ed <- ggplot(dens_dat, aes(x = completed.HS, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() )+labs( x = "Percent completed high school")


```




```{r urban}
dens_dat=data.frame(rbind(cbind(ind_all$SP.URB.TOTL.ZS[!is.na(ind_all$SP.URB.TOTL.ZS)],0),
               cbind(ind$SP.URB.TOTL.ZS[!is.na(ind$SP.URB.TOTL.ZS)],1) ))
colnames(dens_dat)<-c("pc.urban","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

urb <- ggplot(dens_dat, aes(x = pc.urban, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() )+labs( x = "Percent urban")

```




```{r fertility}

dens_dat=data.frame(rbind(cbind(ind_all$fertility[!is.na(ind_all$fertility)],0),
               cbind(ind$fertility[!is.na(ind$fertility)],1) ))
colnames(dens_dat)<-c("N.children.woman","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

fert <- ggplot(dens_dat, aes(x = N.children.woman, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "Total fertility rate")

```

```{r combined}

ggarrange(rich, ed, urb, fert,  
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2)
```


## At the level of the family




A total of `r length(lang_temp)` different languages or language combinations (for bilingual and multilingual children) were reportedly spoken in the corpora. The precise list is (multiple languages spoken within the family are separated by "/", "&" indicates two participant subsamples in terms of languages spoken): `r langs`.

### Educated


We then investigated the extent to which samples themselves captured families with diverse educational backgrounds. There were `r sum(is.na(annotations$Education.Naomi))` missing values (`r round(sum(is.na(annotations$Education.Naomi))/length(annotations$Education.Naomi)*100)`%); and `r sum(annotations$Education.Naomi=="diverse")` were described as diverse, without clarifying the range of education covered. Of the remaining `r sum(!is.na(annotations$Education.min))` samples, `r sum(annotations$Education.min==1 & !is.na(annotations$Education.min))` had at least some parents with primary-level education; `r sum(annotations$Education.min==2 & !is.na(annotations$Education.min))` had parents with secondary school education as the lower bound, and a further `r sum(annotations$Education.min==3 & !is.na(annotations$Education.min))` had some college as the lower bound. Thus, the majority of the samples (N = `r sum(annotations$Education.min>3 & !is.na(annotations$Education.min))`, `r round(sum(annotations$Education.min>3 & !is.na(annotations$Education.min))/sum(!is.na(annotations$Education.min))*100)`%) portrayed children whose parents had a graduate or a post-graduate degree.

### Urban versus rural

Samples were also very homogeneous in terms of the type of community represented: `r sum(annotations$Type.of.community.at.the.time.of.the.recordings=="urban" & !is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` (`r round(sum(annotations$Type.of.community.at.the.time.of.the.recordings=="urban" & !is.na(annotations$Type.of.community.at.the.time.of.the.recordings))/sum(!is.na(annotations$Type.of.community.at.the.time.of.the.recordings))*100)`% of the samples for which this variable was not missing) were described as industrialized or urban, and one additional one as both rural and urban. Only `r sum(annotations$Type.of.community.at.the.time.of.the.recordings!="urban" & !is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` samples were described as farming or rural.


### Household structure

As for household structure, there was a great deal of missing data, with only `r  sum(!is.na(annotations$Household.structure))` (`r round(sum( !is.na(annotations$Household.structure))/length(annotations$Household.structure)*100)`%) of the samples specified. Among these, `r  sum(annotations$Household.structure=="nuclear" & !is.na(annotations$Household.structure))` (`r round(sum(annotations$Household.structure=="nuclear" & !is.na(annotations$Household.structure))/sum(!is.na(annotations$Household.structure))*100)`%) were nuclear; in `r  sum(annotations$Household.structure=="extended" & !is.na(annotations$Household.structure))` (`r round(sum(annotations$Household.structure=="extended" & !is.na(annotations$Household.structure))/sum(!is.na(annotations$Household.structure))*100)`%) were extended; and in one sample structure varied. 

### Siblings

It was not the case that the majority of children were single children. In fact, only `r sum(annotations$Proportion.With.Siblings==0 & !is.na(annotations$Proportion.With.Siblings))` samples (`r round(sum(annotations$Proportion.With.Siblings==0 & !is.na(annotations$Proportion.With.Siblings))/sum(!is.na(annotations$Proportion.With.Siblings))*100)`% of samples having information on siblings) were constituted exclusively by children with no siblings, and the remaining had at least some siblings. However, it should be born in mind that this variable was missing for `r sum(is.na(annotations$Proportion.With.Siblings))` samples (`r round(sum(is.na(annotations$Proportion.With.Siblings))/length(annotations$Proportion.With.Siblings)*100)`% of included samples).

## Confounds

Above, we suggested that our samples are biased in several ways -- but perhaps it is hard to separate these biases. That is, some readers may have the intuition that Western, Educated, Industrialized, Rich, and Infertile are correlated in the world, and thus it will be hard to fix one without fixing the others and vice versa. One way to address this concern which has been suggested in other work is to study within-country variation. But we also want to know -- are these aspects confounded in terms of between-country variation?

```{r}
p <- ggpairs(ind_all[,c('NY.GDP.PCAP.PP.KD','SE.SEC.CUAT.UP.ZS','SP.URB.TOTL.ZS','fertility')], title="Association across vars: all countries") 
p

```


```{r}
p <- ggpairs(ind[,c('NY.GDP.PCAP.PP.KD','SE.SEC.CUAT.UP.ZS','SP.URB.TOTL.ZS','fertility')], title="Association across vars; in CHILDES") 
p

```

- is WEIRD confounded in the world:
		- density plots (Western non Western): education, urban, rich, family size
		- correlation matrix: W/nW, education, urban, rich, family size
		- table
	- is WEIRD confounded in CHILDES
		- table: W/nW, urban,  education, has sibs (N of corpora)