---
title: "How WEIRD-biased is CHILDES data on childrens linguistic input? Supplementary Materials 1"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Supplementary recommendations for CHILDES

As represented in  Figure 1, CHILDES corpora can have two levels of metadata, one at the speaker level and the other at the level of the corpus as a whole. At present, nearly none of the fields are mandatory and seldom filled in. The standardized (yet optional) fields in speaker-level metadata include several fields that would be relevant to the study of diversity in CHILDES, including: language, age, sex, socioeconomic status (WC for working class, UC for upper class, MC for middle class, LI for limited income), role (relationship to the child), and education (in years). At the corpus level, there are already variables to encode the type of activity recorded (interactivity, planning type, involvement, social context, event structure, and task) as well as information about location (at the continent and country levels). There is, at present, no way to signal the geographic origin of CHILDES data more finely, but if this information were added, then we could assess the extent to which localities where education is widespread are over-represented; and the extent to which unusually well-educated parents are over-represented. Perhaps a reasonable goal is to attempt to match the world's distribution in terms of simple educational metrics, like having completed lower secondary school.
 

![Metadata found in CHILDES corpora. The left panel shows a selection of speaker-level metadata fields (reproduced from p. 35 of the CHAT Manual, MacWhinney, 2000). The variable Group can contain any grouping relevant to the corpus producers. The right panel shows corpus-level metadata for Sarvasy (2017).](metadata.jpg)


Incorporating these other metrics would require not only adding fields to the CHILDES' metadata system, but also developing training materials to ensure that all CHILDES contributors provide information in a standardized way. The extra effort required to implement this could be measured with a small number of volunteers before making it a requirement for data deposit, to both check for feasibility and ensure that the benefits of considering and tracking this information outweigh the costs.



## Further acknowledgments

```{r, echo=F}
read.csv("CHILDES corpora contacts - Foglio1.csv")->x


y=NULL
for(i in 1:length(x$name)) y=c(y,unlist(strsplit(as.character(x$name[i]),"    ")))
y=gsub("^ *","",y)
y[y!=""]->y
y[y!="404 Not Found"]->y
z=unique(sort(y))

q=z[1]
for(i in 2:(length(z)-1)) q=paste0(q,", ",z[i])
q=paste0(q,", and ",z[i])
```

We would like to thank the curators of the corpora who replied to our email: `r q`.

## Additional analyses

The following figures on education use WDI's proportion of the population completing high school (rather than Our world in data's proportion of the population completing lower secondary school).

```{r libraries, message=FALSE, warning=FALSE, echo=F}


library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
require(scales)
library(stringr)
library(kableExtra)
library(ggpubr)
library(plotly)
library(GGally)
library(ggthemes)
library(rjson)


library(RColorBrewer)
myPalette <- brewer.pal(5, "Set2") 


group_count <-function(data_, column_){
data_<- data_ %>%
  group_by({{column_}}) %>%
    count() 
      return(data_)
}

group_select<-function(data_, column_){
data_ <- data_%>%
    group_by(!!column_) %>%
      select(Corpus, !!column_)
data_ <- data_[ order( data_[,3] ),]
return(data_)
}
```

```{r data-processing, echo=F,message=F, warning=FALSE}
source("data-processing.R")
```



```{r confounds2b, echo=FALSE, out.height="200%", fig.cap="Equivalent to Figure 1 in the main manuscript, only the education variable has changed.", message=FALSE, warning=FALSE, echo=F}


ed <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, SE.SEC.CUAT.UP.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% compl. HS") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

urb <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% urban") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

rich <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("GDP (log 10)") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())


fert <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Western") +  ylab("Fert. (log 2)") 




ed_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t) & ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, SE.SEC.CUAT.UP.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

urb_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

rich_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)   +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


fert_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Political regime") +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggarrange( ed, ed_d,  rich,rich_d,   
          urb, urb_d,   fert, fert_d,
           ncol = 2, nrow = 4)

```



```{r scatterplot matrix2, fig.cap="Equivalent to Figure 2  in the main manuscript, only the education variable has changed.", message=FALSE, warning=FALSE, echo=F}

p <- ggpairs(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),c('SE.SEC.CUAT.UP.ZS','SP.URB.TOTL.ZS','log_gdp','log2_fert')], title="Association across vars: all countries",columnLabels=c("% compl. high school","% urban","GDP (log 10)","Fert. (log 2)"), 
               lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.7))) +
  theme_bw()
p


```

```{r figcomp2, fig.cap="Equivalent to Figure 3 in the main manuscript, only the education variable has changed.", message=FALSE, warning=FALSE, echo=F}

#  rich
dens_dat=data.frame(rbind(cbind(ind_all$log_gdp[!is.na(ind_all$log_gdp)],0),
               cbind(ind$log_gdp[!is.na(ind$log_gdp)],1) ))
colnames(dens_dat)<-c("GDP.per.capita","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

rich <- ggplot(dens_dat, aes(x = GDP.per.capita, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "GDP per capita (log)")+ scale_fill_colorblind()


#Educated
dens_dat=data.frame(rbind(cbind(ind_all$SE.SEC.CUAT.UP.ZS[!is.na(ind_all$SE.SEC.CUAT.UP.ZS)],0),
               cbind(ind$SE.SEC.CUAT.UP.ZS[!is.na(ind$SE.SEC.CUAT.UP.ZS)],1) ))
colnames(dens_dat)<-c("completed.HS","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

ed <- ggplot(dens_dat, aes(x = completed.HS, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = c(0.25, 0.8), axis.title.y=element_blank() , legend.title = element_blank())+labs( x = "Percent completed high school")+ scale_fill_colorblind()



#urban
dens_dat=data.frame(rbind(cbind(ind_all$SP.URB.TOTL.ZS[!is.na(ind_all$SP.URB.TOTL.ZS)],0),
               cbind(ind$SP.URB.TOTL.ZS[!is.na(ind$SP.URB.TOTL.ZS)],1) ))
colnames(dens_dat)<-c("pc.urban","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

urb <- ggplot(dens_dat, aes(x = pc.urban, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() )+labs( x = "Percent urban")+ scale_fill_colorblind()

#Fertility
dens_dat=data.frame(rbind(cbind(ind_all$Fertility[!is.na(ind_all$Fertility)],0),
               cbind(ind$Fertility[!is.na(ind$Fertility)],1) ))
colnames(dens_dat)<-c("N.children.woman","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")

fert <- ggplot(dens_dat, aes(x = N.children.woman, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "Total fertility rate")+ scale_fill_colorblind()

# combined

#W E I R D
ggarrange(ed, urb, rich, fert,  
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2)
```


# Archive

### Code to reproduce information in the caption to Figure 1

We classified countries as Western and Democratic. For Western, we classified as Western all of the countries in Western Europe (according to EuroVoc: Andorra, Austria, Belgium, France, Germany, Ireland, Liechtenstein, Luxembourg, Monaco, Netherlands, Switzerland, United Kingdom) as well as a few others in Europe (Italy, Portugal, Spain, Norway, Sweden, Iceland, Denmark), most of North America (with the exception of Bermuda and Greenland), and two countries in Oceania (Australia, New Zealand). This resulted in `r sum(ind_all$western == "Western")` Western and `r sum(ind_all$western == "non-Western")` non-Western countries. For democratic, we simplified the OWID's classification of closed and electoral autocracies into autocracies, versus electoral and liberal democracies into democracies (Our World in data, 2022c). 
We had data on political regime for `r sum(!is.na(ind_all$Political.regime_t))` countries (from Our World in Data, OWID, 2011); on the proportion of the population having completed lower secondary school `r sum(!is.na(ind_all$Compl.LS))` (from OWID, 2007-2014); on GDP  for `r sum(!is.na(ind_all$NY.GDP.PCAP.PP.KD))` countries (from WDI, 2011); on the proportion of the population that was urban for `r sum(!is.na(ind_all$SP.URB.TOTL.ZS))` countries (from WDI, 2011); on fertility for `r sum(!is.na(ind_all$Fertility))` countries (from OWID, 2011). 


In this document, we provide code that allows readers to reproduce information provided in the main text that depend on calculations computed here.

## Introduction



### Code to reproduce Figure 1

Note that this Figure appears in a different page.

```{r confounds2, echo=FALSE, out.height="200%", fig.cap="Evidence that values in the multidimensional WEIRD complex are only partially correlated (continued). Violin plots of continuous variables as a function of the two discrete variables: Western and type of political regime. Education is represented by proportion of the population completing lower secondary school; industrialization by percentage of the population living in urban (as opposed to rural) sites; richness by GDP per capita. In addition, we show women's average total fertility.", echo=F,message=F, warning=F, results='hide'}


ed <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, Compl.LS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% compl. LSS") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

urb <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% urban") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

rich <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("GDP (log 10)") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())


fert <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Western") +  ylab("Fert. (log 2)") 




ed_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t) & ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, Compl.LS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

urb_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

rich_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)   +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


fert_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Political regime") +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

x = ggarrange( ed, ed_d,  rich,rich_d,   
          urb, urb_d,   fert, fert_d,
           ncol = 2, nrow = 4)
x
pdf("Fig1.pdf")
x
dev.off()
```


### Code to reproduce information in the caption to Figure 1

We classified countries as Western and Democratic. For Western, we classified as Western all of the countries in Western Europe (according to EuroVoc: Andorra, Austria, Belgium, France, Germany, Ireland, Liechtenstein, Luxembourg, Monaco, Netherlands, Switzerland, United Kingdom) as well as a few others in Europe (Italy, Portugal, Spain, Norway, Sweden, Iceland, Denmark), most of North America (with the exception of Bermuda and Greenland), and two countries in Oceania (Australia, New Zealand). This resulted in `r sum(ind_all$western == "Western")` Western and `r sum(ind_all$western == "non-Western")` non-Western countries. For democratic, we simplified the OWID's classification of closed and electoral autocracies into autocracies, versus electoral and liberal democracies into democracies (Our World in data, 2022c). 
We had data on political regime for `r sum(!is.na(ind_all$Political.regime_t))` countries (from Our World in Data, OWID, 2011); on the proportion of the population having completed lower secondary school `r sum(!is.na(ind_all$Compl.LS))` (from OWID, 2007-2014); on GDP  for `r sum(!is.na(ind_all$NY.GDP.PCAP.PP.KD))` countries (from WDI, 2011); on the proportion of the population that was urban for `r sum(!is.na(ind_all$SP.URB.TOTL.ZS))` countries (from WDI, 2011); on fertility for `r sum(!is.na(ind_all$Fertility))` countries (from OWID, 2011). 


### Code to reproduce text on p. 13

The answer to the question of who surrounds the child may also depend on how large the nuclear family is. There is considerable cross-population variation in terms of the average number of children a woman has in her reproductive lifetime, varying between a little over 1 (`r paste0(ind_all[which(ind_all$Fertility==min(ind_all$Fertility,na.rm=T)),"Fertility"],", ",ind_all[which(ind_all$Fertility==min(ind_all$Fertility,na.rm=T)),"country"])`) to over 6 (`r paste0(ind_all[which(ind_all$Fertility==max(ind_all$Fertility,na.rm=T)),"Fertility"],", ",ind_all[which(ind_all$Fertility==max(ind_all$Fertility,na.rm=T)),"country"])`). 

```{r}
#https://submissions2.mirasmart.com/Verify/PAA2021/Submission/Temp/radd1gjhtkk.pdf
#data extracted from fig 6
x<-read.csv("arg2010pcwch.txt",sep="\t")
x$tfr<-(x$X1+x$X2*2+x$X3*3+x$X4*4)/100
```



## Methods

### Code to reproduce text on p. 16

A total of `r nrow(annotations)` corpora were initially considered. We excluded  `r excluded["Clinical"]` because at least some of the children were not typically developing;  `r excl["In lab/school"]` because data was collected in the lab or school; `r excl["Not available"]` because data was not available; `r excl["Only child transcribed"]` because only the child was transcribed; `r excl["Diary"]` because they were diary studies; `r excl["Task-oriented"]` because speech was triggered by a task (elicitation, story-telling, etc.); `r excl["Unfamiliar adult"]` because the conversation involved exclusively unfamiliar adults. After these exclusions, `r nrow(annotations)` remained. The following analyses will continue only on these included corpora.



### Code to reproduce information on Table 2

```{r tab2, eval=F}

colSums(!is.na(annotations[,c("country","Compl.LS","SP.URB.TOTL.ZS","log_gdp","log2_fert")]))


```

### Code to reproduce information on Table 3

```{r tab3, eval=F}
colSums(!is.na(annotations[,c("Education.STDZD","SES.STDZD","Parental.profession","Language.or.Languages.spoken.in.recordings","Bilingualism.Multilingualism.in.corpus",
                    "Type.of.community.at.the.time.of.the.recordings", "Household.structure","Proportion.With.Siblings","Average.number.of.siblings")]))

```



## Results

### Code to reproduce text on p. 20


The samples are varied in geographic terms, with corpora for every populated continent. Specifically, `r sum(annotations$continent=="Africa")` corpora were collected in Africa; `r sum(annotations$continent=="Asia")` in Asia; `r sum(annotations$continent=="Western Europe")` in Western Europe, and a further `r sum(annotations$continent=="Non-Western Europe")` in Non-Western Europe; `r sum(annotations$continent=="North America")` in North America and `r sum(annotations$continent=="Western Europe")` in Latin America. Only `r sum(annotations$continent=="Oceania")` was collected in Oceania.

### Code to reproduce Figure 3
Note that this Figure appears in a different page.

```{r figcomp, fig.cap="Figure 3: Density plots showing the distribution of country-level descriptors for all countries (dark gray) versus countries represented in CHILDES (orange). GDP stands for Gross Domestic Product. Percent completed lower secondary school is the percent of the country's population completing lower secondary school. Percent urban is the percent of the country's population residing in urban (as opposed to rural) locations. Total fertility rate is the average number of children a woman has over her whole reproductive period.", results='hide'}


#  rich
dens_dat=data.frame(rbind(cbind(ind_all$log_gdp[!is.na(ind_all$log_gdp)],0),
               cbind(ind$log_gdp[!is.na(ind$log_gdp)],1) ))
colnames(dens_dat)<-c("GDP.per.capita","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
gdpt = t.test(dens_dat[,1]~dens_dat$in.childes)

rich <- ggplot(dens_dat, aes(x = GDP.per.capita, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "GDP per capita (log 10)")+ scale_fill_colorblind()


#Educated
dens_dat=data.frame(rbind(cbind(ind_all$Compl.LS[!is.na(ind_all$Compl.LS)],0),
               cbind(ind$Compl.LS[!is.na(ind$Compl.LS)],1) ))
colnames(dens_dat)<-c("completed.LSS","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
edt = t.test(dens_dat[,1]~dens_dat$in.childes)

ed <- ggplot(dens_dat, aes(x = completed.LSS, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = c(0.25, 0.8), axis.title.y=element_blank() , legend.title = element_blank())+labs( x = "% compl. LSS")+ scale_fill_colorblind()



#urban
dens_dat=data.frame(rbind(cbind(ind_all$SP.URB.TOTL.ZS[!is.na(ind_all$SP.URB.TOTL.ZS)],0),
               cbind(ind$SP.URB.TOTL.ZS[!is.na(ind$SP.URB.TOTL.ZS)],1) ))
colnames(dens_dat)<-c("pc.urban","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
urbt = t.test(dens_dat[,1]~dens_dat$in.childes)

urb <- ggplot(dens_dat, aes(x = pc.urban, fill = in.childes)) + geom_density(alpha = 0.5) + theme(legend.position = "none", axis.title.y=element_blank() )+labs( x = "% urban")+ scale_fill_colorblind()

#Fertility
dens_dat=data.frame(rbind(cbind(ind_all$log2_fert[!is.na(ind_all$log2_fert)],0),
               cbind(ind$log2_fert[!is.na(ind$log2_fert)],1) ))
colnames(dens_dat)<-c("N.children.woman","in.childes_T")
dens_dat$in.childes=ifelse(dens_dat$in.childes,"CHILDES","all")
fertt = t.test(dens_dat[,1]~dens_dat$in.childes)

fert <- ggplot(dens_dat, aes(x = N.children.woman, fill = in.childes)) + geom_density(alpha = 0.5)+ theme(legend.position = "none", axis.title.y=element_blank() ) +labs( x = "Fertility (log 2)")+ scale_fill_colorblind()

# combined

#W E I R D
x <- ggarrange(ed, urb, rich, fert,  
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2)
x
pdf("Fig3.pdf")
x
dev.off()
```


### Code to reproduce text on p. 20

Using country-level statistics, we were able to assess the extent to which the countries with data in CHILDES were a representative sample of countries in the world. Density plots are portrayed in Figure 3. The means for each of the variables differed for the countries in CHILDES versus overall in the world using unpaired samples t-tests without assuming equality of variance (Welch's t). Countries in CHILDES had a higher proportion of the population completing lower secondary school than the world wide sample (% compl. LSS, t(`r round(edt$parameter,2)`)=`r round(edt$statistic,2)`, p = `r round(edt$p.value,2)`); they were more urban (% urban, t(`r round(urbt$parameter,2)`)=`r round(urbt$statistic,2)`, p = `r round(urbt$p.value,2)`);  richer (log GDP per capita, t(`r round(gdpt$parameter,2)`)=`r round(gdpt$statistic,2)`, p = `r round(gdpt$p.value,2)`) and had lower fertility rates (log fertility, t(`r round(fertt$parameter,2)`)=`r round(fertt$statistic,2)`, p = `r round(fertt$p.value,2)`).

### Code to reproduce text on p. 22



Education information was missing for more than half of the corpora (see Table 3); and `r sum(annotations$Education.STDZD=="diverse",na.rm=T)` were described as diverse, without clarifying the range of education covered. Of the remaining `r sum(!is.na(annotations$Education.min))` samples, `r sum(annotations$Education.min==1 & !is.na(annotations$Education.min))` had at least some parents with primary-level education;  `r sum(annotations$Education.min==2 & !is.na(annotations$Education.min))` had parents with secondary school education as the lower bound of the education range, and a further `r sum(annotations$Education.min==3 & !is.na(annotations$Education.min))` had some college as the lower bound. Thus, `r round(sum(annotations$Education.min>3 & !is.na(annotations$Education.min))/sum(!is.na(annotations$Education.min))*100)`% of the samples (N = `r sum(annotations$Education.min>3 & !is.na(annotations$Education.min))`) portrayed children whose parents had at least a graduate, if not a postgraduate, degree. Samples were not representative of the countries they were collected, since in those same countries the proportion of the population with tertiary education was only `r round(mean(annotations$College, na.rm=T))`%.



For socioeconomic status, there were `r sum(is.na(annotations$SES.STDZD))` missing values (`r round(sum(is.na(annotations$SES.STDZD))/length(annotations$SES.STDZD)*100)`%). Of the remaining `r sum(!is.na(annotations$SES.STDZD))` samples, `r sum(annotations$SES.STDZD==1 & !is.na(annotations$SES.STDZD))` were described as having low SES; `r sum(annotations$SES.STDZD=="1-2" & !is.na(annotations$SES.STDZD))` were described as spanning both lower and middle or higher SES; and `r sum(annotations$SES.STDZD==2 & !is.na(annotations$SES.STDZD))` were described as middle or higher SES exclusively. Given that most countries represented in CHILDES are in the Organization for Economic Cooperation and Development (OECD, `r sum(annotations$oecd)` out of the `r dim(annotations)[1]` corpora), we can compare this proportion with the proportion of the population in these countries that are middle class. According to a 2016 report, “Almost two-thirds of people live in middle-income households in OECD countries”, for whom “household net income [is] between 0.75 and 2 times the median”. Thus, middle and higher class participants appear to be over-represented in CHILDES data, composing 81% of available data.

<!-- https://www.oecd.org/inclusive-growth/about/centre-for-opportunity-and-equality/Issues-note-Middle-Class-squeeze.pdf -->

Information about parents’ profession or activity was also missing for the majority of the corpora (see Table 3). Professions were overall varied, but it should be noted that `r round(sum(annotations$is.academic=="yes" & !is.na(annotations$is.academic))/sum(!is.na(annotations$Parental.profession))*100)`% of the samples contained parents who were described as (Masters or PhD) students, professors, linguists, researchers, scientists, or academics. To give an idea of the extent to which this is not representative, consider the fact that in 2020, `r round((47000 +135000+20000000+3000000)/392000000*100)`% of the American population would be included in that list of professions. Similar data is hard to find for all countries represented in CHILDES, but we suspect that the proportion of scientists, professors, Masters and PhD students found in most other countries will be the same or lower.

<!-- in 2020, 47,000 people were scientists in the USA (including both in academia and the private sector), 135,000 were professors, 20 million were Masters students, and 3 million were PhD students. Given that the USA’s population is 392 million, -->


Only about a third of the samples had information about whether the community was rural or urban (see Table 3), and the remaining ones were very homogeneous. Setting aside  `r sum(is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` missing values (`r round(sum(is.na(annotations$Type.of.community.at.the.time.of.the.recordings))/length(annotations$Type.of.community.at.the.time.of.the.recordings)*100)`%), and focusing on the remaining `r sum(!is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` samples, `r sum(annotations$Type.of.community.at.the.time.of.the.recordings=="urban" & !is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` were described as industrialized or urban, and an additional one as both rural and urban. Only `r sum(annotations$Type.of.community.at.the.time.of.the.recordings!="urban" & !is.na(annotations$Type.of.community.at.the.time.of.the.recordings))` samples were described as farming or rural. In these same countries, the proportion of the population residing in urban settings was `r round(mean(annotations$SP.URB.TOTL.ZS, na.rm=T))`%, suggesting that samples were not representative of their countries in terms of rural versus urban settings either.

We then turned to the additional factors, starting with how varied language backgrounds were. A total of 62 different languages or language combinations (for bilingual and multilingual children) were reportedly spoken in the corpora. The samples in which only one language was reported spoke `r mono_langs`, and the  samples in which multiple languages were reported spoke `r multi_langs`. 

About a third (`r round(sum(annotations$Bilingualism.Multilingualism.in.corpus=="yes" & !is.na(annotations$Bilingualism.Multilingualism.in.corpus)) /sum(!is.na(annotations$Bilingualism.Multilingualism.in.corpus))*100)`%) of the included corpora that had available data for this variable (N = `r sum(!is.na(annotations$Bilingualism.Multilingualism.in.corpus))`) were not monolingual. It is hard to find reliable estimates of the percentage of the population who are not monolingual in the world or in the countries represented in CHILDES, but for instance, in Europe in 2016, 65% of adults reported knowing multiple languages (Eurostat, 2022). According to such estimates, even if samples are linguistically diverse, it would appear that input data in CHILDES under-represents bilinguals and multilinguals.

<!-- https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Foreign_language_skills_statistics -->

As for family structure, `r  sum(annotations$Household.structure=="nuclear" & !is.na(annotations$Household.structure))` corpora (`r round(sum(annotations$Household.structure=="nuclear" & !is.na(annotations$Household.structure))/sum(!is.na(annotations$Household.structure))*100)`% of those that had data for this variable)  were based on nuclear families; in `r  sum(annotations$Household.structure=="extended" & !is.na(annotations$Household.structure))`  extended families were portrayed; and in `r  sum(annotations$Household.structure=="varied" & !is.na(annotations$Household.structure))` one sample the structure was varied. We do not know of a country-level index that would allow us to check whether CHILDES corpora are representative of their countries for this variable.



A majority of corpora in CHILDES include children who have siblings. In fact, only `r round(sum(annotations$Proportion.With.Siblings==0 & !is.na(annotations$Proportion.With.Siblings))/sum(!is.na(annotations$Proportion.With.Siblings))*100)`% of  samples  (among the `r sum(!is.na(annotations$Proportion.With.Siblings))` corpora having information on siblings) were constituted exclusively by children with no siblings, and the remaining had at least one sibling, with the overall average being `r round(mean(annotations$Average.number.of.siblings,na.rm=T),2)` siblings. Since `r round(sum(annotations$oecd)/length(annotations$oecd)*100)`% of countries in CHILDES are in the OECD, we draw a comparison point for such countries: `r round(mean(nkids$single,na.rm=T))`% of children had no siblings in OECD countries according to 2015 data. In this sense, children with multiple siblings appear to be over-represented in CHILDES. 



## Discussion


### Code to reproduce Table 4

```{r tab4, eval=F}

annotations$ed_all<-as.character(annotations$Education.ac)
annotations$ed_all[is.na(annotations$ed_all)]<-"NA"
annotations$ed_all=factor(annotations$ed_all,levels=c("Some primary","Some secondary","Some college","College and above","NA"))

annotations$sib_all<-ifelse(annotations$Proportion.With.Siblings==0,"None","1+")
annotations$sib_all[is.na(annotations$sib_all)]<-"NA"
annotations$sib_all=factor(annotations$sib_all,levels=c("None","1+","NA"))

table(annotations$ed_all,annotations$sib_all)
```


## SM1:Partial correlations of the four dimensions of interest

```{r confounds2, echo=FALSE, out.height="200%", fig.cap="Evidence that values in the multidimensional WEIRD complex are only partially correlated (continued). Violin plots of continuous variables as a function of the two discrete variables: Western and type of political regime. Education is represented by proportion of the population completing lower secondary school; industrialization by percentage of the population living in urban (as opposed to rural) sites; richness by GDP per capita. In addition, we show women's average total fertility.", echo=F,message=F, warning=F, results='hide'}


ed <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, Compl.LS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% compl. LSS") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

urb <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("% urban") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

rich <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +  ylab("GDP (log 10)") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())


fert <- ggplot(ind_all[ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(western, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Western") +  ylab("Fert. (log 2)") 




ed_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t) & ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, Compl.LS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

urb_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, SP.URB.TOTL.ZS)) + geom_violin() + geom_jitter(height = 0, width = 0.1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

rich_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, log_gdp)) + geom_violin() + geom_jitter(height = 0, width = 0.1)   +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


fert_d <- ggplot(ind_all[!is.na(ind_all$Political.regime_t)& ind_all$Population>10^6 & !is.na(ind_all$Population),], aes(Political.regime_t, log2_fert)) + geom_violin() + geom_jitter(height = 0, width = 0.1) + xlab("Political regime") +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

x = ggarrange( ed, ed_d,  rich,rich_d,   
          urb, urb_d,   fert, fert_d,
           ncol = 2, nrow = 4)
x
pdf("Fig1.pdf")
x
dev.off()
```


